---
title: "Species-level amacrine cell analysis"
output: 
  BiocStyle::html_document:
    toc: true
date: "2023-06-01"
params:
  species: Chicken_reclustered
  initial: TRUE
  batch_int: TRUE
  integrate_by: "animal"
  harmony: TRUE
  contamination_threshold: 6
  contamination: 
  nFeature_threshold: -2.5
  doublet_finder: TRUE
  manual_annotation: 
  sac_annotation: FALSE
  de_expression: TRUE
  save: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("../../utils/dario_functions.R")

# Load libraries
LoadLibraries(load.lisi = TRUE)
SourceFiles()

# Orthologous genes
# orthology_all_others = readRDS("../../utils/orthology_all_others_v4.rds")

# Booleans to control which analyses to perform
INITIAL = params$initial
BATCH_INT = params$batch_int
DOUBLET_FINDER = params$doublet_finder
SAC_ANNOTATION = params$sac_annotation
DE_EXPRESSION = params$de_expression
SAVE = params$save
REPROCESS = TRUE

# File paths to read and write
initial_filepath = paste0("../../Species_Objects/", params$species, "_initial.rds")
amacrine_filepath = paste0("../../Species_Objects/", params$species, "AC_v1.rds")
batchcor_filepath = paste0("../../Species_Objects/", params$species, "AC_v2_seurat.rds")
harmony_filepath = paste0("../../Species_Objects/", params$species, "AC_v2_harmony.rds")
doublet_filepath = paste0("../../Species_Objects/", params$species, "AC_v3.rds")
filtered_filepath = paste0("../../Species_Objects/", params$species, "AC_v4.rds")
final_filepath = paste0("../../Species_Objects/", params$species, "AC_v5.rds")
markers_filepath = paste0("../../Species_Markers/", params$species, "AC_markers.rds")
annotation_filepath = paste0("../../Species_Annotations/", params$species, "AC_anno.rds")
```

Check parameters

```{r}
params
```

# Extract amacrines from full retina object

```{r, eval = INITIAL}
# Read in full object
species_initial = readRDS(initial_filepath)

# Print object to show number of cells and features
message(paste0(capture.output(species_initial), collapse = "\n"))
```

## Convert genes to common mouse symbols if necessary:

Exceptions: 

* Not for rhabdomys and treeshrew since we used custom genomes and not ENSEMBL genomes
* Not for human, as this caused issues with underscores
* Custom conversion for lamprey from StringTie annotations (courtesy of Yirong Peng)

Original features are found in species_initial@misc$orig.features

```{r, eval = INITIAL}
if(params$species %in% c("Macaque", "Marmoset", "Peromyscus", "Squirrel",
                         "Opossum", "Cow", "Sheep", "Lizard", "Mouse", 
                         "Pig", "Chicken", "Ferret", "Zebrafish", 
                         "MouseLemur", "Rat", "Mouse", "Lamprey")){
  species_initial = ConvertGeneSymbols(species_initial, "../../Orthology/martMergeRefHuman.txt", params$species) 
  
  # Check that original feature names were saved correctly and that no genes were lost
  stopifnot(length(rownames(species_initial@assays$RNA@counts)) == length(species_initial@misc$orig.features))
  
} else if(params$species %in% c("Goldfish")){
  # goldfish_key = fread("../../Orthology/ZF_LA_SB_symbol_for_Seurat.txt")
  # new_genes = goldfish_key$V3[match(rownames(object), goldfish_key$V7)]
  # object@misc$orig.features = rownames(object@assays$RNA@counts)
  # rownames(object@assays$RNA@counts) = new_genes
  # rownames(object@assays$RNA@data) = new_genes
  
  # Goldfish has whole genome duplication so all symbols are duplicated
  species_initial = ConvertGeneSymbols(species_initial, "../../Orthology/martMergeRefHuman.txt", params$species, make.unique = FALSE) 
  metadata = species_initial@meta.data
  
  # Aggregate duplicated genes
  mydat = as.data.frame(as.matrix(species_initial@assays$RNA@counts))
  mydat$gene = rownames(species_initial)
  mydat.sum <- aggregate(. ~ gene, data = mydat, sum)
  
  # Check that it worked
  mydat[mydat$gene == "rbfox1",30:40]
  mydat.sum[mydat.sum$gene == "rbfox1",30:40]
  
  # Transfer metadata
  rownames(mydat.sum) = mydat.sum$gene
  species_initial2 <- CreateSeuratObject(mydat.sum[,colnames(mydat.sum) != "gene"])
  species_initial2 = TransferMetadata(from = species_initial, to = species_initial2)
  species_initial = species_initial2
  
  # Change meta features or will cause issues downstream
  # species_initial[["RNA"]]@meta.features <- data.frame(row.names = rownames(species_initial[["RNA"]]))
}
```

## Inspect object to find SACs

```{r, fig.height=2.5, fig.width = 15, eval = INITIAL}
if(params$species %in% c("Macaque")) species_initial$seurat_clusters = species_initial$annotated
Idents(species_initial) = "seurat_clusters"
species_initial = UpperCase_genes(species_initial)
DotPlot(species_initial, features = rev(c("PAX6", "TFAP2A", "TFAP2B", "CHAT", "SLC5A7"))) + coord_flip() + RotatedAxis()
```

```{r, fig.height=4, fig.width=15, eval = INITIAL}
# For previously analyzed datasets, we will use annotated instead of seurat clusters
# if(params$species == "Chicken") species_initial$seurat_clusters = species_initial$annotated

# Fails for chicken and zebrafish because seurat_clusters are not always in the same cell class
if(params$species == "Chicken"){
  # species_initial$cell_class = ifelse(startsWith(as.character(species_initial$annotated), "AC-"), "AC", "Other")
  # DotPlot(species_initial, features = rev(Genes(major_annotation))) + coord_flip()
  
  # Convert to major cell class
  species_initial$cell_class = as.character(species_initial$cell_class)
  species_initial$cell_class[species_initial$cell_class == "GabaAC" | species_initial$cell_class == "GlyAC"] = "AC"
  species_initial$cell_class[species_initial$cell_class == "BP"] = "BC"
  # species_initial$cell_class[!species_initial$cell_class %in% Annotation(major_annotation)] = "Other"
  species_initial$cell_class[species_initial$cell_class == "MicroG"] = "Other"
  
  AnnotatedUmap(species_initial, major_annotation_custom, group.by = "cell_class", color.clusters.by = "cell_class")
  
  # plot_grid(
  #   DimPlot(species_initial, group.by = "cell_class"), 
  #   AnnotatedDotPlot(species_initial, major_annotation, group.by = "cell_class", color.clusters.by = "cell_class", features = rev(Genes(major_annotation)), color.genes = FALSE) + coord_flip() + RotatedAxis(),
  # rel_widths = c(1,2.5))
  
  plot_grid(
      DimPlot(species_initial, group.by = "cell_class"),
      AnnotatedDotPlot(species_initial, major_annotation, group.by = "cell_class", color.clusters.by = "cell_class",
                       features = rev(Genes(major_annotation)), color.genes = FALSE) + coord_flip() + RotatedAxis(),
      rel_widths = c(1,2.5)
    )
  
} else if(params$species %in% c("Zebrafish", "Goldfish")){
  DotPlot(species_initial, features = rev(Genes(major_annotation))) + coord_flip()
} else {
  
  # Convert to major cell class
  species_initial$cell_class = as.character(species_initial$cell_class)
  species_initial$cell_class[species_initial$cell_class == "GabaAC" | species_initial$cell_class == "GlyAC"] = "AC"
  species_initial$cell_class[species_initial$cell_class == "BP"] = "BC"
  # species_initial$cell_class[!species_initial$cell_class %in% Annotation(major_annotation)] = "Other"
  species_initial$cell_class[species_initial$cell_class == "MicroG"] = "Other"

  # Check if cones and rods are in same cluster
  if(max(apply(table(species_initial$cell_class, species_initial$seurat_clusters), 2, function(x) length(which(x != 0)))) > 1){
    species_initial$cell_class[species_initial$cell_class == "Rod" | species_initial$cell_class == "Cone"] = "Rod"
  }
  
  # Default palette for DimPlot
  palette = c("cyan", "chartreuse2", "red", "gold", "magenta", "grey", "blue", "darkred")
  if(!"Cone" %in% unique(species_initial$cell_class)) palette = c("cyan", "chartreuse2", "gold", "magenta", "grey", "blue", "darkred")
  
  # Plot
  plot_grid(
  {if(params$species != "Macaque") DimPlot(species_initial, group.by = "cell_class", cols = palette)}, #c("darkred", "red", "gold", "chartreuse2", "cyan", "blue", "magenta", "grey")
  AnnotatedDotPlot(species_initial, major_annotation, group.by = "seurat_clusters", color.clusters.by = "cell_class", 
                 features = rev(Genes(major_annotation)), color.genes = FALSE) + coord_flip() + RotatedAxis(),
  rel_widths = c(1,2.5))
}
```

## Recover mislabeled ACs if necessary

```{r, eval = INITIAL}
if(params$species == "Human") species_initial$cell_class[species_initial$seurat_clusters %in% c(5, 14, 24)] = "AC"
if(params$species == "Lizard") species_initial$cell_class[species_initial$seurat_clusters %in% c(18)] = "AC"
if(params$species == "Zebrafish") species_initial$cell_class[species_initial$seurat_clusters %in% c(9,18,24)] = "AC"
if(params$species == "Opossum") species_initial$cell_class[species_initial$seurat_clusters %in% c(1,29,30)] = "AC"
if(params$species == "Peromyscus") species_initial$cell_class[species_initial$seurat_clusters %in% c(14,48)] = "AC"
if(params$species == "Cow") species_initial$cell_class[species_initial$seurat_clusters %in% c(8,9,20)] = "AC"
if(params$species == "Lizard_ncbi") species_initial$cell_class[species_initial$seurat_clusters %in% c(21)] = "AC"

# This seemed to be leading to more doublet-enriched clusters downstream so avoid these three clusters
# if(params$species == "TreeShrew") species_initial$cell_class[species_initial$seurat_clusters %in% c(38,39,48)] = "AC"
```

Re-plot if recovered types

```{r, fig.height=8, fig.width=30, eval = INITIAL}
if(params$species %in% c("Human", "Lizard", "Zebrafish", "Opossum", "Peromyscus", "Cow", "Lizard_ncbi")){
  plot_grid(
    {if(params$species != "Macaque") DimPlot(species_initial, group.by = "cell_class", cols = palette)}, #c("darkred", "red", "gold", "chartreuse2", "cyan", "blue", "magenta", "grey")
    AnnotatedDotPlot(species_initial, major_annotation, group.by = "seurat_clusters", color.clusters.by = "cell_class", 
                   features = rev(Genes(major_annotation)), color.genes = FALSE) + coord_flip() + RotatedAxis(),
  rel_widths = c(1,2.5))
}
```

## Create separate species AC object

```{r, eval = INITIAL}
Idents(species_initial) <- "cell_class"

# Subset amacrines. For human, we also subset features because there's so many (60k) which made objects very large
if(params$species %in% c("Human", "TreeShrew")){
  speciesAC <- CreateSeuratObject(
  GetAssayData(
    subset(species_initial, cells = WhichCells(species_initial, idents = c("AC"))), 
    slot="counts", assay="RNA"), 
  min.cells = 3)
  speciesAC$orig.file = subset(species_initial, cells = WhichCells(species_initial, idents = c("AC")))$orig.file
} else if(params$species == "Goldfish"){
  species_initial$orig.file = species_initial$orig.ident
  speciesAC = subset(species_initial, cells = WhichCells(species_initial, idents = c("GabaAC", "GlyAC")))
} else {
  speciesAC = subset(species_initial, cells = WhichCells(species_initial, idents = c("AC")))
}

# Rename empty gene name in Chicken file as this causes problems downstream
if(params$species == "Chicken"){
  rownames(speciesAC@assays$RNA@counts)[rownames(speciesAC@assays$RNA@counts) == ""] <- "EMPTY"
  rownames(speciesAC@assays$RNA@data)[rownames(speciesAC@assays$RNA@data) == ""] <- "EMPTY" 
}

# Add donor information for human
if(params$species == "Human") speciesAC$donor = subset(species_initial, cells = WhichCells(species_initial, idents = c("AC")))$donor

# Add orig.file information for macaque
if(params$species == "Macaque"){
  Idents(speciesAC) <- "orig.ident"
  speciesAC@meta.data$tag <- speciesAC@meta.data$orig.ident
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S1"), "tag"] <- "M1PerCD90S1"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S2"), "tag"] <- "M1PerCD90S2"
  # speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S3"), "tag"] <- "M1PerCD90S3" # This ident is missing from the object...
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S4"), "tag"] <- "M1PerCD90S4"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S5"), "tag"] <- "M2PerCD90S1"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S6"), "tag"] <- "M2PerCD90S2"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S7"), "tag"] <- "M3PerCD90S1"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S8"), "tag"] <- "M3PerCD90S2"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD90S9"), "tag"] <- "M3PerCD90S3"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD73-S1"), "tag"] <- "M1PerCD73-S1"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD73-S2"), "tag"] <- "M1PerCD73-S2"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD73-S3"), "tag"] <- "M2PerCD73-S1"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD73-DP2S1"), "tag"] <- "M4PerCD73-S1"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaqueCD73-DP2S2"), "tag"] <- "M4PerCD73-S2"
  speciesAC@meta.data[WhichCells(speciesAC, idents = "MacaquePeripheryMixedS1"), "tag"] <- "M2PerMixedS1"
  speciesAC$orig.file = speciesAC$tag
}

# Remove full object and save amacrine subset
rm(species_initial)

# Print object to show number of cells and features
message(paste0(capture.output(speciesAC), collapse = "\n"))

# Save
saveRDS(speciesAC, amacrine_filepath)
```

# Quality control 

```{r}
speciesAC = readRDS(amacrine_filepath)
```

## Remove channels that contain the class of interest at a frequency less than 60 cells

```{r}
batch_totals = table(speciesAC$orig.file)
t(t(batch_totals))
cat("Total number of ACs: ", sum(batch_totals))
remove_batches = names(which(batch_totals < 60))

# Custom removal of certain batches in some species due to poor integration 
if(params$species == "Rhabdomys") remove_batches = c(remove_batches, "RhabNeunPos1b1", "RhabNeunPos1b2", "RhabNeunPos2b1") 

# Remove from peromyscus since too few cells compared to other biological replicates
# if(params$species == "Peromyscus") remove_batches = c(remove_batches, "Pero1CD73-dpS1", "Pero1CD73-dpS2", "Pero1CD73-dpS3")
# if(params$species == "Peromyscus") remove_batches = c(remove_batches, "Pero1CD90S1", "Pero1CD90S2", "Pero2CD73-dpS1", "Pero2CD73-dpS2", "Pero3CD73-dpS2") 

if(length(remove_batches) > 0) {
  speciesAC = subset(speciesAC, orig.file %in% remove_batches, invert = TRUE)
  message("\nRemoving batches: ", paste0(remove_batches, collapse = ", "))
}
```

## Filter out cells with abnormally low/high numbers of unique molecules

Look for mitochondrial genes 

Calculate mitochondrial fraction

```{r}
rownames(speciesAC)[startsWith(toupper(rownames(speciesAC)), "MT-")]
speciesAC[["percent.mt"]] <- PercentageFeatureSet(speciesAC, pattern = "^MT-")
```

Plot distributions

```{r, fig.width = 6, fig.height=3}
if(params$species == "Peromyscus"){
  feature.low = 2000 
# } else if(params$species == "Goldfish"){
#   feature.low = 200 
} else {
  feature.low = max(mean(speciesAC$nFeature_RNA) - 3*sd(speciesAC$nFeature_RNA), 400)
}

feature.high = min(mean(speciesAC$nFeature_RNA) + 3*sd(speciesAC$nFeature_RNA), 6000)
mt.high = 20

plot_grid(PrettyHistogram(speciesAC$nFeature_RNA, 
                          vline = c(feature.low, feature.high)), 
          BINplot(speciesAC, 
                  feature.high = feature.high, 
                  feature.low = feature.low, 
                  percent.mt.high = mt.high), 
          ncol = 2)
```

Remove cells

```{r}
message("Removing ", length(which(speciesAC$nFeature_RNA < feature.low)), " cells with nFeatures < ", feature.low)
message("Removing ", length(which(speciesAC$nFeature_RNA > feature.high)), " cells with nFeatures > ", feature.high)
message("Removing ", length(which(speciesAC$percent.mt > mt.high)), " cells with percent.mt > ", mt.high)

speciesAC <- subset(speciesAC, subset = percent.mt < mt.high & nFeature_RNA < feature.high & nFeature_RNA > feature.low)
```

# Batch integration

Assign batch information for species

```{r}
# Read in species metadata
species_metadata = read.csv("../../Species_Objects/sample_metadata2.csv")
# colnames(species_metadata) = c("channel", "description")

if(params$species == "Human") {
  # speciesAC$region = "Peri"; speciesAC$region[grepl("OSm|macula|Mix", speciesAC$orig.file)] = "Mac"
  # speciesAC$enrichment = "NONE"; speciesAC$enrichment[grepl("NeuN|PRet|RGC", speciesAC$orig.file)] = "NeuN+"
  # speciesAC$batch = paste0(speciesAC$region, "_", speciesAC$donor)
  # speciesAC$animal = speciesAC$donor
  speciesAC$region = toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 3)[,1])
  speciesAC$enrichment = toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 3)[,2])
  speciesAC$animal = gsub("A", "", gsub("B", "", gsub("C", "", toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 3)[,3]))))
  speciesAC$donor = speciesAC$animal
} else if(params$species == "Goldfish"){
  speciesAC$enrichment = "NONE"
  speciesAC$animal = speciesAC$orig.ident
} else if(params$species == "Macaque") {
  # Region 
  speciesAC$region = "Periphery"
  speciesAC$region[grepl("Fovea", speciesAC$orig.ident, fixed = TRUE)] = "Fovea"
  
  # Animal
  speciesAC$animal = substr(speciesAC$tag, 0, 2)
  speciesAC$animal[startsWith(speciesAC$tag, "M1Per")] = "M5"
  speciesAC$animal[startsWith(speciesAC$tag, "M2Per")] = "M6"
  speciesAC$animal[startsWith(speciesAC$tag, "M3Per")] = "M7"
  
  # Enrichment type
  speciesAC$enrichment = "NONE"
  speciesAC$enrichment[grepl("CD90|Mixed", speciesAC$tag)] = "CD90+"
  speciesAC$enrichment[grepl("CD73-", speciesAC$tag)] = "CD73-"
  
  # speciesAC$batch = paste0(speciesAC$region, "_", speciesAC$animal, "_", speciesAC$enrichment)
  speciesAC$batch = paste0(speciesAC$region, "_", speciesAC$animal)
} else if(params$species == "Marmoset") {
  speciesAC$region = toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 3)[,1])
  speciesAC$enrichment = toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 3)[,2])
  speciesAC$animal = gsub("A", "", gsub("B", "", gsub("C", "", toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 3)[,3]))))
} else {
  speciesAC$enrichment = toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 2)[,1])
  speciesAC$animal = gsub("A", "", gsub("B", "", gsub("C", "", toupper(str_split_fixed(species_metadata$description[match(speciesAC$orig.file, species_metadata$channel)], "_", 2)[,2]))))
}

addmargins(table(speciesAC$enrichment, speciesAC$orig.file))
addmargins(table(speciesAC$animal, speciesAC$orig.file))

# Dispersion calculation
sd(table(speciesAC$animal))/mean(table(speciesAC$animal))
sd(table(speciesAC$orig.file))/mean(table(speciesAC$orig.file))
```

## Check for batch effects

```{r, fig.width=15, fig.height=7}
speciesAC = ClusterSeurat(speciesAC, cluster_resolution = 1.5)
plotBatches2(speciesAC, batch = 'orig.file')
```

Cluster each batch separately to compute celltype LISI later on

```{r}
speciesAC$batch_clusters = ClusterBatches(speciesAC, batch = params$integrate_by, resolution = 1.5)
```

Check for batch effects using integration LISI and cell type LISI

```{r, fig.width=10, fig.height=6}
# iLISI computation is for all cells together
iLISI_before <- compute_lisi(speciesAC@reductions$pca@cell.embeddings[,1:20], # speciesAC@reductions$umap@cell.embeddings
                              speciesAC@meta.data, 
                              c(params$integrate_by))

# cLISI computation is for each batch since we don't have cell type info a priori
cLISI_before = ComputeCelltypeLISI(speciesAC, batch = params$integrate_by)
```

## Tabulate batches used for integration

```{r}
t(t(table(speciesAC[[params$integrate_by]])))
```

## Seurat integration

```{r, eval = BATCH_INT}
# Save orig.features which would be lost during integration
orig.features = speciesAC@misc$orig.features

# Integrate
speciesAC = ClusterSeurat(speciesAC, integrate.by = params$integrate_by, cluster_resolution = 1.5)

# Preserve orig.features
speciesAC@misc$orig.features = orig.features

# Save
saveRDS(speciesAC, batchcor_filepath)
```

Confirm seurat integration was successful

```{r, fig.width=15, fig.height=7}
speciesAC = readRDS(batchcor_filepath)
plotBatches2(speciesAC, batch = params$integrate_by)
```

Check for batch effects using integration LISI and celltype LISI

```{r, fig.width=10, fig.height=6}
iLISI_seurat <- compute_lisi(speciesAC@reductions$pca@cell.embeddings[,1:20], 
                              speciesAC@meta.data, 
                              c(params$integrate_by))

cLISI_seurat = ComputeCelltypeLISI(speciesAC, batch = params$integrate_by)
```

## Harmony integration

```{r, eval = BATCH_INT, fig.width=15, fig.height=6}
DefaultAssay(speciesAC) = "RNA"
speciesAC = Harmonize(speciesAC, batch = params$integrate_by, cluster_resolution = 1.5)
saveRDS(speciesAC, harmony_filepath)
```

Confirm harmony integration was successful

```{r, fig.width=15, fig.height=7}
speciesAC = readRDS(harmony_filepath)
plotBatches2(speciesAC, batch = params$integrate_by)
```

Check for batch effects using integration LISI and celltype LISI

```{r, fig.width=10, fig.height=6}
# Use harmony embeddings for iLISI
iLISI_harmony <- compute_lisi(speciesAC@reductions$harmony@cell.embeddings, 
                              speciesAC@meta.data, 
                              c(params$integrate_by))

cLISI_harmony = ComputeCelltypeLISI(speciesAC, batch = params$integrate_by, harmony = TRUE)
```

## Summarize batch correction metrics

```{r, fig.height=6, fig.width=8}
par(mfrow = c(1,2))
iLISI = data.frame(before = iLISI_before[[params$integrate_by]], 
                   seurat = iLISI_seurat[[params$integrate_by]], 
                   harmony = iLISI_harmony[[params$integrate_by]])
boxplot(iLISI)
cLISI = data.frame(before = cLISI_before$batch_clusters, 
                   seurat = cLISI_seurat$batch_clusters, 
                   harmony = cLISI_harmony$batch_clusters)
boxplot(cLISI)
```

## Choose batch correction method

```{r, fig.width=14, fig.height=8}
if(params$harmony){
  message("Using Harmony integration")
  speciesAC = readRDS(harmony_filepath)
} else {
  message("Using Seurat integration")
  speciesAC = readRDS(batchcor_filepath)
}
```

Quickly check which clusters were enriched by FACs

```{r, fig.width=14, fig.height=8}
# CD90 is Thy1, CD73 is NT5E, SLC17A6 is VGLUT2, RBFOX3 is NEUN, VSX2 is CHX10 
enrichment_markers = c("THY1", "NT5E", "SLC17A6", "RBFOX3", "VSX2", "ISL2B")
FeaturePlot(speciesAC, features = enrichment_markers, ncol = 3)

VlnPlot(speciesAC, 
        features = enrichment_markers, 
        group.by = "enrichment", pt.size = 0.1, 
        ncol = 3)
```

# Filter out contamination, low quality clusters, and doublets

## Check for contamination from other cell classes

```{r, fig.height=6, fig.width=15}
AnnotatePlot(DotPlot(speciesAC, features = rev(Genes(amacrine_annotation))) + coord_flip() + RotatedAxis(), 
             amacrine_annotation, color_genes_x = TRUE)
```

```{r, fig.height=6, fig.width=15}
AnnotatePlot(DotPlot(speciesAC, features = rev(Genes(major_annotation))) + coord_flip() + RotatedAxis(), 
             major_annotation, color_genes_x = TRUE)
```

```{r, eval = FALSE}
# For interactive assignment
speciesAC@misc$contamin.clusters = list(rod = NA, cone = c(38), hc = c(NA), bc = c(13,18), ac = NA, rgc = NA, mg = NA)
```

```{r, fig.height=5, fig.width=15}
if(FALSE){
  speciesAC = ScaleData(speciesAC, features = Genes(major_annotation))
  
  gs_list = list()
  gs_list$gs_positive = list(Rod = setdiff(Genes(major_annotation)[Annotation(major_annotation) == "Rod"], "RHO"), 
                             Cone = Genes(major_annotation)[Annotation(major_annotation) == "Cone"],
                             HC = "LHX1", 
                             BC = setdiff(Genes(major_annotation)[Annotation(major_annotation) == "BC"], "GRIK1"), 
                             AC = Genes(major_annotation)[Annotation(major_annotation) == "AC"], 
                             RGC = Genes(major_annotation)[Annotation(major_annotation) == "RGC"], 
                             MG = setdiff(Genes(major_annotation)[Annotation(major_annotation) == "MG"], "SLC1A3"))
  gs_list$gs_negative = list(Rod = NULL, 
                             Cone = NULL,
                             HC = NULL, 
                             BC = NULL, 
                             AC = NULL, 
                             RGC = NULL, 
                             MG = NULL)
  
  speciesAC = ScTypeAnnotation(speciesAC, gs_list)
  
  
} else {
  # speciesAC = ProfileContamination(speciesAC, z.score.threshold = params$contamination_threshold)
  if(!is.null(params[["contamination"]])) speciesAC@misc$contamin.clusters = params[["contamination"]]
  contamin.clusters = sort(unique(unlist(speciesAC@misc$contamin.clusters)))
  message("Contaminated clusters: ", paste0(contamin.clusters, collapse = ", "))
}

# AnnotatedDotPlot(speciesAC, major_annotation, group.by = "seurat_clusters", color.clusters.by = "customclassif", 
#                features = rev(Genes(major_annotation)), color.genes = FALSE) + coord_flip() + RotatedAxis()
```

## Check for low quality clusters

```{r, fig.height=12, fig.width = 20}
BrowseSeurat(speciesAC, batch = params$integrate_by)
```

```{r}
lowq.clusters = which(scale(MeanMetadata(speciesAC, feature.1 = "seurat_clusters", feature.2 = "nFeature_RNA")) < params$nFeature_threshold) - 1
message("Low quality clusters: ", paste0(lowq.clusters, collapse = ", "))
```

## Doublet detection with DoubletFinder

```{r, eval = DOUBLET_FINDER, fig.width = 12, fig.height=10}
speciesAC = FindDoublets(speciesAC, channels = "orig.file")
saveRDS(speciesAC, doublet_filepath)
```

Check quality of doublet detection

```{r, fig.width = 12, fig.height=10}
speciesAC = readRDS(doublet_filepath)
DoubletAnalysis(speciesAC)
```

## Remove contaminated and/or low quality clusters and putative doublets

```{r}
if(length(lowq.clusters) > 0) lowq.cells = WhichCells(speciesAC, idents = lowq.clusters) else lowq.cells = NULL
if(length(contamin.clusters) > 0) contamin.cells = WhichCells(speciesAC, idents = contamin.clusters) else contamin.cells = NULL
doublet.cells = WhichCells(speciesAC, expression = DF.classifications == "Doublet")

message("Found ", length(lowq.cells), " low-quality cells")
message("Found ", length(contamin.cells), " contaminated cells")
message("Found ", length(doublet.cells), " putative doublet cells")
message("Removing ", length(unique(c(lowq.cells, contamin.cells, doublet.cells))), " unique cells in total")

# Remove contaminating clusters
speciesAC = subset(speciesAC, cells = c(lowq.cells, contamin.cells, doublet.cells), invert = TRUE)
```

# Re-process after filtration

```{r, eval = REPROCESS}
if(params$harmony){
  speciesAC = Harmonize(speciesAC, batch = params$integrate_by, cluster_resolution = 1.5)
  saveRDS(speciesAC, filtered_filepath)
} else{
  speciesAC = ReprocessIntegrated(speciesAC)
  saveRDS(speciesAC, filtered_filepath)
}
```

```{r, fig.width=15, fig.height=6}
speciesAC = readRDS(filtered_filepath)
plotBatches2(speciesAC, batch = params$integrate_by)
```

## Cluster-resolution curve
Resolution parameter that is most stable based on the cluster-resolution curve

```{r}
if(params$harmony) {
  plotClusterRes(speciesAC)
} else {
  DefaultAssay(speciesAC) = "integrated"
  plotClusterRes(speciesAC)
}
```

Re-cluster if necessary

```{r}
# if(params$species == "Pig") speciesAC = FindClusters(speciesAC, resolution = 1.5)
# if(params$species == "Ferret") speciesAC = FindClusters(speciesAC, resolution = 1.5)
# if(params$species == "TreeShrew") speciesAC = FindClusters(speciesAC, resolution = 2)
```

## Cluster dendrogram 

Check cluster similarity based on tree estimated from distance matrix constructed in PCA space

```{r, fig.width=10, fig.height=5}
DefaultAssay(speciesAC) = "RNA"
speciesAC = FindVariableFeatures(speciesAC, selection.method = "vst", nfeatures = 2000)
speciesAC = BuildClusterTree(object = speciesAC)
PlotClusterTree(speciesAC)
```

## Merge clusters

```{r}
speciesAC$old_seurat_clusters = speciesAC$seurat_clusters
speciesAC = mergeClusters(speciesAC)
```

 Check that merging was successful

```{r, fig.width=10, fig.height=10}
if(length(unique(speciesAC$old_seurat_clusters)) != length(unique(speciesAC$seurat_clusters))) {
  
  # Recompute cluster tree if clustering changed
  speciesAC = BuildClusterTree(object = speciesAC)
  
  # Plot confusion matrix
  plotConfusionMatrix(table(speciesAC$seurat_clusters, speciesAC$old_seurat_clusters), stagger.threshold = 1)
}
```

Visualize final clusters

```{r, fig.height=12, fig.width = 20}
BrowseSeurat(speciesAC, batch = params$integrate_by)
```

Remove any remaining low quality clusters

```{r}
# if(params$species == "TreeShrew") speciesAC = subset(speciesAC, idents = 40, invert = TRUE)
```

Plot final clusters by batch

```{r, fig.height=6, fig.width=15}
plotBatches2(speciesAC, batch = params$integrate_by)
```

# GABAergic vs glycinergic classification

Markers:

* SLC32A1 is transporter required for both GABA and glycine uptake into synaptic vesicles
* Slc6a1, Slc6a13,and Slc6a11 encode GABA transporter types 1, 2, and 3, respectively
* Gad1 and Gad2 encode two forms of glutamate decarboxylase
* Slc6a9 and Slc6a5 encode vesicular glycine transporters 1 and 2, respectively

## GAD vs GlyT1 expression

For treeshrew, GAD1 has low counts due to poor annotation, so we use SLC6A1

```{r, fig.width=12, fig.height=4}
Gaba.gene.1 = ifelse(params$species == "TreeShrew", "SLC6A1", "GAD1")
Gaba.gene.2 = ifelse(params$species == "TreeShrew", "SLC6A11", "GAD2")
speciesAC = GabaGlyClassification(speciesAC, Gaba.gene.1 = Gaba.gene.1, Gaba.gene.2 = Gaba.gene.2)
order = arrange(Metadata(speciesAC, feature.1 = "seurat_clusters", feature.2 = "classification"), classification)$seurat_clusters
speciesAC$ac.order = factor(speciesAC$seurat_clusters, levels = order)
```

```{r}
t(t(table(speciesAC$classification)))
```

## Clustered DotPlot

```{r, fig.height=4, fig.width=15}
ClusteredDotplot(speciesAC, 
                 # dendrogram = phylogram::as.dendrogram(speciesAC@tools$BuildClusterTree), 
                 order = "ac.order",
                 features = toupper(c("Slc32a1", "Slc6a1", "Slc6a13", "Slc6a11", "Gad1", "Gad2", "Slc6a9", "Slc6a5")))
```

# Manual annotation of cell types

## Ensure all contamination was removed 

```{r, fig.height=6, fig.width=15}
AnnotatePlot(DotPlot(speciesAC, features = rev(Genes(major_annotation))) + coord_flip() + RotatedAxis(), 
             major_annotation, color_genes_x = TRUE)
``` 

## Plot amacrine cell type markers

```{r, fig.height=5.5, fig.width=15}
# AnnotatePlot(DotPlot(speciesAC, features = rev(Genes(amacrine_annotation))) + coord_flip() + RotatedAxis(), 
#              amacrine_annotation, color_genes_x = TRUE)

AnnotatePlot(
  OrderedDotPlot(speciesAC, amacrine_annotation, group.by = "seurat_clusters", color.clusters.by = "seurat_clusters", order = order, features = rev(c(Gaba.gene.1, Gaba.gene.2, "SLC6A9", Genes(amacrine_annotation)))), 
  amacrine_annotation, color_genes_x = TRUE) + coord_flip() + RotatedAxis()
```

## Cluster annotations based on marker expression

manual_annotation = list(SAC = c(2), A17 = c(NA), nNOS = c(63), CA1 = c(17), CA2 = c(NA), VG1 = c(NA), A2 = c(1), SEG = c(14,22), VG3 = 40)

```{r}
# manual_annotation = eval(parse(text = params$manual_annotation))
manual_annotation = params$manual_annotation
```

```{r}
# Assign annotations
speciesAC = AssignAnnotations(speciesAC, manual_annotation)

# Concatenate cluster and lit type
speciesAC = AssignName(speciesAC)

t(t(table(speciesAC$type)))
```

Annotated types to one side

```{r, fig.height=5.5, fig.width=15}
order = arrange(Metadata(speciesAC, feature.1 = "type", feature.2 = "classification", feature.3 = "lit_type"), classification, lit_type)$type
AnnotatePlot(
  OrderedDotPlot(speciesAC, amacrine_annotation, group.by = "type", color.clusters.by = "type", order = order, features = rev(c(Gaba.gene.1, Gaba.gene.2, "SLC6A9", Genes(amacrine_annotation)))), 
  amacrine_annotation, object_y = speciesAC, y_annotation = "lit_type") + coord_flip() + RotatedAxis()
```

## ON/OFF SAC annotation

```{r}
# SAC_ANNOTATION = "SAC" %in% speciesAC$lit_type
```

```{r, fig.height=8, fig.width=15, eval = SAC_ANNOTATION}
speciesSAC = subset(speciesAC, subset = lit_type == "SAC")

# speciesSAC = ClusterSeurat(speciesSAC)

ON_SAC_markers = c("FEZF1", "GFRA3", "CNTN5", "SLIT2")
OFF_SAC_markers = c("TENM3", "ZFHX3", "RND3")

plot_grid({if(any(ON_SAC_markers %in% rownames(speciesSAC))) FeaturePlot(speciesSAC, features = ON_SAC_markers, ncol = length(ON_SAC_markers), reduction = "umap")}, 
          {if(any(OFF_SAC_markers %in% rownames(speciesSAC))) FeaturePlot(speciesSAC, features = OFF_SAC_markers, ncol = length(OFF_SAC_markers),  reduction = "umap")}, 
            ncol = 1, nrow = 2)
```

```{r, fig.height=5, fig.width=5, eval = SAC_ANNOTATION}
speciesSAC$on.sac.markers = PercentageFeatureSet(speciesSAC, features = ON_SAC_markers)
speciesSAC$off.sac.markers = PercentageFeatureSet(speciesSAC, features = OFF_SAC_markers)

plot_grid(
  gghistogram(speciesSAC@meta.data, x = "on.sac.markers", xlab = FALSE),
  NULL, 
  ggscatter(speciesSAC@meta.data, x = "on.sac.markers", "off.sac.markers"), 
  gghistogram(speciesSAC@meta.data, x = "off.sac.markers", ylab = FALSE) + coord_flip(), 
  ncol = 2, nrow = 2, rel_widths = c(2,1), rel_heights = c(1,2), align = "hv")
```

## Compare our clustering to previous annotations, if they exist

```{r, fig.height=8, fig.width=10}
if("annotated" %in% colnames(speciesAC@meta.data)){
  # plotConfusionMatrix(table(speciesAC$annotated, speciesAC$type))
  JSHeatmap(JSMatrix(t(table(speciesAC$annotated, speciesAC$type))), heatmap = FALSE, stagger.threshold = 0.05)
}
```

# Effects of enrichment procedure

Exclude zebrafish from analysis since all samples used were enriched for Isl2b
Old: Exclude rhabdomys (only using neun- and none)

```{r}
ENRICHED = "enrichment" %in% colnames(speciesAC@meta.data) & !params$species %in% c("Zebrafish", "Goldfish", "Chicken")
```

```{r, fig.width=16, fig.height=4, eval = ENRICHED}
# Get enrichment groups and corresponding markers
enrichment_groups = toupper(unique(speciesAC$enrichment)[!unique(speciesAC$enrichment) %in% c("NONE", "ALL", "CD73-")])
markers = (gsub("-", "", gsub("\\+", "", enrichment_groups)))
markers[markers == "NEUN"] = "RBFOX3"
markers[markers == "CD90"] = "THY1"
markers[markers == "CHX10"] = "VSX2"

# Subset to markers whose gene is present in the dataset
enrichment_groups = enrichment_groups[markers %in% rownames(speciesAC)]
markers = markers[markers %in% rownames(speciesAC)]

# Plot UMAPs showing enriched populations
lapply(seq_along(markers), function(index){
  plot_grid(FeaturePlot(speciesAC, features = markers[[index]]) + NoLegend(), 
          DimPlot(speciesAC, split.by = "enrichment", group.by = "enrichment") + NoLegend(),
          VlnPlot(speciesAC, group.by = "enrichment", features = markers[[index]], pt.size = 0),
          ncol = 3, 
          rel_widths = c(1, length(unique(speciesAC$enrichment)), 1.5))
})
```

## Identify cell types affected by enrichment procedure

```{r, fig.width=10, fig.height=4, eval = ENRICHED}
lapply(seq_along(markers), function(index){
  StackedPlots(list(stackedBarGraph(speciesAC, feature.1 = "enrichment", feature.2 = "type") + NoLegend(), 
          VlnPlot(speciesAC, features = markers[[index]], pt.size = 0, group.by = "type") + NoLegend()))
})
```

## Correlation between enriched marker expression and proportion of cells from enriched collections

```{r, fig.width=6, fig.height=6, eval = ENRICHED}
lapply(seq_along(markers), function(index){
  enriched.expression = as.data.frame(t(log1p(AverageExpression(speciesAC, features = c(markers[[index]], "PAX6"), group.by = "type")$RNA)))
  prop.enriched = proportionTable(speciesAC, feature.1 = "enrichment", feature.2 = "type")[,enrichment_groups[[index]] ]
  enriched.expression$prop.enriched = prop.enriched
  enriched.expression$cluster = as.character(rownames(enriched.expression))
  
  ScatterPlot(enriched.expression, 
              x = "prop.enriched", 
              y = markers[[index]], 
              ylab = paste0(markers[[index]], " expression"), 
              xlab = paste0("Proportion from ", enrichment_groups[[index]], " samples"), 
              label = "cluster")
})
```

If not enriched, denote proportion enriched to all zero

```{r, eval = !ENRICHED}
prop.enriched = rep(0, length(unique(speciesAC$seurat_clusters)))
if(params$species == "Zebrafish") prop.enriched = rep(1, length(unique(speciesAC$seurat_clusters)))
```

# Regional enrichment for primates

Identify cell types with different proportions in fovea or periphery (only in unenriched samples)

```{r, fig.width=15, fig.height=6}
if(params$species %in% c("Human", "Macaque", "Marmoset")){
  plotBatches2(speciesAC, batch = "region", clusters = "type")
}

# barplot((table(temp$region, temp$type)/rowSums(table(temp$region, temp$type))), beside=T, las = 2, legend = TRUE)
```

# Find DE markers

```{r, eval = DE_EXPRESSION}
# Using faster Rcpp implementation provided through presto 
speciesACmarkers = wilcoxauc(speciesAC, 'seurat_clusters')
saveRDS(speciesACmarkers, markers_filepath)
```

## Peptide annotation

```{r}
neuropeptides = read.table("../../data/complete_neuropeptide_symbols.txt")$V1
speciesACmarkers = readRDS(markers_filepath)

padj.threshold = 0.001
logFC.threshold = 1

peptide_markers = lapply(sort(unique(speciesAC$seurat_clusters)), function(cluster){
                    markers = subset(speciesACmarkers, group == cluster & 
                             logFC > logFC.threshold & 
                             padj < padj.threshold)$feature
                    return(markers[markers %in% toupper(neuropeptides)])
                  })
names(peptide_markers) = sort(unique(speciesAC$seurat_clusters))
peptides_found = sort(unique(unlist(peptide_markers)))
```

Neuropeptide genes from HUGO (http://www.neuropeptides.nl/tabel%20neuropeptides%20linked.htm)

```{r, fig.height=10, fig.width=15}
if(length(peptides_found) > 0) {
  # ClusteredDotplot(speciesAC, dendrogram = phylogram::as.dendrogram(speciesAC@tools$BuildClusterTree), features = peptides_found)
  DotPlot(speciesAC, group.by = "ac.order", features = peptides_found) + coord_flip()
}
```

# Summarize annotations

```{r}
species_anno = SeuratAnnotation(speciesAC, peptide.list = peptide_markers)
                                # prop.enriched = prop.enriched)
species_anno
saveRDS(species_anno, annotation_filepath)
```

Save final annotated object

```{r, eval = SAVE}
saveRDS(speciesAC, final_filepath)
```

```{r}
sessionInfo()
```

