---
title: "Literature proportion analysis"
output: 
  BiocStyle::html_document:
    toc: true
date: "2023-07-28"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../../utils/dario_functions.R")
LoadLibraries()
SourceFiles()
```

# Read in data

```{r, fig.height=4, fig.width=8}
speciesList = phylogenetic_order[phylogenetic_order %in% c("Human", "Ferret", "Lizard", "Rhabdomys", "Marmoset", "Lamprey", "Macaque", "Chicken", "MouseLemur", "Rat", "Pig", "Peromyscus")] #setdiff(unique(vertebrateAC$species), c("Mouse", "Rhabdomys", "Sheep", "Squirrel", "TreeShrew"))
objectList = lapply(speciesList, function(species) readRDS(paste0("../../Species_Objects/", species, "AC_v5.rds")))
names(objectList) = speciesList

# Make sure enrichment group is properly assigned
species_metadata = read.csv("../../Species_Objects/sample_metadata2.csv")
objectList = lapply(seq_along(objectList), function(index) {
  object = objectList[[index]]
  if(!names(objectList)[index] %in% c("Human", "Macaque", "Marmoset")){
    object$enrichment = toupper(str_split_fixed(species_metadata$description[match(object$orig.file, species_metadata$channel)], "_", 2)[,1])
    object$animal = gsub("A", "", gsub("B", "", gsub("C", "", toupper(str_split_fixed(species_metadata$description[match(object$orig.file, species_metadata$channel)], "_", 2)[,2]))))
  }
  return(object)
})
names(objectList) = speciesList
```

# Validation of cell type proprotions using literature

A2	Human	10-15 of total
A2	Macaque	10-15 of total
A2	Marmoset	10-15 of total
KT2	Macaque	5 of total
A8	Macaque	3-5 of total
VG3	Rat	1 of total
A17	Mouse	5 of total
CA1	Human	0.1
CA1	Macaque	0.1
nNOS Marmoset	2.5

```{r, eval = prop.differences}
humanAC = readRDS("../../Species_Objects/HumanAC_v6.rds")
# humanAC = objectList$Human
human.props = getProportions(humanAC)

macaqueAC = readRDS("../../Species_Objects/MacaqueAC_v6.rds")
# macaqueAC = objectList$Macaque
macaque.props = getProportions(macaqueAC)

marmosetAC = readRDS("../../Species_Objects/MarmosetAC_v6.rds")
# marmosetAC = objectList$Marmoset
marmoset.props = getProportions(marmosetAC)

ratAC = readRDS("../../Species_Objects/RatAC_v6.rds")
# ratAC = objectList$Rat
rat.props = getProportions(ratAC)

# mouseAC = readRDS("../../Species_Objects/MouseAC_v6.rds")
mouseAC = readRDS("../../Species_Reference/MouseACref_v4.rds")
mouseAC$enrichment = NA
mouseAC$enrichment[mouseAC$source %in% c("Tran", "Macosko")] = "NONE"
mouse.props = getProportions(mouseAC)

# Add mouse to objectList
# objectList$Mouse = mouseAC
```

```{r lit-prop-scatter, eval = prop.differences, fig.height=3.5, fig.width=3.5}
# A2 (human 3)
human.A2 = human.props[[1]]["3"]

# A2 (macaque 15)
macaque.A2 = macaque.props[[1]]["15"]

# A2 (marmoset 5)
marmoset.A2 = marmoset.props[[1]]["5"]

# Knotty type 2 (macaque 38)
macaque.KT2 = macaque.props[[1]]["38"]

# A8 (macaque 4)
macaque.A8 = macaque.props[[1]]["4"]

# VG3 (rat 35)
rat.VG3 = rat.props[[1]]["35"]

# CA1 (human 21)
human.CA1 = human.props[[1]]["21"]

# CA2 (macaque 43)
macaque.CA1 = macaque.props[[1]]["43"]

# nNOS (marmoset 21)
marmoset.nNOS = marmoset.props[[1]]["24"]

# SAC (mouse 7)
mouse.SAC = mouse.props[[1]]["15"]

# A17 (mouse 10)
mouse.A17 = mouse.props[[1]]["4"]

literatureProportions = data.frame(type_species = c("A2-human", "A2-macaque", "A2-marmoset", "KT2-macaque", "A8-macaque", "VG3-rat", 
                                                    "A17-mouse", "CA1-human", "CA1-macaque", "nNOS-marmoset", "SAC-mouse"), 
                                   `IHC` = c(0.10, 0.10, 0.10, 0.05, 0.04, 0.01, 0.05, 0.001, 0.001, 0.025, 0.052), 
                                   `scRNAseq` = c(human.A2, macaque.A2, marmoset.A2, macaque.KT2, macaque.A8, rat.VG3, 
                                                  mouse.A17, human.CA1, macaque.CA1, marmoset.nNOS, mouse.SAC))

literatureProportions = subset(literatureProportions, type_species != "KT2-macaque")

# ScatterPlot(literatureProportions,
#             x = "IHC",
#             y = "scRNAseq",
#             labels = "type_species",
#             xlab = "Literature (% total)",
#             ylab = "scRNA-seq (% total)",
#             title = "Cell type proportions", 
#             lm = FALSE, r = TRUE, r_pval = FALSE)+
#   geom_abline(slope=1, intercept = 0)+
#   scale_x_continuous(limits = c(0,11))+
#   scale_y_continuous(limits = c(0,11))

ggplot(literatureProportions, aes(x = IHC, y = scRNAseq)) +
  theme_cowplot()+
  # geom_point(
  #   data = data.frame(x = seq(0,0.1,length.out = 100), 
  #                     y = seq(0,0.1,length.out = 100)),
  #   aes(x, y),
  #   alpha = 1,        # invisible
  #   size = 3
  # )+
  geom_abline(slope=1, intercept = 0, color = 'lightgrey', alpha = 1, linewidth = 0.5)+
  geom_point(color = '#584B9FFF', fill = '#584B9FFF', shape = 21, size = 2.5)+
  xlab('IHC frequency')+
  ylab('sc/snRNA-seq frequency')+
  geom_text_repel(aes(label = type_species), 
                  size = 3, 
                  min.segment.length = 0.2)+
                  # nudge_x = ifelse(literatureProportions$IHC > literatureProportions$scRNAseq, 0.002, -0.002))+
  # scale_x_continuous(limits = c(0,0.11))+
  # scale_y_continuous(limits = c(0,0.11))+
  scale_ticks_first_last(expand = c(0.01, 0.01), limits = c(0, 0.12))+
  scale_ticks_first_last(expand = c(0.01, 0.01), axis = 'x', limits = c(0, 0.12))+
  stat_cor(aes(label = ..r.label..))+
  ArialFont()
  # coord_cartesian(clip = "off") +
  # theme_dario()

ggsave('../../figures/my_figs/lit-prop-scatter.pdf', height=3.3, width=3.5)
# p1
# dev.off()
```


```{r, fig.height=3, fig.width=6}
# No need to read in orthotype object at the moment
# vertebrateAC = readRDS("../../Ortho_Objects/vertebrateAC_v2.rds")

# Compute proportions for each object
proportionsList = lapply(c(objectList), getProportions)
nCells = sapply(proportionsList, function(x) x[2])
data.frame(nCells = unlist(nCells))
```


```{r, fig.height=3, fig.width=6}
# Extract proportion vectors
proportions = sapply(proportionsList, function(x) x[1])

# Convert literature names to new versions
objectList = lapply(objectList, function(object) {
  object$lit_type[object$lit_type == "AII"] = "A2" 
  object$lit_type[object$lit_type == "CA-I"] = "CA1" 
  object$lit_type[object$lit_type == "CA-II"] = "CA2" 
  return(object)
})
```

## SAC proportions

```{r}
# Find best match for SACs
PlotCelltypeProportions(objectList, proportions, "SAC", color = "red", reference.line = 5.2)
```

We do observe fewer SACs in mouse lemur. All pig and opossum SACs originated from neun-enriched samples, suggesting these low numbers were due to a wide NeuN gate that depleted SACs from the un-enriched samples. Cow SACs are neun+ but in general SACs are neun-, but may be present in neun enriched samples due to a wide gate used in FACS.   

## A2 proportions

```{r}
objectList$Lizard$lit_type[objectList$Lizard$seurat_clusters == 18] = "A2"
PlotCelltypeProportions(objectList, proportions, "A2", color = "skyblue", reference.line = 10)
```

```{r}
PlotCelltypeProportions(objectList, proportions, "VG3", color = "magenta", reference.line = 1)
```

```{r}
sessionInfo()
knitr::knit_exit()
```

## Proportion differences between fovea and periphery

Only possible in human because marmoset and macaque have no un-enriched peripheral 
samples. However, we can't compare between periphery and fovea because there is no 
reference. The region is confounded by the dissection. Before we were comparing 
the number of type cells to the total number of cells. We would have to compare as 
proportion of total cells profiled, but even then different channels will obtain
different numbers of cells. 

```{r, fig.width=15, fig.height=6, eval = FALSE}
plotBatches2(subset(humanAC, enrichment == "NONE"), batch = "region", clusters = "type")
# plotBatches2(subset(macaqueAC, enrichment == "NONE"), batch = "region", clusters = "type")
# marmosetAC = AssignName(marmosetAC)
# plotBatches2(subset(marmosetAC, enrichment == "NONE"), batch = "region", clusters = "type")

table(subset(humanAC, enrichment == "NONE")$region, subset(humanAC, enrichment == "NONE")$type)
# table(subset(macaqueAC, enrichment == "NONE")$region, subset(macaqueAC, enrichment == "NONE")$type)
# table(subset(marmosetAC, enrichment == "NONE")$region, subset(marmosetAC, enrichment == "NONE")$type)
```

## Compare proportions

```{r}

```

