---
title: "Single-cell cluster robustness analysis"
output: 
  BiocStyle::html_document:
    toc: true
date: "2023-07-28"
params:
  filepath: "../../Ortho_Objects/vertebrateAC_v2_100.rds"
  nPermutations: 50
  nCores: 1
  method: "seurat"
  group.by: "species"
  proportion: 0.8
  run.perm: TRUE
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../../utils/dario_functions.R")
LoadLibraries(load.lisi = FALSE)
SourceFiles()
# options(ggplot2.discrete.colour = list(dtColors()), 
        # ggplot2.discrete.fill = list(dtColors()))
```

```{r}
params
```

# Introduction

Here we will be analyzing the reproducibility of our single cell clusters by varying the key clustering parameters: 

* The cells supplied 
* The number of PCs used
* The number of neighbors in the KNN graph 
* The resolution parameter

Our metrics for determining cluster reproducibility are: 

* Metric 1: the distribution of types across subsamples clusters, measured by Shannon entropy 
* Metric 2: the tendency of cells from a type to cluster with cells of other types, measured as the fraction of other cells in the cluster

# Example: walk-through one iteration

## Load data

```{r}
seurat = readRDS(params$filepath)

# DefaultAssay(seurat) = 'integrated'
# seurat = FindClusters(seurat, resolution = 1.25)
# DefaultAssay(seurat) = 'RNA'
```

## Resample data

```{r}
# Set seed for reproducibility
set.seed(12345)

system.time({ 
  sample1 = SubsampleCluster(seurat, 
                             sample(colnames(seurat), size = floor(length(Cells(seurat))*params$proportion), replace=F), 
                             nPCs = 15, 
                             k.param = 18, 
                             resolution = 1, 
                             method = params$method, 
                             group.by = params$group.by)
})
```

## Visualize similarity

```{r, fig.width=8, fig.height=6}
JSHeatmap(JSMatrix(table(sample1$old_seurat_clusters, sample1$seurat_clusters)), stagger.threshold = 0.10, heatmap = FALSE)
adj.rand.index(sample1$old_seurat_clusters, sample1$seurat_clusters)
```

## Compute metrics 

Metric 1 (co-clustering): 

1. For the proportion of cells from a type across the k newly computed clusters, $p_1, p_2, ..., p_k$, we compute the Shannon entropy: $H_{T} = -\sum_i p_i log_2 p_i$. 
2. The minimum is bounded by 0 (least disordered) when all cells of a type are in a single cluster: $H_{T} = - (1 log_2 1 + 0 log_20 + ... + 0 log_20) = 0$. When a type is equally distributed across the k clusters (most disordered), $p_i = \frac{1}{k}$ so the maximum is bounded by: $-k(\frac{1}{k} log_2 \frac{1}{k}) = log_2 k$. Thus, we take $\frac{-\sum_i p_i log_2 p_i}{log_2 k}$ to normalize between 0 and 1. 
3. In order to have higher scores represent more co-clustering, we take one minus this value: $1+\frac{\sum_i p_i log_2 p_i}{log_2 k}$. 


Metric 2 (promiscuity): 

1. For each cell $c_1, c_2, ..., c_n$, from a type, we find the fraction of cells, $f_1, f_2, ..., f_n$, in their cluster that is from cells other than their type. 
2. We summarize this score as the average fraction: $\frac{1}{n}\sum_i^n f_i$. Since all the cells that are in the same cluster will have the same fraction, we can rewrite the formula as the proportion of other cells in each cluster, scaled by the number of cells $N_i$ from a type in that cluster: $\frac{1}{n}\sum_i^k N_i(1-p_i)$. This value will be zero when there are no cells from that type in cluster i ($N_i=0$), and when all the cells in cluster i are from that type ($1-p_i=0$). Note that this proportion $p_i$ is different from the previous one, since it is a proportion of the cluster rather than a proportion of the type. 
3. Since higher fractions represent more promiscuity, we subtract this value from 1 so that higher scores represent less promiscuity: $1 - \frac{1}{n}\sum_i^k N_i(1-p_i)$. 


Summarize in a data frame

```{r, fig.height=3.5, fig.width=8}
df.test = data.frame(Type = factor(levels(seurat$seurat_clusters), levels = levels(seurat$seurat_clusters)), 
                     M1 = ComputeM1(sample1$old_seurat_clusters, sample1$seurat_clusters), 
                     M2 = ComputeM2(sample1$old_seurat_clusters, sample1$seurat_clusters))

df.test = reshape2::melt(df.test)
colnames(df.test) = c("Type", "Metric", "Score")
df.test$Group = "True"
```

## Shuffled type labels as control

```{r, fig.width=8, fig.height=6}
# Shuffle type labels to generate control set
sample1.shuffled = sample1
sample1.shuffled$old_seurat_clusters = sample(as.integer(sample1$old_seurat_clusters))

df.rand = data.frame(Type = factor(levels(seurat$seurat_clusters), levels = levels(seurat$seurat_clusters)), 
                     M1 = ComputeM1(sample1.shuffled$old_seurat_clusters, sample1.shuffled$seurat_clusters), 
                     M2 = ComputeM2(sample1.shuffled$old_seurat_clusters, sample1.shuffled$seurat_clusters))

df.rand = reshape2::melt(df.rand)
colnames(df.rand) = c("Type", "Metric", "Score")
df.rand$Group = "Shuffled"

JSHeatmap(JSMatrix(table(sample(sample1.shuffled$old_seurat_clusters), sample1.shuffled$seurat_clusters)), stagger.threshold = 0.10, heatmap = FALSE)
```

```{r, fig.height=6, fig.width=15}
df = rbind(df.test, df.rand)
df$Type = as.character(df$Type)

p1 = ggbarplot(subset(df, Metric == "M1"), x = "Type", y = "Score",
          fill = "Group", position = position_dodge()) +
  ylab("Co-clustering score (M1)")+
  theme(legend.position = "right") + 
  scale_y_continuous(expand = c(0,0))+
  RotatedAxis()

p2 = ggbarplot(subset(df, Metric == "M2"), x = "Type", y = "Score",
          fill = "Group", position = position_dodge()) +
  ylab("1 - promiscuity score (M2)")+
  theme(legend.position = "right") + 
  scale_y_continuous(expand = c(0,0))+
  RotatedAxis()

ggarrange(p1, p2, nrow = 2, common.legend = T, legend = "right")
```

# Run analysis 

```{r}
RunSubsamplingParallel = function(object, nPermutations = 50, fraction.use = 0.8, shuffle.types = FALSE, nCores = 1){
  
  # Sample cells
  subsample.size = floor(length(Cells(object))*fraction.use)
  rand.cells.use = lapply(seq_len(nPermutations), function(x) sample(colnames(object), size = subsample.size, replace=F))
  
  # Sample parameters
  rand.nPCs = sample(c(15:30), nPermutations, replace = TRUE)
  rand.k = sample(c(16:24), nPermutations, replace = TRUE)
  rand.res = sample(seq(1, 2, by = 0.1), nPermutations, replace = TRUE)

  # Make cluster
  cl <- parallel::makeCluster(
    nCores, 
    type = "PSOCK"
    )
  
  # register it to be used by %dopar%
  doParallel::registerDoParallel(cl = my.cluster)
  
  # Export functions and objects
  clusterExport(cl=cl, list("SubsampleCluster", "rand.cells.use", "rand.nPCs", "rand.k",
                          "rand.res", "shuffle.types", "params"),
              envir=environment())

  # Parallel loop
  clusters = foreach(iteration = seq_len(nPermutations)) %dopar% {
    message("Running iteration #", iteration, " with the following params: \n nPCs: ", rand.nPCs[[iteration]], "\n k.param: ", rand.k[[iteration]], "\n res: ", rand.res[[iteration]])
    
    sample = suppressMessages(SubsampleCluster(object, 
                                               cells.use = rand.cells.use[[iteration]],
                                               nPCs = rand.nPCs[[iteration]], 
                                               k.param = rand.k[[iteration]], 
                                               resolution = rand.res[[iteration]], 
                                               shuffle.types = shuffle.types, 
                                               method = params$method))
    
    return(list(original = sample$old_seurat_clusters, permuted = sample$seurat_clusters))
  }
  
  stopCluster(cl)
  
  # Name by permutation
  names(clusters) = seq_len(nPermutations)
  
  return(list(clusters = clusters, 
              rand.cells.use = rand.cells.use, 
              rand.nPCs = rand.nPCs, 
              rand.k = rand.k, 
              rand.res = rand.res))
}
```

## Run permutations

```{r, eval = params$run.perm}
system.time({ 
  perm.data = RunSubsampling(seurat, 
                             nPermutations = params$nPermutations, 
                             fraction.use = params$proportion,
                             nCores = 1, 
                             method = params$method, 
                             group.by = params$group.by)
})

saveRDS(perm.data, paste0("permuted_clusters/", gsub(".rds", "", basename(params$filepath)), "_permdata.rds"))

# message("--- Now in parallel with future --- ")
# 
# library(future)
# plan("multiprocess", workers = 30)
# 
# system.time({ 
#   perm.data = RunSubsampling(seurat, 
#                              nPermutations = params$nPermutations, 
#                              fraction.use = params$proportion,
#                              nCores = 1)
# })
```

Parallelization makes it go slower!

## Calculate scores

```{r}
perm.data = readRDS(paste0("permuted_clusters/", gsub(".rds", "", basename(params$filepath)), "_permdata.rds"))

scores = lapply(perm.data$clusters, function(clusters){
  
    # Object "seurat" has original clusters
    m1.scores = ComputeM1(clusters$original, clusters$permuted)
    m2.scores = ComputeM2(clusters$original, clusters$permuted)
    
    # Shuffle type labels to generate control set
    random_seurat_clusters = sample(clusters$original)
    m1.random = ComputeM1(random_seurat_clusters, clusters$permuted)
    m2.random = ComputeM2(random_seurat_clusters, clusters$permuted)
    
    return(list(m1.scores = m1.scores, 
                m2.scores = m2.scores, 
                m1.random = m1.random, 
                m2.random = m2.random))
})
```

## Plot results

```{r, fig.height=6, fig.width=15}
# M1 scores
m1.scores = do.call(rbind, lapply(scores, function(x) x$m1.scores))
colnames(m1.scores) = levels(seurat$seurat_clusters)

df.m1test = reshape2::melt(m1.scores)
colnames(df.m1test) = c("Iteration", "Type", "Score")
df.m1test$Group = "True"

m1.random = do.call(rbind, lapply(scores, function(x) x$m1.random))
colnames(m1.random) = levels(seurat$seurat_clusters)

df.m1rand = reshape2::melt(m1.random)
colnames(df.m1rand) = c("Iteration", "Type", "Score")
df.m1rand$Group = "Shuffled"

df.m1 = rbind(df.m1test, df.m1rand)
df.m1$Metric = "M1"
df.m1$Type = as.character(df.m1$Type)

# M2 scores
m2.scores = do.call(rbind, lapply(scores, function(x) x$m2.scores))
colnames(m2.scores) = levels(seurat$seurat_clusters)

df.m2test = reshape2::melt(m2.scores)
colnames(df.m2test) = c("Iteration", "Type", "Score")
df.m2test$Group = "True"

m2.random = do.call(rbind, lapply(scores, function(x) x$m2.random))
colnames(m2.random) = levels(seurat$seurat_clusters)

df.m2rand = reshape2::melt(m2.random)
colnames(df.m2rand) = c("Iteration", "Type", "Score")
df.m2rand$Group = "Shuffled"

df.m2 = rbind(df.m2test, df.m2rand)
df.m2$Metric = "M2"
df.m2$Type = as.character(df.m2$Type)

p1 = ggbarplot(df.m1, x = "Type", y = "Score",
          fill = "Group", position = position_dodge(), add = c("mean_se")) +
  ylab("Co-clustering score (M1)")+
  theme(legend.position = "right") + 
  scale_y_continuous(expand = c(0,0))+
  RotatedAxis()

p2 = ggbarplot(df.m2, x = "Type", y = "Score",
          fill = "Group", position = position_dodge(), add = c("mean_se")) +
  ylab("1 - promiscuity score (M2)")+
  theme(legend.position = "right") + 
  scale_y_continuous(expand = c(0,0))+
  RotatedAxis()

ggarrange(p1, p2, nrow = 2, common.legend = T, legend = "right")
```

M2 scores have a tendency of being smaller for smaller clusters because small clusters might not be resolved when fewer cells are sampled or when the resolution parameter is lower. 

# Adjusted rand indices across permutations

```{r}
ari.df = data.frame(Permutation = names(perm.data$clusters), 
                      ARI = sapply(perm.data$clusters, function(x) adj.rand.index(x$original, x$permuted)))

ggbarplot(ari.df, x = "Permutation", y = "ARI", position = position_dodge(), fill = "grey") +
  ylab("Adjusted rand index (ARI)")+
  scale_y_continuous(expand = c(0,0), limits = c(0,0.91))+
  RotatedAxis()
```

# Average jaccard of each type

```{r}

```

# Permutation plot

Adapted from Langfelder and Horvath 2012

```{r}
library(WGCNA)

nCells = ncol(seurat)
nRuns = params$nPermutations
labels = matrix(0, nRuns + 1, nCells)

ref_clusters = as.numeric(as.character(seurat$seurat_clusters))
ref_barcodes = names(seurat$seurat_clusters)

labels[1,] = ref_clusters

# Match labels from each resampling to the labels obtained for the full data set
# This may also take some time. Hence, the results are saved and next time the user can skip this part and
# load the data from disk. 

for(r in 2:(nRuns+1)){
  this_barcodes = names(perm.data$clusters[[r-1]]$original)
  converted = matchLabels(as.numeric(as.character(perm.data$clusters[[r-1]]$permuted))+1,
                          as.numeric(as.character(perm.data$clusters[[r-1]]$original))+1)-1
  labels[r,] = converted[match(ref_barcodes, this_barcodes),]
}

rownames(labels) = c("Ref", paste0("Perm", 1:nRuns))

# Sort by dendro.order
if(startsWith(basename(params$filepath), "vertebrateAC")){
  dendro = CorrelationHeatmap(seurat, group.by = 'seurat_clusters', return.dendrogram = TRUE)
  seurat$dendro.order = factor(seurat$seurat_clusters, levels = labels(dendro))
  dendro.order = levels(seurat$dendro.order)
  sorted.matrix = labels[,order(factor(labels[1,], levels = dendro.order))]
} else {
  seurat = ClusterDendrogram(seurat, assay = "RNA")
  dendro.order = labels(phylogram::as.dendrogram(seurat@tools$BuildClusterTree))
  sorted.matrix = labels[,order(factor(labels[1,], levels = dendro.order))]
}

# Relabel un-mapped cell modules to -1 (grey) but keep NAs as NAs (white)
sorted.matrix[!sorted.matrix %in% unique(ref_clusters) & !is.na(sorted.matrix)] = -1

# Account for zero based indexing
zeroIndex = any(seurat$seurat_clusters == 0)
if(zeroIndex) {
  colors = structure(labels2colors_dt(0:(max(ref_clusters)+1), omit = "white"), names = as.character(-1:max(ref_clusters)))
} else {
  colors = structure(labels2colors_dt(0:(max(ref_clusters)), omit = "white"), names = as.character(c(-1, 1:max(ref_clusters))))
}
```

```{r, fig.height=15, fig.width=20}
# pdf('../../figures/my_figs/orthotype-reproducibility.pdf', height=8, width=20)

# Row and column splits
row_km = seq_along(rownames(sorted.matrix)) %>% setNames(rownames(sorted.matrix))
col_km = factor(sorted.matrix[1,], levels = dendro.order)

# Label clusters
midpoints = sapply(dendro.order, function(cluster) floor(mean(which(sorted.matrix[1,] == cluster))))
top_ha = columnAnnotation(foo = anno_mark(at = midpoints, 
                                      labels = as.numeric(dendro.order), 
                                      labels_rot = 0, 
                                      link_height = unit(7,"mm"), 
                                      # link_width = unit(3,"mm"), 
                                      link_gp = gpar(col="black", lty = 1, lwd = 1)))

# Parameter annotation
# right_ha = rowAnnotation(nPCs = anno_numeric(c(20, perm.data$rand.nPCs), bg_gp = gpar(fill = myPalette(5)[1], col = "white")), 
#                          k = anno_numeric(c(20, perm.data$rand.k), bg_gp = gpar(fill = myPalette(5)[2], col = "white")), 
#                          Res = anno_numeric(c(1.5, perm.data$rand.res), bg_gp = gpar(fill = myPalette(5)[3], col = "white")), 
#                          nClusters = anno_numeric(c(nClusters(seurat), sapply(perm.data$clusters, function(x) length(levels(x$permuted)))), bg_gp = gpar(fill = myPalette(5)[4], col = "white")),
#                          ARI = anno_numeric(c(1, round(ari.df$ARI, 2)), bg_gp = gpar(fill = myPalette(5)[5], col = "white")),
#                          annotation_name_rot = 45, 
#                          width = unit(5, "cm"))

right_ha = rowAnnotation(nPCs = anno_barplot(c(20, perm.data$rand.nPCs), border = FALSE, gp = gpar(fill = myPalette(5)[1], col = "white")), 
                         k = anno_barplot(c(20, perm.data$rand.k), border = FALSE, gp = gpar(fill = myPalette(5)[2], col = "white")),
                         Res = anno_barplot(c(1.5, perm.data$rand.res), border = FALSE, gp = gpar(fill = myPalette(5)[3], col = "white")),
                         nClusters = anno_barplot(c(nClusters(seurat), sapply(perm.data$clusters, function(x) length(levels(x$permuted)))), border = FALSE, gp = gpar(fill = myPalette(5)[4], col = "white")),
                         ARI = anno_barplot(c(1, round(ari.df$ARI, 2)), border = FALSE, gp = gpar(fill = myPalette(5)[5], col = "white")),
                         annotation_name_rot = 45, 
                         width = unit(5, "cm"), 
                         gap = unit(2, "mm"))

# custom_colors = c('grey', colorspace::lighten(myPalette(floor(length(colors)/2)), 0.1), colorspace::lighten(myPalette(floor(length(colors)/2)), 0.3)) %>% setNames(seq(-1, length(colors)-2))

# Heatmap
Heatmap(sorted.matrix, name = "mat2", show_row_names = FALSE, 
        show_column_names = FALSE, na_col = "white", col = colors, 
        cluster_rows = FALSE, cluster_columns = FALSE, row_split = row_km, 
        # column_split = col_km, column_gap = unit(0,"mm"), 
        row_gap = unit(0, "mm"), column_title = NULL, border = FALSE, 
        show_heatmap_legend = FALSE, 
        row_title = names(row_km), row_title_rot = 0, 
        top_annotation = top_ha, right_annotation = right_ha, 
        use_raster = FALSE, 
        raster_quality = 2)

# Add horizontal bars separating rows
for(i_slice in seq_along(unique(row_km))) {
  	decorate_heatmap_body("mat2", row_slice = i_slice, {
  	  grid.lines(c(0, 1), c(0, 0), gp = gpar(lty = 1, lwd = 1))
  	})
}

decorate_heatmap_body("mat2", row_slice = 1, {
  grid.lines(c(1,0), c(1, 1), gp = gpar(lty = 1, lwd = 1))
})

# dev.off()
```

# Analysis of parameters predictive of ARI

```{r}
parameters.df = data.frame(ARI = ari.df$ARI, 
                          nPCs = perm.data$rand.nPCs, 
                          k.param = perm.data$rand.k, 
                          resolution = perm.data$rand.res,
                          nClusters = sapply(perm.data$clusters, function(x) length(levels(x$permuted))))

head(parameters.df)
```


```{r}
# Run lm
model = lm(ARI ~ nPCs + k.param + resolution, data = parameters.df)
summary(model)
```

```{r, fig.height=6, fig.width=6}
library(psych)
pairs.panels(parameters.df, 
             method = "pearson", # correlation method
             hist.col = "#00AFBB",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
             )
```

```{r}
knitr::knit_exit()
```

# Figure for paper

```{r, fig.height=9, fig.width=20}
pdf('../../figures/my_figs/orthotype-reproducibility.pdf', height=9, width=20)

# Row and column splits
row_km = seq_along(rownames(sorted.matrix)) %>% setNames(rownames(sorted.matrix))
col_km = factor(sorted.matrix[1,], levels = dendro.order)

# Label clusters
midpoints = sapply(dendro.order, function(cluster) floor(mean(which(sorted.matrix[1,] == cluster))))
top_ha = columnAnnotation(foo = anno_mark(at = midpoints, 
                                      labels = as.numeric(dendro.order), 
                                      labels_rot = 0, 
                                      link_height = unit(7,"mm"), 
                                      # link_width = unit(3,"mm"), 
                                      link_gp = gpar(col="black", lty = 1, lwd = 1)))

right_ha = rowAnnotation(nPCs = anno_barplot(c(20, perm.data$rand.nPCs), border = FALSE, gp = gpar(fill = myPalette(5)[1], col = "white")), 
                         k = anno_barplot(c(20, perm.data$rand.k), border = FALSE, gp = gpar(fill = myPalette(5)[2], col = "white")),
                         Res = anno_barplot(c(1.5, perm.data$rand.res), border = FALSE, gp = gpar(fill = myPalette(5)[3], col = "white")),
                         nClusters = anno_barplot(c(nClusters(seurat), sapply(perm.data$clusters, function(x) length(levels(x$permuted)))), border = FALSE, gp = gpar(fill = myPalette(5)[4], col = "white")),
                         ARI = anno_barplot(c(1, round(ari.df$ARI, 2)), border = FALSE, gp = gpar(fill = myPalette(5)[5], col = "white")),
                         annotation_name_rot = 45, 
                         width = unit(5, "cm"), 
                         gap = unit(2, "mm"))

custom_colors = c('grey', colorspace::lighten(myPalette(floor(length(colors)/2)), 0.1), colorspace::lighten(myPalette(floor(length(colors)/2)), 0.3)) %>% setNames(seq(-1, length(colors)-2))

# Heatmap
Heatmap(sorted.matrix, name = "mat2", show_row_names = FALSE, 
        show_column_names = FALSE, na_col = "white", col = custom_colors, 
        cluster_rows = FALSE, cluster_columns = FALSE, row_split = row_km, 
        # column_split = col_km, column_gap = unit(0,"mm"), 
        row_gap = unit(0, "mm"), column_title = NULL, border = FALSE, 
        show_heatmap_legend = FALSE, 
        row_title = names(row_km), row_title_rot = 0, 
        top_annotation = top_ha, right_annotation = right_ha, 
        use_raster = TRUE, 
        raster_quality = 2)

# Add horizontal bars separating rows
for(i_slice in seq_along(unique(row_km))) {
  	decorate_heatmap_body("mat2", row_slice = i_slice, {
  	  grid.lines(c(0, 1), c(0, 0), gp = gpar(lty = 1, lwd = 1))
  	})
}

decorate_heatmap_body("mat2", row_slice = 1, {
  grid.lines(c(1,0), c(1, 1), gp = gpar(lty = 1, lwd = 1))
})

dev.off()
```

# Leave one out cross validation

```{r}
vertebrateAC = readRDS('../../../storage/Ortho_Objects/vertebrateAC_v3_100.rds')

dendro = CorrelationHeatmap(vertebrateAC, 
                   group.by = 'seurat_clusters', 
                   cols = c('white', "#922144FF"), 
                   # annotate.by = 'classification', 
                   # annotation_cols = c('GABA' = "#DBAD5DFF", 'Gly' = "#4F4782FF"), 
                   row_names_gp = gpar(fontsize = 6), 
                   show_column_names = FALSE, 
                   return.dendrogram = TRUE)
vertebrateAC$dendro.order = factor(vertebrateAC$seurat_clusters, levels = labels(dendro))

speciesList = setdiff(names(SPECIESFILES), c("Zebrafish", "Goldfish", "Lamprey"))
loo.clustering = lapply(speciesList, function(species) {
  object = readRDS(paste0('../../../storage/Ortho_Objects/cross_validation/', species, '.rds'))
  return(data.frame(cell = Cells(object), seurat_clusters = object$seurat_clusters) %>% setNames(c('cell', species)))
})
saveRDS(loo.clustering, 'loo.clustering.rds')

loo.clustering = readRDS('loo.clustering.rds')
loo.merged = Reduce(function(dtf1, dtf2) merge(dtf1, dtf2, by = "cell", all=T), loo.clustering)
loo.merged$type = vertebrateAC$type[match(loo.merged$cell, Cells(vertebrateAC))]
loo.merged$seurat_clusters = vertebrateAC$seurat_clusters[match(loo.merged$cell, Cells(vertebrateAC))]
loo.merged$species = vertebrateAC$species[match(loo.merged$cell, Cells(vertebrateAC))]

# Order by species
loo.merged = loo.merged %>% arrange(species)

lapply(speciesList, function(species) ConfusionMatrix2(loo.merged$type, loo.merged[[species]]))
lapply(speciesList, function(species) length(unique(loo.merged[[species]])))
```

Adapted from Langfelder and Horvath 2012

```{r}
library(WGCNA)

nCells = ncol(vertebrateAC)
nRuns = length(speciesList)
labels = matrix(0, nRuns + 1, nCells)

ref_clusters = as.numeric(as.character(loo.merged$seurat_clusters))
ref_barcodes = loo.merged$cell
labels[1,] = ref_clusters

for(r in 2:(nRuns+1)){
  this_barcodes = loo.merged$cell
  converted = matchLabels(as.numeric(as.character(loo.merged[,r]))+1,
                          as.numeric(as.character(loo.merged$seurat_clusters))+1)-1
  labels[r,] = converted[match(ref_barcodes, this_barcodes),]
}

rownames(labels) = c("17 species", paste0('- ', speciesList))

# Sort by dendro.order
dendro.order = levels(vertebrateAC$dendro.order)
sorted.matrix = labels[,order(factor(labels[1,], levels = dendro.order))]

# Relabel un-mapped cell modules to -1 (grey) but keep NAs as NAs (white)
sorted.matrix[!sorted.matrix %in% unique(ref_clusters) & !is.na(sorted.matrix)] = -1

# Make dendrogram
# hclust(AvgExpr(seurat, group.by = "seurat_clusters", features = VariableFeatures(seurat), assay = "RNA"))

# Account for zero based indexing
zeroIndex = any(vertebrateAC$seurat_clusters == 0)
if(zeroIndex) {
  colors = structure(labels2colors_dt(0:(max(ref_clusters)+1), omit = "white"), names = as.character(-1:max(ref_clusters)))
} else {
  colors = structure(labels2colors_dt(0:(max(ref_clusters)), omit = "white"), names = as.character(c(-1, 1:max(ref_clusters))))
}
```

```{r, fig.height=5, fig.width=17, dpi = 600}

pdf('../../figures/my_figs/cross-validation.pdf', height=5, width=17)
# Row and column splits
row_km = seq_along(rownames(sorted.matrix)) %>% setNames(rownames(sorted.matrix))
col_km = factor(sorted.matrix[1,], levels = dendro.order)

# Label clusters
midpoints = sapply(dendro.order, function(cluster) floor(mean(which(sorted.matrix[1,] == cluster))))
top_ha = columnAnnotation(foo = anno_mark(at = midpoints, 
                                      labels = dendro.order, 
                                      labels_rot = 0, 
                                      link_height = unit(7,"mm"), 
                                      # link_width = unit(3,"mm"), 
                                      link_gp = gpar(col="black", lty = 1, lwd = 1)))

# Parameter annotation
# right_ha = rowAnnotation(#nPCs = anno_numeric(c(20, perm.data$rand.nPCs), bg_gp = gpar(fill = "red", col = "white")),
#                          #k = anno_numeric(c(20, perm.data$rand.k), bg_gp = gpar(fill = "yellow", col = "white")),
#                          #Res = anno_numeric(c(1.5, perm.data$rand.res), bg_gp = gpar(fill = "green", col = "white")),
#                          nClusters = anno_numeric(c(nClusters(seurat), sapply(perm.data$clusters, function(x) length(levels(x$permuted)))), bg_gp = gpar(fill = "blue", col = "white")),
#                          ARI = anno_numeric(c(0, round(ari.df$ARI, 2)), bg_gp = gpar(fill = "purple", col = "white")),
#                          annotation_name_rot = 45,
#                          width = unit(5, "cm"))
right_ha = NULL

custom_colors = c(colorspace::lighten(myPalette(length(colors)/2), 0.1), colorspace::lighten(myPalette(length(colors)/2), 0.3)) %>% setNames(seq(-1, length(colors)-2))
# custom_colors = colorspace::lighten(myPalette(length(colors)), 0.5) %>% setNames(seq(-1, length(colors)-2))

# Heatmap
Heatmap(sorted.matrix, name = "mat2", show_row_names = FALSE, 
        show_column_names = FALSE, 
        na_col = "black", 
        col = custom_colors, 
        cluster_rows = FALSE, 
        cluster_columns = FALSE, 
        row_split = row_km, 
        # column_split = col_km, column_gap = unit(0,"mm"), 
        row_gap = unit(0, "mm"), 
        column_title = NULL, 
        border = FALSE, 
        show_heatmap_legend = FALSE, 
        row_title = names(row_km), 
        row_title_rot = 0, 
        top_annotation = top_ha, 
        right_annotation = right_ha, 
        use_raster = TRUE, 
        raster_quality = 2)

# Add horizontal bars separating rows
for(i_slice in seq_along(unique(row_km))) {
  	decorate_heatmap_body("mat2", row_slice = i_slice, {
  	  grid.lines(c(0, 1), c(0, 0), gp = gpar(lty = 1, lwd = 1))
  	})
}

decorate_heatmap_body("mat2", row_slice = 1, {
  grid.lines(c(1,0), c(1, 1), gp = gpar(lty = 1, lwd = 1))
})

dev.off()
```

```{r}
sessionInfo()
```
