---
title: "TF tree analysis"
output: 
  BiocStyle::html_document:
    toc: true
params:
  nIter: 10
  metadata: TRUE
  save: TRUE
---

Load necessary libraries for analysis

```{r setup}
knitr::opts_chunk$set(dev = 'png', dpi = 300, fig.height = 10, fig.width = 10, 
                      root.dir = '/clusterfs/kslab/Evolution/paper_2/projects/AC') # set output device to svg

source("../../utils/dario_functions.R")
LoadLibraries(load.lisi = FALSE)
SourceFiles()

# Global variables
NCLUSTERS = 42
type_cols = AcPalette2(42) %>% setNames(1:NCLUSTERS)
speciesList = names(SPECIESFILES)
speciesList.no.gf = setdiff(names(SPECIESFILES), 'Goldfish')
speciesList.liz = setdiff(speciesList.no.gf, c('Zebrafish', 'Lamprey')) %>% setNames(setdiff(speciesList.no.gf, c('Zebrafish', 'Lamprey')))
JACCARD.CUTOFF = 0.25
KNOWNTYPES = c('SAC', 'A2', 'VG3') %>% setNames(c('SAC', 'A2', 'VG3'))

# Global files
neuropeptides = read.table("../../reference_files/complete_neuropeptide_symbols.txt")$V1
hk_genes = read.table("../../reference_files/Eisenberg_Cell_2013_hk_genes.txt")$V1
tf_genes = read.csv("../../reference_files/cisbp_20220920.csv")$Name

# Feature plot colors
fcols = c('grey', 'deeppink3')

# Palettes
major_annotation_palette = c(unique(Colors(major_annotation_pr_gaba_gly)) %>% setNames(unique(Annotation(major_annotation_pr_gaba_gly))), 'Other' = 'grey')
major_annotation_palette2 = c(unique(Colors(major_annotation_pr_gaba_gly_on_off)) %>% setNames(unique(Annotation(major_annotation_pr_gaba_gly_on_off))), 'Other' = 'grey')
CELL_CLASSES = setdiff(names(major_annotation_palette), 'Other') %>% 
  setNames(setdiff(names(major_annotation_palette), 'Other'))
CELL_CLASSES2 = setdiff(names(major_annotation_palette2), 'Other') %>% 
  setNames(setdiff(names(major_annotation_palette2), 'Other'))

prefix.key = c(ch = "Chicken", li = "Lizard", ze = "Zebrafish", kf = 'Killifish', mf = 'Macaque', mm = 'Mouse', op = 'Opossum')
NM_SPECIES = c('Chicken', 'Lizard', 'Zebrafish', 'Killifish') %>% setNames(c('Chicken', 'Lizard', 'Zebrafish', 'Killifish'))
nm_palette = species_palette2[NM_SPECIES]
nm_palette2 = nm_palette %>% setNames(prefix.key[match(names(nm_palette), prefix.key)])

ac.ortho = LoadACOrtho()
dend_custom_order = readRDS('../../Ortho_Objects/ac.dendrogram.v2.rds')
# plot(dend_custom_order)
clusters <- cutree(as.hclust(dend_custom_order), k = 10)
type_cols2.no = colorspace::lighten(ClusterPalette(sort(clusters), colors = c('deeppink', 'cyan', 'orange', 'gold', 'chartreuse', '#584B9FFF', 'blue', 'violet', 'tan', 'antiquewhite', 'brown'), vary.by = 0.4), 0.2)
names(type_cols2.no) = as.numeric(names(sort(clusters))) + 1
type_cols2.no = type_cols2.no[order(as.numeric(names(type_cols2.no)))]
type_cols2 = type_cols2.no %>% setNames(levels(ac.ortho$type_unordered))
# type_cols3 = AcPalette3(42, vary.by = 0.9) %>% setNames(levels(ac.ortho$type))
# type_cols3.no = type_cols3 %>% setNames(Metadata(ac.ortho, 'type', 'type_no')$type_no)
# type_cols3.no[names(type_cols3.no) == '20'] = 'tan'
# type_cols3.no = type_cols3 %>% setNames(1:42)

METHODS = c('UPGMA', 'NJ', 'MP', 'ML')
```

# Set up

```{r}
ac.ortho = readRDS("../../Ortho_Objects/vertebrateAC_BIGSEURAT.rds")
```

Steps:

* Select top conserved TFs
* Binarize expression so that each OT is a vector of 1s and 0s of length nTFs
* Compute maximum parsimony tree using Phangorn
* Understand tree logic

## Feature selection

```{r, fig.height=4, fig.width=5}
conservation.summary = readRDS("../../Ortho_Objects/vertebrateAC_conservation_v2.rds")
# conservation.summary$is.tf = conservation.summary$gene %in% tf_genes
cor.score.cutoff = 0.3 #max(conservation.summary$cor.score.rand, na.rm = TRUE)
PrettyHistogram(subset(conservation.summary, is.tf)$cor.score, 
                vline = cor.score.cutoff, 
                xlab = "Conservation score", 
                title = paste0("Conservation scores for ", nrow(subset(conservation.summary, is.tf)), " TFs"))
PrettyHistogram(sample(conservation.summary$cor.score, nrow(subset(conservation.summary, is.tf))), 
                vline = cor.score.cutoff, 
                xlab = "Conservation score", 
                title = paste0("Conservation scores for ", nrow(subset(conservation.summary, is.tf)), " random genes"))

# Significantly more conserved 
t.test(cor.score ~ is.tf, conservation.summary)
```

```{r, fig.height=4, fig.width=9}
ac.ortho = LoadACOrtho()
conserved.tfs = subset(conservation.summary %>% arrange(desc(cor.score)), cor.score > cor.score.cutoff & is.tf)
head(conserved.tfs)

# Examples of highly conserved TFs
lapply(head(conserved.tfs$gene, 5), function(gene) geneDotPlotFast(ac.ortho, gene, group.by = "type"))
lapply(tail(conserved.tfs$gene, 5), function(gene) geneDotPlotFast(ac.ortho, gene, group.by = "type"))
```

## Binarization using k-means

Average expression

```{r, eval = FALSE}
# Within each species
features.use = rownames(ac.ortho)
vertebrateAC.list = SplitObject(ac.ortho, 'species')
avg.expr.list = lapply(vertebrateAC.list, function(object){
  print(object)
  object$seurat_clusters = factor(object$seurat_clusters)
  LogAvgExpr(object, group = 'seurat_clusters', features = features.use, assay = 'RNA')
})

# Reshape to be n_gene matrices of size n_species x n_OT
avg.mat.list = mclapply(features.use, function(feature) {
  do.call(gtools::smartbind, lapply(avg.expr.list, function(x) data.frame(as.list(x[feature,]))))
}, mc.cores = 16)
names(avg.mat.list) = features.use

saveRDS(avg.mat.list, "../../Ortho_Objects/AC_OT_average_expression.rds")
```

Percentage expressed

```{r, eval = FALSE}
# Within each species
features.use = rownames(ac.ortho)
vertebrateAC.list = SplitObject(ac.ortho, 'species')
percent.list = lapply(vertebrateAC.list, function(object){
  
  print(object)
  object$seurat_clusters = factor(object$seurat_clusters)
  
  # Percentage expressed
  pct.data = PercentageExpressed2(object, group.by = "seurat_clusters", features = features.use) # was PercentageExpressed
  
  return(pct.data)
})

saveRDS(percent.list, "../../Ortho_Objects/AC_OT_percent_expressed_list.rds")

# Reshape to be n_gene matrices of size n_species x n_OT
pct.list = mclapply(features.use, function(feature) {
  do.call(gtools::smartbind, lapply(percent.list, function(x) data.frame(as.list(x[feature,]))))/100 # divide by 100
}, mc.cores = 16)
names(pct.list) = features.use

saveRDS(pct.list, "../../Ortho_Objects/AC_OT_percent_expressed.rds")
```

New method using percentage expression and scaled, log-transformed expression

```{r}
# matrix.list = readRDS("../../Ortho_Objects/scaled_OT_expression_v4_nonlog.rds")

# Log-transformed and scaled
matrix.list = readRDS("../../Ortho_Objects/scaled_OT_expression_log.rds")

# Log transformed, not scaled
# matrix.list = readRDS("../../Ortho_Objects/AC_OT_average_expression.rds")

# rescaled.matrix.list = readRDS("../../Ortho_Objects/RGC_OT_rescaled_expr.rds")
pct.list = readRDS("../../Ortho_Objects/AC_OT_percent_expressed.rds")
# score.list = readRDS("../../Ortho_Objects/RGC_OT_expr_score.rds")
conservation.summary = readRDS("../../Ortho_Objects/vertebrateAC_conservation_v2.rds")
cor.score.cutoff = 0.3
```

```{r}
conserved.genes = subset(conservation.summary %>% arrange(desc(cor.score)), cor.score > cor.score.cutoff)
k.masks.res = lapply(conserved.genes$gene, function(gene){
  
  # Ensure that rows with all zeroes are set to NA to avoid scaling artifacts
  mat = apply(matrix.list[[gene]], 1, function(row){
    if(var(row, na.rm = TRUE) == 0) rep(NA, length(row)) else row
  }) %>% t %>% as.data.frame()
  
  bin.res = BinarizeExpressionMV2(gene = gene, 
                        avg.expr = unlist(mat) %>% as.data.frame %>% setNames(gene) %>% t, 
                        pct.expr = unlist(pct.list[[gene]]) %>% as.data.frame %>% setNames(gene) %>% t, 
                        fast = TRUE, 
                        compute.r.squared = FALSE, 
                        plot = TRUE, 
                        nstart = 50)
  
  message(paste0(gene, ':', bin.res$r.squared.k))
  
  df = as.data.frame(
    matrix(bin.res$k.vector, 
         nrow = nrow(matrix.list[[1]]), 
         ncol = ncol(matrix.list[[1]]))
    )
  rownames(df) = speciesList.liz
  colnames(df) = levels(ac.ortho$type_unordered)
  list(mat = df, r.squared = bin.res$r.squared.k)
}) %>% setNames(conserved.genes$gene)

k.masks = lapply(k.masks.res, function(x) x$mat)
r.squared.list = sapply(k.masks.res, function(x) x$r.squared)
saveRDS(r.squared.list, 'tf_analysis/r.squared.list.rds')
PrettyHistogram(r.squared.list, vline = 0.64)

pdf('tf_analysis/ac-gene-dotplots.pdf', height = 5, width = 9)
lapply(seq_along(k.masks), function(i) {
  gene = names(k.masks)[[i]]
  TitlePlot(geneDotPlotFast(ac.ortho, gene = gene, binarization = k.masks[[gene]]), 
            title = paste0(gene, ' (', round(r.squared.list[[i]], 2), ')'))
})
dev.off()

# Flagged TFs that should be removed
# remove.genes = c('DPF3', 'PAX6', 'FOXN3', 'PEG3', 'CUX1', 'RORA', 'KLF12')
remove.genes = c() # Filter later on, try toggling the threshold

# Thresholding
masks = lapply(k.masks, function(full_mask){
  new = apply(full_mask, 2, mean, na.rm = TRUE)
  new[new > 0.5] = 1
  new[new <= 0.5] = 0
  return(new)
})
names(masks) = names(k.masks)

# Remove genes with no variance
variance = sapply(masks, var) # CUX1 RORA 
which(variance == 0)
filtered.masks = masks[!(variance == 0 | names(masks) %in% remove.genes)]
length(filtered.masks)

# Save
tf.expr.mask = do.call(rbind, filtered.masks)
colnames(tf.expr.mask) = gsub('\\*', '', Metadata(ac.ortho, "seurat_clusters", "type")$type)
saveRDS(tf.expr.mask, "tf_analysis/k.mask.collapsed.rds")
saveRDS(k.masks, "tf_analysis/k.masks.rds")
```

# Checkpoint 1

```{r}
ac.ortho = readRDS("../../Ortho_Objects/vertebrateAC_BIGSEURAT.rds")
k.masks = readRDS('tf_analysis/k.masks.rds')
tf.expr.mask = readRDS("tf_analysis/k.mask.collapsed.rds")
r.squared.list = readRDS('tf_analysis/r.squared.list.rds')
tfs.use = intersect(tf_genes, names(which(r.squared.list > 0.65)))
```

# TF code

We need a nTF-dimensional binary vector for each OT in each OT. Final output: OT x TF matrix

```{r, fig.height=60, fig.width=8}
Heatmap((tf.expr.mask), 
        rect_gp = gpar(col = "black", lwd = 1),
        border_gp = gpar(col = "black", lwd = 1),
        col = c("lightgrey", "#584B9FFF"), 
        show_row_dend = TRUE, 
        show_column_dend = TRUE, 
        clustering_distance_rows = "manhattan", 
        clustering_method_rows = "average", 
        clustering_distance_columns = "manhattan", 
        clustering_method_columns = "average", 
        show_heatmap_legend = FALSE)
```

```{r, fig.width=30, fig.height=10}
Heatmap(t(tf.expr.mask), 
        rect_gp = gpar(col = "black", lwd = 1),
        border_gp = gpar(col = "black", lwd = 1),
        col = c("lightgrey", "#584B9FFF"), 
        show_row_dend = TRUE, 
        show_column_dend = TRUE, 
        clustering_distance_rows = "binary", 
        clustering_method_rows = "average", 
        clustering_distance_columns = "binary", 
        clustering_method_columns = "average", 
        show_heatmap_legend = FALSE)
```

# Hierarchical trees with pvclust

```{r, fig.height=6, fig.width=20}
library(pvclust)
result.man <- pvclust(tf.expr.mask, method.hclust = "average", method.dist = "manhattan", nboot = 1000)
plot(result.man)
mean(result.man$edges[,'au'])
```

```{r, fig.height=6, fig.width=20}
library(pvclust)
result.bin <- pvclust(tf.expr.mask, method.hclust = "average", method.dist = "binary", nboot = 1000)
plot(result.bin)
mean(result.bin$edges[,'au'])
```


```{r}
library(dendextend)
dnd1 <- as.dendrogram(result.man$hclust)
dnd2 <- as.dendrogram(result.bin$hclust)
## rearrange in ladderized fashion
# dnd1 <- ladder(dnd1)
# dnd2 <- ladder(dnd2)
## plot the tanglegram
dndlist <- dendextend::dendlist(dnd1, dnd2)
dndlist %>% entanglement
dndlist %>% untangle(method = "step1side") %>% entanglement
dendextend::tanglegram(dndlist %>% untangle(method = "step1side"), 
                       fast = TRUE, 
                       margin_inner = 5, 
                       main_left = "all species NJ", 
                       main_right = "all species MP")
```

Metrics: 

* Species consistency: similarity to species trees
* Self consistency: robustnesss to gene removal 
* Cross-method compatibility: compared to other method

Parameters:

* Method (4): NJ and MP seem more robust than UPGMA and ML
* Choice of genes (2): TFs seem more robust than non-TFs
* Gene cutoff

8 combinations

# Species consistency

```{r}
# Filtration
PrettyHistogram(r.squared.list, vline = c(0.55, 0.61, 0.64), fill = 'grey')
# features.use = names(which(r.squared.list > 0.61))
# features.use = names(which(r.squared.list > 0.64))

# Conserved genes with high R^2
features.use = names(which(r.squared.list > 0.7))

# Conserved TFs
features.use = intersect(tf_genes, names(which(r.squared.list > 0.65)))

# Make binarization within each species
species.mats = lapply(speciesList.liz, function(species){
  df = do.call(rbind, lapply(k.masks, function(mask) mask[species,]))
  colnames(df) = gsub('\\*', '', colnames(df))
  df
})

# Remove rows with all NAs (missing gene) and columns with all NAs (missing type)
species.mats.clean = lapply(species.mats, function(mat) {
  mat[!apply(mat, 1, function(x) all(is.na(x))),!apply(mat, 2, function(x) all(is.na(x)))]
}) 

# Check that NAs have been removed
stopifnot(all(!sapply(species.mats.clean, function(x) any(is.na(x)))))
```

```{r, fig.width=12, fig.height=10}
# Compute species trees
species.trees = lapply(species.mats.clean, function(mat){
  mat = mat[intersect(features.use, rownames(mat)),]
  ConstructTrees(t(mat), nIter = 100, mc.cores = 1, bootstrap = FALSE)
}) 

# With collapsed data
col.trees = ConstructTrees(t(tf.expr.mask[features.use,]), nIter = 100, mc.cores = 1, bootstrap = FALSE)

# Hierarchical clustering tree
h.tree = ape::as.phylo(as.hclust(readRDS('../../Ortho_Objects/ac.dendrogram.v2.rds')))
h.tree$tip.label = gsub('\\*', '', levels(ac.ortho$type_unordered))

# Robinson Foulds distances
# all.trees = c(unlist(species.trees, recursive = FALSE), col = col.trees, hc = list(h.tree))
all.trees = c(species.trees, all = list(col.trees))# hc = list(h.tree))

rf.mat = Iterate(c(
    upgma = lapply(all.trees, function(x) x$UPGMA),
    nj = lapply(all.trees, function(x) x$NJ),
    mp = lapply(all.trees, function(x) x$MP),
    ml = lapply(all.trees, function(x) x$ML),
    hc = list(h.tree)
  ), function(x, y) {
    tryCatch({
      tips.remove = union(setdiff(x$tip.label, y$tip.label), setdiff(y$tip.label, x$tip.label))
      print(tips.remove)
      x = drop.tip(x, tip = tips.remove)
      y = drop.tip(y, tip = tips.remove)
      RF.dist(x, y, normalize = TRUE)
  }, error = function(e) {
      # Print the error message to the console
      message("An error occurred: ", conditionMessage(e))
      browser()
  })
})

Heatmap2(rf.mat, 
         top_anno = HeatmapAnnotation(`Marginal\nRFD` = anno_barplot(min_max_norm(colMedians(rf.mat, na.rm = T)))), 
         legend.name = 'Robinson-\nFoulds\ndistance')
```


```{r, fig.height=10, fig.width=10}
saveRDS(all.trees, 'tf_analysis/all.trees.v2.rds')
```

# Compare trees

```{r, fig.height=20, fig.width=20}
library(phylogram)
library(ape)

# all.trees$col$NJ$node.label = all.trees$col$MP$node.label

plot(all.trees$all$UPGMA, type = 'unrooted', show.node.label = TRUE)
nodelabels(1:Nnode(all.trees$all$UPGMA) + 42)
plot(all.trees$all$NJ, type = 'unrooted', show.node.label = TRUE) # 60, 70
nodelabels(1:Nnode(all.trees$all$NJ) + 42)
plot(all.trees$all$MP, type = 'unrooted', show.node.label = TRUE) # 68, 47
# nodelabels(1:Nnode(all.trees$col$MP))
plot(all.trees$all$ML, type = 'unrooted', show.node.label = TRUE)
# nodelabels(1:Nnode(all.trees$col$ML))
```

```{r, fig.height=10, fig.width=10}
all.trees = readRDS('tf_analysis/all.trees.v2.rds')
CompareTrees(all.trees$all$NJ, all.trees$all$MP, 'NJ', 'MP')
CompareTrees(all.trees$all$UPGMA, all.trees$all$NJ, 'UPGMA', 'NJ')
CompareTrees(all.trees$all$MP, all.trees$all$ML, 'MP', 'ML')
```

# TFs versus non-TFs

```{r, fig.height=2, fig.width=4}
# Maximum parsimony tree with 70 conserved TFs
tfs.use = intersect(tf_genes, names(which(r.squared.list > 0.65)))
tf.trees = ConstructTrees(t(tf.expr.mask[tfs.use,]), nIter = 100, mc.cores = 10, bootstrap = TRUE)
tf.bs = tf.trees$MP$bs.support
saveRDS(tf.trees, 'tf_analysis/tf.trees.rds')

# Non-tfs
nontf.trees.list = lapply(seq_len(100), function(i){
  nontfs.use = sample(setdiff(names(which(r.squared.list > 0.65)), tf_genes), size = length(tfs.use))
  ConstructTrees(t(tf.expr.mask[nontfs.use,]), nIter = 100, mc.cores = 10, bootstrap = TRUE)
})
saveRDS(nontf.trees.list, 'tf_analysis/nontf.trees.list.rds')

# Get bs values
bs.values = unlist(sapply(nontf.trees.list, function(x) x$MP$bs.support))
mean.bs = unlist(sapply(nontf.trees.list, function(x) mean(x$MP$bs.support)))
PrettyHistogram(mean.bs, vline = mean(tf.bs), xlab = 'Mean bootstrap support')

t.test(tf.bs, bs.values)
wilcox.test(tf.bs, nontf.bs)
```

```{r, fig.height=10, fig.width=10}
plot(tf.trees$MP, type = 'unrooted')
plot(tf.trees$MP)
```

TFs lead to more robust trees than comparable sets of non-TFs. 

```{r}
boxplot(tf.trees$UPGMA$bs.support, tf.trees$NJ$bs.support, tf.trees$MP$bs.support)
boxplot(unlist(lapply(nontf.trees.list, function(x) x$UPGMA$bs.support)), 
        unlist(lapply(nontf.trees.list, function(x) x$NJ$bs.support)), 
        unlist(lapply(nontf.trees.list, function(x) x$MP$bs.support)))
```

Robinson foulds distance to all the bootstrap trees

```{r}
bs.tf.trees = BootstrapTrees(t(tf.expr.mask[tfs.use,]), mc.cores = 10)
nontfs.use = sample(setdiff(names(which(r.squared.list > 0.65)), tf_genes), size = length(tfs.use))
bs.nontf.trees = BootstrapTrees(t(tf.expr.mask[nontfs.use,]), mc.cores = 10)
```

```{r, fig.height=5, fig.width=3}
Heatmap2(data.frame(UPGMA = sapply(bs.tf.trees$UPGMA, function(this) RF.dist(tf.trees$UPGMA, this, normalize = TRUE)), 
                NJ = sapply(bs.tf.trees$NJ, function(this) RF.dist(tf.trees$NJ, this, normalize = TRUE)), 
                MP = sapply(bs.tf.trees$MP, function(this) RF.dist(tf.trees$MP, this, normalize = TRUE))), 
         col = colorRamp2(c(0.4,1), c('white', 'red')))

Heatmap2(data.frame(UPGMA = sapply(bs.nontf.trees$UPGMA, function(this) RF.dist(tf.trees$UPGMA, this, normalize = TRUE)), 
                NJ = sapply(bs.nontf.trees$NJ, function(this) RF.dist(tf.trees$NJ, this, normalize = TRUE)), 
                MP = sapply(bs.nontf.trees$MP, function(this) RF.dist(tf.trees$MP, this, normalize = TRUE))), 
         col = colorRamp2(c(0.4,1), c('white', 'red')))


```

# Cross method compatibility

```{r}
# Randomly sample some TFs, calculate trees, and check similarity
resampled.tf.trees = mclapply(seq_len(params$nIter), function(i){
  genes.use = sample(tfs.use, size = length(tfs.use)-5)
  ConstructTrees(t(tf.expr.mask[genes.use,]), bootstrap = FALSE, minit = 100)
}, mc.cores = 1)

resampled.tf.trees
```
```{r}
rf.mat = Iterate(c(
    upgma = lapply(resampled.tf.trees, function(x) x$UPGMA),
    nj = lapply(resampled.tf.trees, function(x) x$NJ),
    mp = lapply(resampled.tf.trees, function(x) x$MP),
    ml = lapply(resampled.tf.trees, function(x) x$ML)
  ), function(x, y) {
      tryCatch({
      tips.remove = union(setdiff(x$tip.label, y$tip.label), setdiff(y$tip.label, x$tip.label))
      print(tips.remove)
      x = drop.tip(x, tip = tips.remove)
      y = drop.tip(y, tip = tips.remove)
      RF.dist(x, y, normalize = TRUE)
  }, error = function(e) {
      # Print the error message to the console
      message("An error occurred: ", conditionMessage(e))
      browser()
  })
})

colnames(rf.mat) = sub("(\\d)", "\\.\\1", colnames(rf.mat))
rownames(rf.mat) = sub("(\\d)", "\\.\\1", rownames(rf.mat))
rf.mat = MakeSmartMatrix(rf.mat, delimiter.row = '\\.', delimiter.col = '\\.')

Heatmap2(rf.mat@matrix, 
         top_anno = HeatmapAnnotation(`Marginal\nRFD` = anno_barplot(min_max_norm(matrixStats::colMedians(rf.mat@matrix, na.rm = T)))), 
         legend.name = 'Robinson-\nFoulds\ndistance')
```

mp2 is more robust than mp1

```{r}
CompareTrees(resampled.tf.trees[[1]]$MP, resampled.tf.trees[[2]]$MP)
CompareTrees(tf.trees$MP, resampled.tf.trees[[2]]$MP)
CompareTrees(resampled.tf.trees[[1]]$MP, tf.trees$MP)
```

# Monophyletic genes

```{r}
mono.res = lapply(tf.trees[METHODS], function(tree) {
  names(which((apply(tf.expr.mask, 1, function(row){
    is.monophyletic(tree, names(which(row > 0)))
  }))))
})
mono.res

library(UpSetR)
upset(fromList(mono.res),
                    # A few optional arguments to customize the plot:
                    nsets = 5,                 # The number of sets to display
                    nintersects = 20,          # The number of intersections to display
                    mainbar.y.label = "Intersection Size",   # Label for the main bar chart
                    sets.x.label = "Set Size", # Label for the set size bar chart
                    text.scale = c(1.5, 1.3, 1, 1, 2, 1.2), # Adjust font sizes
                    point.size = 2.5,
                    line.size = 1)
```

Visualize monophyletic

```{r, fig.height=8, fig.width=10}
Heatmap(t(tf.expr.mask[mono.res$MP,]), 
        rect_gp = gpar(col = "black", lwd = 1),
        border_gp = gpar(col = "black", lwd = 1),
        col = c("lightgrey", "#584B9FFF"), 
        show_row_dend = TRUE, 
        show_column_dend = TRUE, 
        clustering_distance_rows = "binary", 
        clustering_method_rows = "average", 
        clustering_distance_columns = "binary", 
        clustering_method_columns = "average", 
        show_heatmap_legend = FALSE)
```

Most are trivially monophyletic

# Visualize tree with gene expression

```{r, fig.width=15, fig.height=8}
lapply(tf.trees[METHODS], function(tree){
  Heatmap(t(tf.expr.mask[,tree$tip.label]), 
        # rect_gp = gpar(col = "black", lwd = 1),
        border_gp = gpar(col = "black", lwd = 1),
        col = c("lightgrey", "#584B9FFF"), 
        show_row_dend = TRUE, 
        show_column_dend = TRUE, 
        cluster_rows = as.dendrogram.phylo(midpoint(tree)), 
        clustering_distance_columns = "binary", 
        clustering_method_columns = "average", 
        show_heatmap_legend = FALSE)
})
```


```{r}
plot(tf.trees$ML)
```

Plot TFs and adhesion molecules

```{r, fig.height=6, fig.width=10}
lapply(tf.trees[METHODS], function(tree){
  Heatmap(t(tf.expr.mask[tfs.use,tree$tip.label]), 
        # rect_gp = gpar(col = "black", lwd = 1),
        border_gp = gpar(col = "black", lwd = 1),
        col = c("lightgrey", "#584B9FFF"), 
        show_row_dend = TRUE, 
        show_column_dend = TRUE, 
        cluster_rows = as.dendrogram.phylo(midpoint(tree)), 
        clustering_distance_columns = "binary", 
        clustering_method_columns = "average", 
        show_heatmap_legend = FALSE)
})
```



```{r, fig.height=6, fig.width=6}
gene.lists = readRDS('../../reference_files/gene-lists-ac.rds')

adhesion.use = intersect(gene.lists$adhesion, names(which(r.squared.list > 0.60)))
lapply(tf.trees[METHODS], function(tree){
  Heatmap(t(tf.expr.mask[adhesion.use,tree$tip.label]), 
        # rect_gp = gpar(col = "black", lwd = 1),
        border_gp = gpar(col = "black", lwd = 1),
        col = c("lightgrey", "#584B9FFF"), 
        show_row_dend = TRUE, 
        show_column_dend = TRUE, 
        cluster_rows = as.dendrogram.phylo(midpoint(tree)), 
        clustering_distance_columns = "binary", 
        clustering_method_columns = "average", 
        show_heatmap_legend = FALSE)
})
```

# More plots

```{r, fig.height=12, fig.width=8}
ac.ortho$type_clean = factor(gsub('\\*', '', ac.ortho$type_unordered), 
                             levels = gsub('\\*', '', levels(ac.ortho$type_unordered)))

genes.omit = c('HLF', 'RORB', 'ZFPM2', 'ZNF503', 'FOXP1', 'DACH1', 'CUX2', 'KLF7', 'ZNF488', 'NR2F2', 'NR2F1', 
               'PBX1', 'ZNF521', 'POU3F3', 'ZBTB18', 'FEZF1', 'FOXN2', 
               'GBX2', 'FEZF2', 'ZNF536', 'BNC1', 'NR1H4', 'ONECUT1', 'TFAP2E', 'EOMES', 'SOX5', 'ESRRG', 'BCL11B')

single.type.genes = c('NR1H4', 'ONECUT1', 'EOMES', 'SOX2', 'ISL1', 'GBX2')

family.genes = setdiff(setdiff(c(tfs.use, 'TFAP2A'), genes.omit), single.type.genes)
clust = hclust(dist(tf.expr.mask[family.genes,], method = 'binary'))
family.genes.ordered = clust$labels[clust$order]

# Family-defining TFs
ClusteredDotplot2(ac.ortho, 
                  features = family.genes.ordered, 
                  group.by = 'type_clean', 
                  cluster_columns = midpoint(tf.trees$MP),
                  max.pct = 40, 
                  dot.scale.factor = 0.5, 
                  pseudocount = 5, 
                  rotate = FALSE, 
                  show_row_dend = FALSE)

# Type-defining TFs
```


```{r, fig.height=6, fig.width=8.5}
# Rotate
dend = as.dendrogram.phylo(midpoint(tf.trees$MP))
tf.trees$MP2 = dendextend::rotate(dend, order = c("26_A8", "24_SEG", "2_NNgly", "34", "17_ROBO3", "15", "18_TRHDE", "27_VG3", "41_TRHDE", "22_A2", "21_PDGFRA", "28", "1_MAF", "3", "9", "20", "6_NPY", "13", "40", "8_SAC", "42", "38", "14_CA2", "16_NNgaba", "4_nNOS", "31_nNOS", "10_MAF", "23_RXRG", "39", "35", "7", "33_CA1", "11", "25", "12_nNOS", "19_CRH", "30_A17", "32_A17", "29_SLC35D3", "5", "36_VIP", "37"))

ClusteredDotplot2(ac.ortho, 
                  features = family.genes.ordered, 
                  group.by = 'type_clean', 
                  cluster_rows = tf.trees$MP2,
                  max.pct = 40, 
                  dot.scale.factor = 0.3, 
                  pseudocount = 5, 
                  rotate = TRUE)
```


```{r, fig.height=6, fig.width=5}
adhesion.use = intersect(gene.lists$adhesion, names(which(r.squared.list > 0.60)))
ClusteredDotplot2(ac.ortho, 
                  features = adhesion.use, 
                  group.by = 'type_clean', 
                  cluster_rows = tf.trees$MP2,
                  cluster_columns = TRUE,
                  max.pct = 40, 
                  dot.scale.factor = 0.3, 
                  pseudocount = 5, 
                  rotate = TRUE)

adhesion.genes = c('UNC5D', 'CLMP', 'CDH7', 'DSCAM', 'CDH10', 'CDH8', 'CDH4', 'PCDH10', 'IGSF11', 'DSCAML1', 'CDH22', 'TENM3')
ClusteredDotplot2(ac.ortho, 
                  features = adhesion.genes, 
                  group.by = 'type_clean', 
                  cluster_rows = tf.trees$MP2,
                  cluster_columns = TRUE,
                  max.pct = 40, 
                  dot.scale.factor = 0.3, 
                  pseudocount = 5, 
                  rotate = TRUE)
```

# Explore the tree space

```{r, fig.width=10, fig.height=10}
set.seed(42)
phydat = BinaryPhyDat(t(tf.expr.mask[tfs.use,]))
treeMPs <- pratchet(phydat, trace = 0, minit = 1000, all = TRUE)

# Add branch lengths (will be needed for consensus branch lengths)
treeMPs = lapply(treeMPs, function(tree){
  tree = acctran(tree, phydat)
  di2multi(tree, tol = 0)
})
plot(treeMPs[[1]])

# Consensus tree
ctree = consensus(treeMPs, p = 0.5, check.labels = TRUE)
plot(ctree)
saveRDS(ctree, 'tf_analysis/ctree.rds')
ctree_bl <- phytools::consensus.edges(treeMPs, method = "mean.edge", consensus.tree = ctree)
plot(ctree_bl)
saveRDS(ctree_bl, 'tf_analysis/ctree_bl.rds')

# Show that the consensus tree is particularly robust

# 132 equivalent trees
# nni.matrix = treeMPs

ctree2  <- acctran(ctree, phydat)
ctree2  <- di2multi(ctree2, tol = 0)
plot(ctree2)
saveRDS(ctree2, 'tf_analysis/ctree2.rds')
```


```{r, fig.width=10, fig.height=10}
CompareTrees(ctree2, tf.trees$NJ)
```

What is robust? 
* 22, 41, 27, 18, 15, 17, 34, 2, 26, 24
* 33,7
* 8, 42
* 21, 28, 36, 5, 37
* 1,3,9,20,23,39
* 12, 11, 25, 19, 32, 30
* 40, 13, 6
* 16, 14, 4, 31

Unclear: 
29, 35, 38, 10

```{r, fig.height=10, fig.width=10}
plotBS(ctree, as.multiPhylo(treeMPs)) # all > 0.5
```

# How many genes to include (when does robustness saturate?)

```{r}
loo.trees.list = lapply(c(5,10,20,40,70), function(nTFs) LOOTrees(t(tf.expr.mask[tfs.use[1:nTFs],]), consensus = TRUE))
loo.trees = lapply(c(40), function(nTFs) LOOTrees(t(tf.expr.mask[tfs.use[1:nTFs],]), consensus = TRUE))
loo.trees.list[[4]] = loo.trees[[1]]
RFMatrix(loo.trees.list[[1]])
RFMatrix(loo.trees.list[[2]])
RFMatrix(loo.trees.list[[3]])
RFMatrix(loo.trees.list[[4]])
saveRDS(loo.trees.list, 'tf_analysis/loo.trees.list.rds')
```

# Robustness diagram (using LOO trees)

```{r}
# loo.trees = LOOTrees(t(tf.expr.mask[tfs.use[1:5],]), consensus = TRUE)

# Check if clade is monophyletic

clades <- list(
  c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
  c(33, 7), # CA1
  c(8, 42), # SAC
  c(21, 28), # PDGFRA
  c(36, 5, 37), # VIP
  c(1, 3, 9, 20), # MAF
  c(23, 39),
  c(12, 11, 25, 19, 32, 30), # CRH
  c(40, 13, 6), # NPY
  c(16, 14, 4, 31), # CA2
  c(29, 35, 38, 10) # not robust
)
clades = lapply(clades, as.character)
names(clades) = 1:length(clades)

# Binary presence doesn't capture some cases
robust.res = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
  sapply(clades, function(clade){
    tree = this$UPGMA
    tree$tip.label = ExtractString(tree$tip.label, after = '_')
    is.monophyletic(tree, tips = as.character(clade))
  })
})), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      is.monophyletic(tree, tips = as.character(clade))
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades, function(clade){
        tree = this$MP
        tree$tip.label = ExtractString(tree$tip.label, after = '_')
        is.monophyletic(tree, tips = as.character(clade))
      })
    })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
        sapply(clades, function(clade){
          tree = this$ML
          tree$tip.label = ExtractString(tree$tip.label, after = '_')
          is.monophyletic(tree, tips = as.character(clade))
        })
    }))
)

Heatmap2(t(apply(robust.res, 1, as.numeric)))
saveRDS(robust.res, 'tf_analysis/robust.res.binary.rds')
```

Clades broad

```{r}
# loo.trees = LOOTrees(t(tf.expr.mask[tfs.use[1:5],]), consensus = TRUE)

# Check if clade is monophyletic

clades.broad <- list(
  c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
  c(21, 28), # PDGFRA
  c(36, 5, 37), # VIP
  c(1, 3, 9, 20, 23, 39, 33, 7, 8, 42, 40, 13, 6, 16, 14, 4, 31, 10, 35, 38), # gaba
  c(12, 11, 25, 19, 32, 30), # CRH
  c(29) # not robust
)
clades.broad = lapply(clades.broad, as.character)
names(clades.broad) = 1:length(clades.broad)

# Binary presence doesn't capture some cases
robust.res = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
  sapply(clades.broad, function(clade){
    tree = this$UPGMA
    tree$tip.label = ExtractString(tree$tip.label, after = '_')
    is.monophyletic(tree, tips = as.character(clade))
  })
})), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades.broad, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      is.monophyletic(tree, tips = as.character(clade))
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades.broad, function(clade){
        tree = this$MP
        tree$tip.label = ExtractString(tree$tip.label, after = '_')
        is.monophyletic(tree, tips = as.character(clade))
      })
    })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
        sapply(clades.broad, function(clade){
          tree = this$ML
          tree$tip.label = ExtractString(tree$tip.label, after = '_')
          is.monophyletic(tree, tips = as.character(clade))
        })
    }))
)

Heatmap2(t(apply(robust.res, 1, as.numeric)))
saveRDS(robust.res, 'tf_analysis/robust.res.binary.broad.rds')
```

Nearest neighbors

```{r}

# Binary presence doesn't capture some cases
robust.res.nn = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$UPGMA
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      are.neighbors(cophenetic(tree), nodes = as.character(clade), binary = FALSE)
    })
})), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      are.neighbors(cophenetic(tree), nodes = as.character(clade), binary = FALSE)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades, function(clade){
        tree = this$MP
        tree$tip.label = ExtractString(tree$tip.label, after = '_')
        are.neighbors(cophenetic(tree), nodes = as.character(clade), binary = FALSE)
      })
    })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades, function(clade){
        tree = this$ML
        tree$tip.label = ExtractString(tree$tip.label, after = '_')
        are.neighbors(cophenetic(tree), nodes = as.character(clade), binary = FALSE)
      })
    }))
)

Heatmap2(t(apply(robust.res.nn, 1, as.numeric)))
saveRDS(robust.res, 'tf_analysis/robust.res.nn.rds')
```

Normalized Average Within-Cluster Distance

```{r}
robust.res.dist = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$UPGMA
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_normalized_distance(cophenetic(tree), clade)/sum(tree$edge.length)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_normalized_distance(cophenetic(tree), clade)/sum(tree$edge.length)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$MP
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_normalized_distance(cophenetic(tree), clade)/sum(tree$edge.length)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$ML
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_normalized_distance(cophenetic(tree), clade)/sum(tree$edge.length)
    })
  }))
)

Heatmap2(t(apply(robust.res.dist, 1, as.numeric)))#, col = colorRamp2(c(-1, 0, 1), c('blue', 'white', 'red')))
```

Sihloutte score

```{r}
robust.res.dist = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$UPGMA
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_silhouette_coefficient(cophenetic(tree), clade, clades)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_silhouette_coefficient(cophenetic(tree), (clade), clades)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$MP
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_silhouette_coefficient(cophenetic(tree), (clade), clades)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$ML
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      calculate_silhouette_coefficient(cophenetic(tree), (clade), clades)
    })
  }))
)

Heatmap2(t(apply(robust.res.dist, 1, as.numeric)), col = colorRamp2(c(-1, 0, 1), c('blue', 'white', 'red')))
```

Proceeding with presence 

```{r, fig.height=3, fig.width=4}
plot(ctree)
nodelabels(1:Nnode(ctree) + 42)
ctree3 = root(ctree, node = 59, edgelabel = TRUE)
ctree3$node.label = as.numeric(ctree3$node.label)
ctree2$node.label = as.numeric(ctree2$node.label)
plotBS(ctree)
plotBS(ctree2)
plotBS(ctree3)

dend = as.dendrogram(ctree3)
dend = dendextend::rotate(dend, order = c("26_A8", "24_SEG", "2_NNgly", "34", "17_ROBO3", "15", "18_TRHDE", "41_TRHDE", "27_VG3", "22_A2", "21_PDGFRA", "28", "5", "36_VIP", "37", "1_MAF", "3", "9", "20", "6_NPY", "13", "40", "8_SAC", "42", "38", "14_CA2", "16_NNgaba", "4_nNOS", "31_nNOS", "10_MAF", "23_RXRG", "39", "35", "7", "33_CA1", "11", "25", "12_nNOS", "19_CRH", "30_A17", "32_A17", "29_SLC35D3"))
plot(dend)
saveRDS(dend, 'tf_analysis/dend.rds')
type.order = labels(dend)
type.order.no = ExtractString(type.order, after = '_')


# Make a type by presence matrix
pres.mat = matrix(rep(type.order.no, nrow(robust.res)), ncol = length(type.order.no), nrow = nrow(robust.res), byrow = T)
pres.mat = t(apply(pres.mat, 1, function(x) as.character(unlist(convert_values(x, invert_list(clades))))))
head(pres.mat)

# Fill with results
pres.mat.filled = pres.mat
for(i in seq_len(nrow(robust.res))){
  for(clade in names(clades)){
    row = pres.mat[i,]
    if(!robust.res[i,as.numeric(clade)]) pres.mat.filled[i,row == clade] = NA
  }
}

Heatmap(pres.mat.filled[,order(as.numeric(type.order.no))], cluster_columns = dend, 
        col = AcPalette3(11), 
        # rect_gp = gpar(col = "grey", lwd = 0.1),
        border_gp = gpar(col = "black", lty = 1), 
        show_heatmap_legend = FALSE
        )
```

# Gene panels

```{r, fig.height=7, fig.width=12}
ac.ortho$type_clean = factor(gsub('\\*', '', ac.ortho$type_unordered), 
                             levels = gsub('\\*', '', levels(ac.ortho$type_unordered)))

genes.omit = c()
# c('HLF', 'RORB', 'ZFPM2', 'ZNF503', 'FOXP1', 'DACH1', 'CUX2', 'KLF7', 'ZNF488', 'NR2F2', 'NR2F1', 
               # 'PBX1', 'ZNF521', 'POU3F3', 'ZBTB18', 'FEZF1', 'FOXN2', 
               # 'GBX2', 'FEZF2', 'ZNF536', 'BNC1', 'NR1H4', 'ONECUT1', 'TFAP2E', 'EOMES', 'SOX5', 'ESRRG', 'BCL11B')

single.type.genes = c('NR1H4', 'ONECUT1', 'EOMES', 'SOX2', 'ISL1', 'GBX2')

family.genes = setdiff(setdiff(c(tfs.use, 'TFAP2A'), genes.omit), single.type.genes)
clust = hclust(dist(tf.expr.mask[family.genes,], method = 'binary'))
family.genes.ordered = clust$labels[clust$order]

# Family-defining TFs
ClusteredDotplot2(ac.ortho, 
                  features = family.genes.ordered, 
                  group.by = 'type_clean', 
                  cluster_rows = dend,
                  max.pct = 40, 
                  dot.scale.factor = 0.5, 
                  pseudocount = 5, 
                  rotate = TRUE, 
                  show_row_dend = FALSE)

# Type-defining TFs
```


```{r, fig.height=6, fig.width=8.5}
# Rotate
dend = as.dendrogram.phylo(midpoint(tf.trees$MP))
tf.trees$MP2 = dendextend::rotate(dend, order = c("26_A8", "24_SEG", "2_NNgly", "34", "17_ROBO3", "15", "18_TRHDE", "27_VG3", "41_TRHDE", "22_A2", "21_PDGFRA", "28", "1_MAF", "3", "9", "20", "6_NPY", "13", "40", "8_SAC", "42", "38", "14_CA2", "16_NNgaba", "4_nNOS", "31_nNOS", "10_MAF", "23_RXRG", "39", "35", "7", "33_CA1", "11", "25", "12_nNOS", "19_CRH", "30_A17", "32_A17", "29_SLC35D3", "5", "36_VIP", "37"))

ClusteredDotplot2(ac.ortho, 
                  features = family.genes.ordered, 
                  group.by = 'type_clean', 
                  cluster_rows = tf.trees$MP2,
                  max.pct = 40, 
                  dot.scale.factor = 0.3, 
                  pseudocount = 5, 
                  rotate = TRUE)
```


```{r, fig.height=6, fig.width=5}
adhesion.use = intersect(gene.lists$adhesion, names(which(r.squared.list > 0.60)))
ClusteredDotplot2(ac.ortho, 
                  features = adhesion.use, 
                  group.by = 'type_clean', 
                  cluster_rows = tf.trees$MP2,
                  cluster_columns = TRUE,
                  max.pct = 40, 
                  dot.scale.factor = 0.3, 
                  pseudocount = 5, 
                  rotate = TRUE)

adhesion.genes = c('UNC5D', 'CLMP', 'CDH7', 'DSCAM', 'CDH10', 'CDH8', 'CDH4', 'PCDH10', 'IGSF11', 'DSCAML1', 'CDH22', 'TENM3')
ClusteredDotplot2(ac.ortho, 
                  features = adhesion.genes, 
                  group.by = 'type_clean', 
                  cluster_rows = tf.trees$MP2,
                  cluster_columns = TRUE,
                  max.pct = 40, 
                  dot.scale.factor = 0.3, 
                  pseudocount = 5, 
                  rotate = TRUE)
```

# Checkpoint 2

```{r}
conservation.summary = readRDS("../../Ortho_Objects/vertebrateAC_conservation_v2.rds")
dend = readRDS('tf_analysis/dend.rds')
ctree = readRDS('tf_analysis/ctree.rds')
ctree_bl = readRDS('tf_analysis/ctree_bl.rds')
ctree_bl$tip.label = ctree_bl$tip.label2
ctree2 = readRDS('tf_analysis/ctree2.rds')
# ctree2$edge.length = ctree2$edge.length/11 # similar to UPGMA and NJ tree sizes
# Set 0 branch lengths to a small value
ctree2$edge.length[ctree2$edge.length == 0] = 1e-5
tf.expr.mask = readRDS("tf_analysis/k.mask.collapsed.rds")
loo.trees.list = readRDS('tf_analysis/loo.trees.list.rds')
gene.lists = readRDS('../../reference_files/gene-lists-ac.rds')
r.squared.list = readRDS('tf_analysis/r.squared.list.rds')
tfs.use = intersect(tf_genes, names(which(r.squared.list > 0.65)))

ac.ortho$type_clean = factor(gsub('\\*', '', ac.ortho$type_unordered), 
                             levels = gsub('\\*', '', levels(ac.ortho$type_unordered)))
```

# Phylogenetic signal

```{r}
library(phylolm)
library(ape)

# PhylogeneticSignalTest(tree, min_max_norm(trait_data$trait_B) %>% setNames(tree$tip.label))
test = rtree(42)
test$tip.label = colnames(tf.expr.mask)


# Rank genes by PRT p-value
phylo.res = do.call(rbind, apply(tf.expr.mask, 1, function(gene){
  # names(gene) = ExtractString(gene, after = '_')
  tryCatch({
    PhylogeneticSignalTest(ctree2, gene)}, 
    error = function(e) {
      # Print the error message to the console
      message("An error occurred: ", conditionMessage(e))
      NULL
  })
}))

phylo.res = phylo.res %>% arrange(pval)
```

```{r, fig.height=50, fig.width=10}
ClusteredDotplot2(ac.ortho, 
                  features = rownames(phylo.res), 
                  group.by = 'type_clean', 
                  cluster_columns = dend,
                  max.pct = 40, 
                  dot.scale.factor = 0.5, 
                  pseudocount = 5, 
                  rotate = FALSE)
```

Just TFs

```{r, fig.height=20, fig.width=10}
ClusteredDotplot2(ac.ortho, 
                  features = intersect(rownames(phylo.res), gene.lists$tf_genes), 
                  group.by = 'type_clean', 
                  cluster_columns = dend,
                  max.pct = 40, 
                  dot.scale.factor = 0.5, 
                  pseudocount = 5, 
                  rotate = FALSE)


glycinergic_families <- list(
  Pan = c("TCF4"),
  NFI_family = rev(c("NR2F2","NFIX", "NFIB", "NEUROD4", "SATB2", "EBF2",
                 "EBF3", "ST18", "PROX1", "EBF1", "PRDM13")),
  Neurod2_family = c("NEUROD2", "TFAP2B", 'TFAP2A'),
  A2 = c("ONECUT1", "NFIA"),
  VG3_family = c( "FOXP2", "BNC2", "TFAP2E", "NR1H4", "LHX2", "BNC1")
)

clades <- list(
  c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
  c(21, 28), # PDGFRA SOX5
  c(36, 5, 37), # VIP SOX5 NR4A2
  
  c(40, 13, 6), # NPY ZFHX3 ZFHX4 MEIS2 ZBTB16 ESRRG
  c(8, 42), # SAC ZFHX3 ZFHX4 MEIS2 TBX2 SIX3 ZBTB16 TFAP2D
  c(1, 3, 9, 20), # MAF ZFHX3 ZFHX4 MEIS2 SIX3 ZBTB16 BHLHE22 PRDM8
  c(23, 39), # ZFHX3 ZFHX4 MEIS2 SIX3 ZBTB16
  
  c(16, 14, 4, 31), # CA2 ZFHX3 ZFHX4 MEIS2 LHX2
  c(33, 7), # CA1 ZFHX3 ZFHX4 MEIS2 TSHZ1
  c(12, 11, 25, 19, 32, 30), # CRH MEIS2 ZNF536 DACH2 LHX9
  c(29, 35, 38, 10) # not robust
)

```


```{r}
# adhesion.show = sort(unique(c('UNC5D', 'CLMP', 'CDH7', 'DSCAM', 'CDH10', 'CDH8', 'CDH4', 'PCDH10', 'IGSF11', 
#                               'DSCAML1', 'CDH22', 'TENM3', 'IGSF11', 'DSCAM', 'DSCAML1', 'CDH10', 'PCHD19', 
#                               'CDH8', 'PCDH10', 'PCDH17', 'CDH22')))
ClusteredDotplot2(ac.ortho, 
                  features = intersect(rownames(phylo.res), gene.lists$adhesion), 
                  group.by = 'type_clean', 
                  cluster_columns = dend,
                  max.pct = 40, 
                  dot.scale.factor = 0.5, 
                  pseudocount = 5, 
                  rotate = FALSE)
```

# Reconstruction of developmental sequence


```{r, fig.height=12, fig.width=20}
# Run the checkpoint 1 before this
library(phangorn)
ctree2 = readRDS('tf_analysis/ctree2.rds')
treeRooted = ctree2
for(this.node in (c(67,69,68,70,72))){
  treeRooted = ape::rotate(treeRooted, node = this.node)
}
# treeRooted = ctree2
# treeRooted = makeNodeLabel(treeRooted)
# treeRooted = root(treeRooted, node = 66)
# treeRooted = ape::rotate(treeRooted, node = 43)
plot(treeRooted, direction = 'downward')
nodelabels(1:treeRooted$Nnode + length(treeRooted$tip.label))

# treeRooted = ape::rotate(treeRooted, node = 70)
# plot(treeRooted, direction = 'downward')
# nodelabels(1:treeRooted$Nnode + length(treeRooted$tip.label))

phydat = BinaryPhyDat(t(tf.expr.mask[tfs.use,]))
anc.acctran = anc_pars(treeRooted, phydat, type = 'ACCTRAN')
tf.order = tfs.use
# anc.mpr <- ancestral.pars(treeRooted, phydat, "MPR")

pdf("tf_analysis/reconstruction_v4_acctran.pdf", height = 6.5, width = 6.5)
for(gene in tfs.use){
  # plotAnc(treeLabeled, anc.mpr, which(tf.order == gene), type = "phylogram", cex.pie = 0.4, col = c("white", "black"))
  plotAnc(anc.acctran, which(tfs.use == gene), type = "phylogram", cex.pie = 0.3, col = c("white", "black"), cex = 0.6, show.node.label = FALSE)
  title(gene)
}
dev.off()
```

Find TF switches

```{r, fig.height=30, fig.width=20}
anc.acctran <- ancestral.pars(treeRooted, phydat, "ACCTRAN")
gained = apply(treeRooted$edge, 1, function(row){
  gained = tf.order[which(ceiling(anc.acctran[[row[2]]][,2]) - ceiling(anc.acctran[[row[1]]][,2]) > 0)] #& anc.acctran[[row[1]]][,2] == 0)] # an increase from 0 (not expressed)
  if(length(gained) > 0){
    gained = paste0(gained, "+")
    paste0(gained, collapse = "\n")
  } else {
    NA
  }
})

lost = apply(treeRooted$edge, 1, function(row){
  lost = tf.order[which(ceiling(anc.acctran[[row[2]]][,2]) - ceiling(anc.acctran[[row[1]]][,2]) < 0)] #& anc.acctran[[row[1]]][,2] == 1)] # a decrease from 1 (expressed)
  if(length(lost) > 0){
    lost = paste0(lost, "-")
    paste0(lost, collapse = "\n")
  } else {
    NA
  }
})

# Plot gained TFs
pdf("tf_analysis/labeled_tree_gained_lost_v4.pdf", height=30, width=20)
plot(treeRooted, show.node.label = TRUE)
edgelabels(gained, srt = 0, bg = "green")

# Plot lost TFs
plot(treeRooted, show.node.label = TRUE)
edgelabels(lost, srt = 0, bg = "red")
dev.off()

# Plot all TF switches
switches = gsub('^\n|\n$', '', gsub("NA", "", paste0(gained, "\n", lost)))
switches[switches == "\n"] = NA

pdf("tf_analysis/labeled_tree_v4.pdf", height = 20, width = 30)
plot(treeRooted, show.node.label = FALSE, direction = "downwards")
edgelabels(switches, srt = 0, bg = "white")
add.scale.bar()
dev.off()

pdf("../../figures/my_figs/figS-labeled_tree_v6.pdf", height = 30, width = 22, family = "ArialMT")
treeRooted3 = treeRooted
treeRooted3$tip.label = gsub(' ', '\n', orthotype_labels(treeRooted$tip.label, asterisk = FALSE, as.factor = FALSE))
plot(treeRooted3, show.node.label = FALSE, 
     edge.width = 2,
     # direction = "downwards", 
     show.tip.label = FALSE)
tiplabels(
  text = treeRooted3$tip.label,
  tip = 1:length(treeRooted3$tip.label),
  srt = 0,              # rotate 90 degrees
  adj = 0,               # vertical alignment
  frame = "none",
  col = "black"
)
edgelabels(switches, srt = 0, bg = "white", cex = 1, font = 3)
add.scale.bar(lwd = 2)
dev.off()
```

Save genes for later

```{r}
tf.switches.clean = lapply(str_split(switches, '\n'), function(x) gsub('\\-|\\+', '', x))
saveRDS(tf.switches.clean, 'tf_analysis/tf.switches.clean.rds')
```

Plot tree skeleton

```{r}
pdf('tf_analysis/tree_skeleton.pdf')
treeRooted2 = root(treeRooted, node = 66, edgelabel = TRUE)
plot(treeRooted2, show.node.label = FALSE, direction = "downwards")
add.scale.bar()
dev.off()
```


```{r}
#' Could also do reconstruction on multifurcating tree now? 
# Run the checkpoint 1 before this
library(phangorn)
treeRooted = readRDS('tf_analysis/ctree.rds')
# for(this.node in (c(67,69,68,70,72))){
#   treeRooted = ape::rotate(treeRooted, node = this.node)
# }
plot(treeRooted, direction = 'downward')
nodelabels(1:treeRooted$Nnode + length(treeRooted$tip.label))
phydat = BinaryPhyDat(t(tf.expr.mask[tfs.use,]))
anc.acctran = anc_pars(treeRooted, phydat, type = 'ACCTRAN')
```

# Figure 5

## Tree

```{r}
library(ggtree)
library(treeio)

beast_file <- system.file("examples/MCC_FluA_H3.tree", package="ggtree")
beast_tree <- read.beast(beast_file)

genotype_file <- system.file("examples/Genotype.txt", package="ggtree")
genotype <- read.table(genotype_file, sep="\t", stringsAsFactor=F)
colnames(genotype) <- sub("\\.$", "", colnames(genotype))
p <- ggtree(beast_tree, mrsd="2013-01-01") + 
    geom_treescale(x=2008, y=1, offset=2) + 
    geom_tiplab(size=2)
gheatmap(p, genotype, offset=5, width=0.5, font.size=3, 
        colnames_angle=-45, hjust=0) +
    scale_fill_manual(breaks=c("HuH3N2", "pdm", "trig"), 
        values=c("steelblue", "firebrick", "darkgreen"), name="genotype")
p <- ggtree(beast_tree, mrsd="2013-01-01") + 
    geom_tiplab(size=2, align=TRUE, linesize=.5) + 
    theme_tree2()
gheatmap(p, genotype, offset=8, width=0.6, 
        colnames=FALSE, legend_title="genotype") +
    scale_x_ggtree() + 
    scale_y_continuous(expand=c(0, 0.3))
```

```{r}
plot(ctree)
plot(dend)
# plot(ctree2)

ggtree(as.treedata(ctree))+
  geom_tiplab()
# ggtree(as.treedata(treeio::as.phylo(dend)))
```

```{r}
set.seed(2016-05-29)
x <- rtree(50)
p <- ggtree(x) + geom_tiplab()

## nn <- unique(reorder(x, 'postorder')$edge[,1]) 
## to traverse all the internal nodes

nn <- sample(unique(reorder(x, 'postorder')$edge[,1]), 4)

pp <- lapply(nn, function(n) {
    p <- rotate(p, n)
    p + geom_point2(aes(subset=(node == n)), color='red', size=3)
})

plot_list(gglist=pp, tag_levels='A', nrow=1)
```

## Low resolution clades

```{r, fig.height=7, fig.width=4}
clades.broad <- list(
  c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
  c(21, 28), # PDGFRA
  c(36, 5, 37), # VIP
  c(1, 3, 9, 20, 23, 39, 33, 7, 8, 42, 40, 13, 6, 16, 14, 4, 31, 10, 35, 38), # gaba
  c(12, 11, 25, 19, 32, 30), # CRH
  c(29) # not robust
)
clades.broad = lapply(clades.broad, as.character)
names(clades.broad) = 1:length(clades.broad)

# Binary presence doesn't capture some cases
robust.res = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
  sapply(clades.broad, function(clade){
    tree = this$UPGMA
    tree$tip.label = ExtractString(tree$tip.label, after = '_')
    is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades.broad, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades.broad, function(clade){
        tree = this$MP
        tree$tip.label = ExtractString(tree$tip.label, after = '_')
        is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
      })
    })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
        sapply(clades.broad, function(clade){
          tree = this$ML
          tree$tip.label = ExtractString(tree$tip.label, after = '_')
          is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
        })
    }))
)

# Heatmap2(t(apply(robust.res, 1, as.numeric)))
# saveRDS(robust.res, 'tf_analysis/robust.res.binary.broad.rds')

dend = readRDS('tf_analysis/dend.rds')
type.order = labels(dend)
type.order.no = ExtractString(type.order, after = '_')

# Make a type by presence matrix
pres.mat = matrix(rep(type.order.no, nrow(robust.res)), ncol = length(type.order.no), nrow = nrow(robust.res), byrow = T)
pres.mat = t(apply(pres.mat, 1, function(x) as.character(unlist(convert_values(x, invert_list(clades.broad))))))

# Fill with results
pres.mat.filled = pres.mat
for(i in seq_len(nrow(robust.res))){
  for(clade in names(clades.broad)){
    row = pres.mat[i,]
    if(!robust.res[i,as.numeric(clade)]) pres.mat.filled[i,row == clade] = NA
  }
}

method = c(rep('UPGMA', 70), rep('NJ', 70), rep('MP', 70), rep('ML', 70))
final.pres = t(pres.mat.filled[,order(as.numeric(type.order.no))])
rownames(final.pres) = levels(ac.ortho$type_unordered)

# Set 29 to NA
final.pres['29_SLC35D3',] = NA

rownames(final.pres) = orthotype_labels(rownames(final.pres), as.factor = FALSE)

pdf('../../figures/my_figs/tf-robustness-broad.pdf', height = 7, width = 2.2)
draw(Heatmap(final.pres, 
        name = "mat",
        cluster_rows = dend, 
        col = c('1' = "#FF94BB", '2' = "#FFA500", '3' = "#FFCA9E", '4' = "#A5FFFF", '5' = "#FFB1FF"),
        # rect_gp = gpar(col = "grey", lwd = 0.1),
        border_gp = gpar(col = "black", lwd = NA), 
        show_heatmap_legend = FALSE, 
        show_row_dend = TRUE,
        show_row_names = TRUE,
        row_dend_gp = gpar(lwd = 2),
        row_names_gp = gpar(fontsize = 9),
        row_names_side = 'left',
        top_annotation = HeatmapAnnotation(method = method, 
                                   col = list(method = c('white', 'grey', 'brown', 'black') %>% setNames(METHODS)), 
                                   border = TRUE, show_legend = FALSE, show_annotation_name = FALSE)
        ), 
  padding = unit(c(0.1, 0.1, 0.1, 0.1), "mm") # Top, right, bottom, left padding
)
breaks = seq_len(ncol(pres.mat.filled))
for(index in breaks) {
  x_coord = index/length(breaks)
  decorate_heatmap_body("mat", row_slice = 1, {
    # grid.lines(c(0, 1), c(1-x_coord, 1-x_coord), gp = gpar(lty = 1, lwd = 0.75))
  })
}
for(index in 1:3) {
  y_coord = (index * 70)/280
  decorate_heatmap_body("mat", row_slice = 1, {
    grid.lines(c(y_coord, y_coord), c(0, 1), gp = gpar(lty = 2, lwd = 0.75))
  })
}
dev.off()
```

Some fraction of the clade has to be present

```{r, fig.height=7, fig.width=4}
clades.broad <- list(
  c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
  c(21, 28), # PDGFRA
  c(36, 5, 37), # VIP
  c(1, 3, 9, 20, 23, 39, 33, 7, 8, 42, 40, 13, 6, 16, 14, 4, 31, 10, 35, 38), # gaba
  c(12, 11, 25, 19, 32, 30), # CRH
  c(29) # not robust
)
clades.broad = lapply(clades.broad, as.character)
names(clades.broad) = 1:length(clades.broad)

# Just get the fraction 
robust.res = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades.broad[1:5], function(clade){
      find_fraction_nodes(this$UPGMA, clade)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades.broad[1:5], function(clade){
      find_fraction_nodes(this$NJ, clade)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades.broad[1:5], function(clade){
        find_fraction_nodes(this$MP, clade)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades.broad[1:5], function(clade){
        find_fraction_nodes(this$ML, clade)
    })
  }))
)
robust.res = cbind(robust.res, 1) 
colnames(robust.res)[[6]] = '6'

dend = readRDS('tf_analysis/dend.rds')
type.order = labels(dend)
type.order.no = ExtractString(type.order, after = '_')

# Make a type by presence matrix
pres.mat = matrix(rep(type.order.no, nrow(robust.res)), ncol = length(type.order.no), nrow = nrow(robust.res), byrow = T)
pres.mat = t(apply(pres.mat, 1, function(x) as.character(unlist(convert_values(x, invert_list(clades.broad))))))

# Fill with results
pres.mat.filled = pres.mat
for(i in seq_len(nrow(robust.res))){
  for(clade in names(clades.broad)){
    if(robust.res[i,as.numeric(clade)] <= 0.5){
      # row = pres.mat[i,]
      pres.mat.filled[i, row == clade] = NA
      # if(!robust.res[i,as.numeric(clade)]) pres.mat.filled[i,row == clade] = NA
    }
  }
}

method = c(rep('UPGMA', 70), rep('NJ', 70), rep('MP', 70), rep('ML', 70))
final.pres = t(pres.mat.filled[,order(as.numeric(type.order.no))])
rownames(final.pres) = levels(ac.ortho$type_unordered)

# Set 29 to NA
final.pres['29_SLC35D3',] = NA

pdf('../../figures/my_figs/tf-robustness-broad-2.pdf', height = 7, width = 2.2)
draw(Heatmap(final.pres, 
        name = "mat",
        cluster_rows = dend, 
        col = c('1' = "#FF94BB", '2' = "#FFA500", '3' = "#FFCA9E", '4' = "#A5FFFF", '5' = "#FFB1FF"),
        # rect_gp = gpar(col = "grey", lwd = 0.1),
        border_gp = gpar(col = "black", lwd = NA), 
        show_heatmap_legend = FALSE, 
        show_row_dend = TRUE,
        show_row_names = TRUE,
        row_names_side = 'left',
        top_annotation = HeatmapAnnotation(method = method, 
                                           col = list(method = c('white', 'grey', 'brown', 'black') %>% setNames(METHODS)), 
                                           border = TRUE, 
                                           show_legend = FALSE, 
                                           show_annotation_name = FALSE)
        ), 
  padding = unit(c(0.1, 0.1, 0.1, 0.1), "mm") # Top, right, bottom, left padding
)
breaks = seq_len(ncol(pres.mat.filled))
for(index in breaks) {
  x_coord = index/length(breaks)
  decorate_heatmap_body("mat", row_slice = 1, {
    # grid.lines(c(0, 1), c(1-x_coord, 1-x_coord), gp = gpar(lty = 1, lwd = 0.75))
  })
}
for(index in 1:3) {
  y_coord = (index * 70)/280
  decorate_heatmap_body("mat", row_slice = 1, {
    grid.lines(c(y_coord, y_coord), c(0, 1), gp = gpar(lty = 2, lwd = 0.75))
  })
}
dev.off()
```

What's going on with NJ

```{r}
phy1 <- root(all.trees$all$NJ, node = 72)
phy2 <- root(all.trees$all$MP, node = 49)

plot_tanglegram(force.ultrametric(midpoint(loo.trees.list[[4]][[1]]$NJ), method = "extend"), 
                force.ultrametric(midpoint(loo.trees.list[[4]][[1]]$UPGMA), method = "extend"))

plot(midpoint(loo.trees.list[[4]][[1]]$NJ))
# plot(force.ultrametric(midpoint(loo.trees.list[[4]][[1]]$NJ), method = "extend"))
# plot(force.ultrametric(midpoint(loo.trees.list[[4]][[1]]$MP), method = "extend"))
```


```{r}
# Let's document the issues


# Clade 1 is fine 

# Clade 2 is simply not robust
tree = loo.trees.list[[4]][[3]]$NJ
tree$tip.label = ExtractString(tree$tip.label, after = '_')
tip_cols <- ifelse(tree$tip.label %in% clades.broad$`2`, "red", "black")
plot(midpoint(tree), tip.color = tip_cols, cex = 1)
are_adjacent_tips(tree, clades.broad$`2`)

# Clade 3 is robust everywhere except ML
tree = loo.trees.list[[4]][[1]]$ML
tree$tip.label = ExtractString(tree$tip.label, after = '_')
tip_cols <- ifelse(tree$tip.label %in% clades.broad$`3`, "red", "black")
plot(midpoint(tree), tip.color = tip_cols, cex = 1)

# Clade 4 is everywhere except in NJ
# There is an insertion of a clade disrupting it
tree = loo.trees.list[[4]][[1]]$NJ
tree$tip.label = ExtractString(tree$tip.label, after = '_')
tip_cols <- ifelse(tree$tip.label %in% clades.broad$`4`, "red", "black")
plot(midpoint(tree), tip.color = tip_cols, cex = 1)

# Clade 5
tree = loo.trees.list[[4]][[2]]$ML
tree$tip.label = ExtractString(tree$tip.label, after = '_')
tip_cols <- ifelse(tree$tip.label %in% clades.broad$`5`, "red", "black")
plot(midpoint(tree), tip.color = tip_cols, cex = 1)

# New clades to try
# clades.broad <- list(
#   c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
#   c(21, 28, 36, 5, 37), # PDGFRA and VIP
#   c(1, 3, 9, 20, 23, 39, 
#     40, 13, 6), 
#     16, 14, 4, 31, 10, 35, 38), # gaba 33, 7, 8, 42, 
#   c(12, 11, 25, 19, 32, 30), # CRH
#   c(29) # not robust
# )

# clades.broad <- list(
#   c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
#   # c(15, 17, 34, 2, 26, 24), # gly
#   c(41, 27), 
#   c(21, 28), # PDGFRA
#   c(36, 5), # VIP
#   c(40, 13), # NPY
#   c(1, 3, 9, 20), # MAF
#   c(23, 39),
#   c(8, 42), # SAC
#   c(16, 14, 4, 31), # CA2
#   c(33, 7), # CA1
#   c(11, 12, 25), # CRH
#   c(32, 30), # A17
#   c(29, 35, 38, 10, 22, 18, 19, 37, 6) # not robust
# )
```

## High resolution clades

```{r, fig.height=7, fig.width=4}
clades <- list(
  # c(22, 41, 27, 18, 15, 17, 34, 2, 26, 24), # gly
  c(15, 17, 34, 2, 26, 24), # gly
  c(41, 27), 
  c(21, 28), # PDGFRA
  c(36, 5), # VIP
  c(40, 13), # NPY
  c(1, 3, 9, 20), # MAF
  c(23, 39),
  c(8, 42), # SAC
  c(16, 14), # CA2
  c(33, 7), # CA1
  c(11, 12, 25, 19), # CRH
  c(32, 30), # A17
  c(29, 35, 38, 10, 22, 18, 37, 6, 31, 4) # not robust
)
clades = lapply(clades, as.character)
names(clades) = 1:length(clades)

# Binary presence doesn't capture some cases
robust.res = rbind(
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
  sapply(clades, function(clade){
    tree = this$UPGMA
    tree$tip.label = ExtractString(tree$tip.label, after = '_')
    is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
  })
})), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
    sapply(clades, function(clade){
      tree = this$NJ
      tree$tip.label = ExtractString(tree$tip.label, after = '_')
      is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
    })
  })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
      sapply(clades, function(clade){
        tree = this$MP
        tree$tip.label = ExtractString(tree$tip.label, after = '_')
        is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
      })
    })), 
  do.call(rbind, lapply(loo.trees.list[[4]], function(this){
        sapply(clades, function(clade){
          tree = this$ML
          tree$tip.label = ExtractString(tree$tip.label, after = '_')
          is.monophyletic(tree, tips = as.character(clade)) | are_adjacent_tips(tree, clade)
        })
    }))
)

# Heatmap2(t(apply(robust.res, 1, as.numeric)))
# saveRDS(robust.res, 'tf_analysis/robust.res.binary.rds')

dend = readRDS('tf_analysis/dend.rds')
type.order = labels(dend)
type.order.no = ExtractString(type.order, after = '_')

# Make a type by presence matrix
pres.mat = matrix(rep(type.order.no, nrow(robust.res)), ncol = length(type.order.no), nrow = nrow(robust.res), byrow = T)
pres.mat = t(apply(pres.mat, 1, function(x) as.character(unlist(convert_values(x, invert_list(clades))))))

# Fill with results
pres.mat.filled = pres.mat
for(i in seq_len(nrow(robust.res))){
  for(clade in names(clades)){
    row = pres.mat[i,]
    if(!robust.res[i,as.numeric(clade)]) pres.mat.filled[i,row == clade] = NA
  }
}

pdf('../../figures/my_figs/tf-robustness.pdf', height = 7, width = 0.75)
method = c(rep('UPGMA', 70), rep('NJ', 70), rep('MP', 70), rep('ML', 70))
final.pres = t(pres.mat.filled[,order(as.numeric(type.order.no))])
rownames(final.pres) = levels(ac.ortho$type_unordered)
draw(Heatmap(final.pres, 
        name = "mat",
        cluster_rows = dend, 
        col = AcPalette3(length(clades)-1) %>% setNames(names(clades)[1:(length(clades)-1)]), 
        # rect_gp = gpar(col = "grey", lwd = 0.1),
        border_gp = gpar(col = "black", lwd = NA), 
        show_heatmap_legend = FALSE, 
        show_row_dend = FALSE,
        show_row_names = FALSE,
        row_names_side = 'left',
        top_annotation = HeatmapAnnotation(method = method, 
                                           col = list(method = c('white', 'grey', 'brown', 'black') %>% setNames(METHODS)), 
                                           border = TRUE, show_legend = FALSE, show_annotation_name = FALSE)
        ),
padding = unit(c(0.1, 0.1, 0.1, 0.1), "mm") # Top, right, bottom, left padding
)
breaks = seq_len(ncol(pres.mat.filled))
for(index in breaks) {
  x_coord = index/length(breaks)
  decorate_heatmap_body("mat", row_slice = 1, {
    # grid.lines(c(0, 1), c(1-x_coord, 1-x_coord), gp = gpar(lty = 1, lwd = 0.75))
  })
}
for(index in 1:3) {
  y_coord = (index * 70)/280
  decorate_heatmap_body("mat", row_slice = 1, {
    grid.lines(c(y_coord, y_coord), c(0, 1), gp = gpar(lty = 2, lwd = 0.75))
  })
}
dev.off()
```

## Gene panels

```{r, fig.width=11, fig.height=7.5}
tf.show = c(
  
  # Pan
  # 'TFAP2A', 
  'TFAP2B', 
  
  # Gaba
  'MEIS2', 'ZEB2', 
  'SOX5', 'DACH2', 'LHX9', 'ZNF536',
  'POU6F2', 'ZFHX3', 'ZFHX4', 
  'TSHZ1', 
  'TBX3',
  'LHX2',
  'TFAP2D',
  'TBX2', 
  # 'SIX3', 
  'NR4A2', 
  'BHLHE22',
  'PRDM8', 
  
  # Gly
  "TCF4", 'ZNF804A',
  
  "NEUROD2", 
  # "TFAP2B",
  # "TFAP2A", 
  'NEUROD1', 'ESRRG', 'DACH1', 'NFIA', "PROX1", "ST18", 'ONECUT1', # A2 enriched
  "FOXP2", "BNC2", "BNC1",
  
  "PRDM13", "EBF1", #"PROX1", "ST18",
  "EBF3",
  "SATB2", "NEUROD4", "NFIB", "NFIX"
)

ac.ortho$orthotype.tf = factor(orthotype_labels(ac.ortho$type),
                               levels = (c("oAC11", "oAC14 [A17]", "oAC13 [A17]", "oAC37* [CRH]", "oAC38* [nNOS]",
                                          "oAC23*", "oAC22*", "oAC39* [CA1]", "oAC28", "oAC12", "oAC21*",
                                          "oAC40 [nNOS]", "oAC18 [nNOS]", "oAC36* [nGnG-ga]", "oAC19 [CA2]",
                                          "oAC27", "oAC26", "oAC42* [SAC]", "oAC35", "oAC34", "oAC29", "oAC15",
                                          "oAC17*", "oAC16*", "oAC20 [NPY]", "oAC25", "oAC24", "oAC41",
                                          "oAC33 [VIP]", "oAC32", "oAC31", "oAC30 [PDGFRA]", "oAC1 [A2]",
                                          "oAC2 [VG3]", "oAC4", "oAC3", "oAC9", "oAC10", "oAC5",
                                          "oAC8 [nGnG-gl]", "oAC7 [SEG]", "oAC6 [A8]")))

# pdf('../../figures/my_figs/tf-show.pdf', height=9, width=5.2)
# ClusteredDotplot2(ac.ortho, 
#                   features = tf.show, 
#                   group.by = 'orthotype.tf', 
#                   # cluster_rows = dend,
#                   max.pct = 40, 
#                   dot.scale.factor = 0.35, 
#                   pseudocount = 5, 
#                   rotate = TRUE, 
#                   show.legend = FALSE, 
#                   show_row_names = FALSE, 
#                   show_row_dend = FALSE, 
#                   border_gp = gpar(col = "black", lwd = 3),
#                   column_names_side = 'top')
# dev.off()

bin.mat = tf.expr.mask[tf.show,]
colnames(bin.mat) = orthotype_labels(colnames(bin.mat), asterisk = FALSE)
bin.mat = bin.mat[,levels(ac.ortho$orthotype.tf)]
DotPlot3(ac.ortho, 
        features = tf.show, 
        group.by = 'orthotype.tf', 
         max.pct = 40, 
        max.size = 4, 
        binarization = bin.mat) + 
  # scale_y_discrete(position = "right")+
  guides(y = guide_axis(position = "right")) +
  theme(
    axis.text.y.left  = element_blank(),
    axis.ticks.y.left = element_blank(),
    axis.title.y.left = element_blank(),
    axis.text.y = element_text(size = 11), 
    axis.ticks.y = element_blank(), 
    axis.ticks.length.y = unit(0, 'cm')
  )+
  theme(axis.text.x = element_blank(), 
       axis.ticks.x = element_blank(), 
       axis.line = element_blank(),
       panel.border = element_rect(fill = NA, colour = "black", linewidth=1))

ggsave('../../figures/my_figs/tf-show.pdf', width=11, height=5.85)
```


```{r, fig.width=11, fig.height=5.2}
# adhesion.show = c('DSCAM', 'UNC5D', 'TENM3', 
#                   'CLMP', 'CDH10', 'CDH8', 'PCDH10', 
#                   'PCDH19', 
#                   #'PCDH17',
#                   'DSCAML1', 'IGSF11', 'CDH4')
# 
# pdf('../../figures/my_figs/adhesion-show.pdf', height=9, width=2)
# ClusteredDotplot2(ac.ortho, 
#                   features = adhesion.show, 
#                   group.by = 'type_clean', 
#                   cluster_rows = dend,
#                   max.pct = 40, 
#                   dot.scale.factor = 0.35, 
#                   pseudocount = 5, 
#                   col.high = '#934E93',
#                   rotate = TRUE, 
#                   show.legend = FALSE, 
#                   show_row_names = FALSE, 
#                   show_row_dend = FALSE, 
#                   border_gp = gpar(col = "black", lwd = 3),
#                   column_names_side = 'top')
# dev.off()
```

Other TFs

```{r}
ac.ortho$type_tf = factor(gsub('\\*', '', ac.ortho$type), levels = labels(dend))
geneDotPlotFast(ac.ortho, group.by = 'type_tf', gene = 'TBX3')
```


# Co-mentions in literature

```{r}
tfs.considered = intersect(gene.lists$tf_genes, conservation.summary$gene)
write.table(tfs.considered, 'tf_analysis/tfs.considered.tsv', quote = FALSE, row.names = FALSE, col.names = FALSE)
pubmed.counts = lapply(tfs.considered, function(gene){
  pubmed_command <- paste0("curl -s 'https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi?db=pubmed&term=", gene, "+%22amacrine%22' | grep '<Count>' | sed -e 's/<Count>//' -e 's/\\s*<\\/Count>//'")
  result <- system(pubmed_command, intern = TRUE)
  as.numeric(ExtractString(result, before = '<eSearchResult>', after = '<RetMax>'))
}) %>% setNames(tfs.considered)
pubmed.counts

saveRDS(pubmed.counts, 'tf_analysis/pubmed.counts.rds')

pubmed.counts = readRDS('tf_analysis/pubmed.counts.rds')
t(t(sort(unlist(pubmed.counts[pubmed.counts>0]))))
```

```{r}
system.time({
  pubmed.counts = lapply(tf.show, function(gene){
  pubmed_command <- paste0("curl -s 'https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi?db=pubmed&term=", gene, "+%22amacrine%22' | grep '<Count>' | sed -e 's/<Count>//' -e 's/\\s*<\\/Count>//'")
  result <- system(pubmed_command, intern = TRUE)
  as.numeric(ExtractString(result, before = '<eSearchResult>', after = '<RetMax>'))
}) %>% setNames(tf.show)
  })

# Doesn't store all values properly
system.time({
  pubmed.counts = mclapply(tf.show, function(gene){
    pubmed_command <- paste0("curl -s 'https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi?db=pubmed&term=", gene, "+%22amacrine%22' | grep '<Count>' | sed -e 's/<Count>//' -e 's/\\s*<\\/Count>//'")
    result <- system(pubmed_command, intern = TRUE)
    as.numeric(ExtractString(result, before = '<eSearchResult>', after = '<RetMax>'))
  }, mc.cores = 20, mc.cleanup = FALSE) %>% setNames(tf.show)
})

t(t(sort(unlist(pubmed.counts[pubmed.counts>0]))))
```

Enrichment test

```{r}
# All TFs
all.tfs = names(pubmed.counts)
mentioned.tfs = names(pubmed.counts[pubmed.counts > 0])
# Used: tfs.use
mentioned.used = intersect(tfs.use, mentioned.tfs)
VennDiagram(mentioned.tfs, tfs.use)

t(t(sort(unlist(pubmed.counts[names(pubmed.counts) %in% tfs.use]))))
```

# Enrichment for protein-protein interactions

## Cooperativity events

```{r, fig.height=2, fig.width=4}
tf.switches.clean = readRDS('tf_analysis/tf.switches.clean.rds')

library(STRINGdb)

# Setup
db <- STRINGdb$new(species=10090, score_threshold = 0, version="12")

# Run it all in a batch
genes <- unique(unlist(tf.switches.clean))
map <- db$map(data.frame(symbol = genes), "symbol", removeUnmappedRows=TRUE)

# Get interactions
interactions <- db$get_interactions(map$STRING_id)

# Merge to get symbols back
merged <- merge(
    interactions,
    map[, c("STRING_id", "symbol")],
    by.x="from", by.y="STRING_id", all.x=TRUE
)

merged <- merge(
    merged,
    map[, c("STRING_id", "symbol")],
    by.x="to", by.y="STRING_id", all.x=TRUE,
    suffixes = c("_from", "_to")
)

PrettyHistogram2(merged$combined_score, vline = c(400, 700))

# 2.4k / 2 = 1.2k interactions (each is given twice)

merged.unique = unique(merged)

# How many interactions are predicted using our tree? 
ints.list = lapply(tf.switches.clean, function(x) subset(merged.unique, symbol_from %in% x & symbol_to %in% x))
ints.all = do.call(rbind, ints.list)
saveRDS(ints.all, 'tf_analysis/ints.all.rds')
ints = (subset(ints.all, combined_score > 400))
# They need not be unique, since we want to count interactions at muliple sites in the tree
# Found 43 interactions

# Scramble the list, while keeping the structure and repeat
null.dist = sapply(1:10000, function(i){
  shuffled_list = shuffle_values(tf.switches.clean)
  ints.list = lapply(shuffled_list, function(x) subset(merged.unique, symbol_from %in% x & symbol_to %in% x))
  ints.all = do.call(rbind, ints.list)
  ints = (subset(ints.all, combined_score > 400)) 
  nrow(ints)
})
```


```{r, fig.height=2, fig.width=4}
observed = nrow(ints)
PrettyHistogram2(null.dist, vline = observed, bins = 20, line.col = 'deeppink', 
                 fill = 'white', xlab = 'Significant protein-protein interactions') +
  ArialFont()
ggsave('../../figures/my_figs/fig-ppi-hist.pdf')
message('permutation p-value: ', (sum(null.dist >= observed) + 1)/(length(null.dist) + 1))
```

## Novel roles for known transcriptional modules

```{r, fig.height=4, fig.width=7}
#' Run check point 2 first 
pubmed.counts = unlist(readRDS('tf_analysis/pubmed.counts.rds'))
ints.all = readRDS('tf_analysis/ints.all.rds')

ints.all$refs_from = pubmed.counts[ints.all$symbol_from]
ints.all$refs_to = pubmed.counts[ints.all$symbol_to]
ints.all$refs_min = apply(ints.all[,c('refs_from', 'refs_to')], 1, min)
ints.all$label = paste0(ints.all$symbol_from, '-', ints.all$symbol_to)
ints.all$label[ints.all$combined_score < 400] = ''

# ScatterPlot(unique(ints.all), x = 'combined_score', y = 'refs_min', labels = 'label', max.overlaps = 100) + 
#   geom_vline(xintercept = c(400), linetype = 'dashed', color = 'deeppink')

# Add expression correlations
expr.mat = LogAvgExpr(ac.ortho, group.by = 'type', assay = 'RNA')
cor.mat = cor(t(expr.mat[tfs.use,]))
ints.all$expr.cor = apply(ints.all, 1, function(x) cor.mat[x[['symbol_from']], x[['symbol_to']]])

ScatterPlot(unique(ints.all), 
            x = 'expr.cor', 
            y = 'combined_score', 
            fill = 'expr.cor',
            pt.col = 'black',
            # labels = 'label', 
            xlab = 'Expression correlation', 
            ylab = 'PPI score',
            max.overlaps = 5, 
            pt.size = 1.5,
            # pt.shape = 21,
            label.size = 3) + 
  theme_cowplot()+
  NoLegend()+
  scale_fill_gradient2(low = 'cyan2', mid = 'grey', high = 'deeppink', midpoint = 0)+
  geom_hline(yintercept = c(400), linetype = 'dashed', color = 'grey') + 
  geom_text_repel(aes(label = label), fontface = 'italic', family = 'ArialMT', size = 2.5) + 
  ArialFont()

ggsave('../../figures/my_figs/fig-ppi-cor.pdf', width = 6.5, height = 3)
```



```{r, fig.height=3}
ClusteredDotplot2(ac.ortho, 
                  features = c('NEUROD1', 'ONECUT1', 'ZEB1', 'ZEB2', 'TBX3', 'PBX1', 'MEIS1', 'MEIS2', 'PBX3'), 
                  group.by = 'type_clean', 
                  cluster_columns = dend,
                  max.pct = 40, 
                  dot.scale.factor = 0.65, 
                  pseudocount = 5, 
                  rotate = FALSE, 
                  species.size = 0.40, 
                  show.legend = FALSE, 
                  show_row_names = TRUE, 
                  show_row_dend = FALSE, 
                  border_gp = gpar(col = "black", lwd = 3),
                  column_names_side = 'top')

DotPlot3(mouse.ac, features = c('NOS1', 'LHX9'))
```


