---
title: "Sub-cluster SACs"
output:
  html_document:
    keep_md: true
    toc: true
date: "2023-07-28"
params:
  OT: FALSE
  manual_annotation: list(Mouse = 2, Rat = 0, Peromyscus = c(2,3), Squirrel = 1, Rhabdomys = 1, TreeShrew = 1, Cow = 0, Sheep = 0, Opossum = 1, Chicken = 0, Lizard = 1, Zebrafish = 0, Killifish = 1, Lamprey = 1)
  color1: 'deeppink'
  color2: 'orange'
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 8, dev = 'png', dpi = 300)
source("../../utils/dario_functions.R")
LoadLibraries(load.lisi = FALSE)
SourceFiles()

ON_SAC_markers = c("FEZF1", 'FEZF2', "GFRA3", "CNTN5", "SLIT2")
OFF_SAC_markers = c("TENM3", "ZFHX3", "RND3")
```

# Using OrthoType analysis fails 

We reach the batch correction limit

```{r, eval = params$OT}
mammalAC = readRDS('../../Ortho_Objects/mammalAC_v2.rds')
# subsetMammalAC = subset(mammalAC, species %in% c("Rat", "Peromyscus", "Squirrel", "Rhabdomys", "TreeShrew", "Cow", "Sheep", "Chicken", "Lizard", "Lamprey") & lit_type == "SAC")
subsetMammalAC = ClusterSeurat(subsetMammalAC, integrate.by = "species", cluster_resolution = 0.1)

# Find resolution
DefaultAssay(subsetMammalAC) = "integrated"
plotClusterRes(subsetMammalAC, from = 0.05, to = 1, by = 0.05)
subsetMammalAC = FindClusters(subsetMammalAC, resolution = 0.05)

# Plot
DefaultAssay(subsetMammalAC) = "RNA"
BrowseSeurat(subsetMammalAC, batch = "species")
FeaturePlot(subsetMammalAC, features = c(ON_SAC_markers, OFF_SAC_markers))
```

Split by quality, not ON and OFF

```{r, eval = params$OT}
verterbateAC = readRDS('../../Ortho_Objects/vertebrateAC_v2.rds')
# subsetVertAC = subset(verterbateAC, species %in% c("Rat", "Peromyscus", "Squirrel", "Rhabdomys", "TreeShrew", "Cow", "Sheep", "Chicken", "Lizard", "Lamprey") & lit_type == "SAC")
subsetVertAC = ClusterSeurat(subsetVertAC, integrate.by = "species", cluster_resolution = 0.2)
DefaultAssay(subsetVertAC) = "integrated"
subsetVertAC = FindClusters(subsetVertAC, resolution = 0.4)
DefaultAssay(subsetVertAC) = "RNA"
BrowseSeurat(subsetVertAC, batch = "species")
FeaturePlot(subsetVertAC, features = c(ON_SAC_markers, OFF_SAC_markers))

# Score
subsetVertAC = ScaleData(subsetVertAC, features = c(OnSacMarkers, OffSacMarkers))
subsetVertAC$ON.SAC.score = ScoreCelltype(subsetVertAC, OnSacMarkers)
subsetVertAC$OFF.SAC.score = ScoreCelltype(subsetVertAC, OffSacMarkers)
plotSACMarkers(subsetVertAC)
```

# Read in species data

```{r, fig.height=4, fig.width=8}
# No SACs found in ferret due to low sampling of ACs, will omit
speciesList = phylogenetic_order2[!phylogenetic_order2 %in% c("Ferret", 'Goldfish')]
objectList = lapply(speciesList, function(species) {
    if(species == "Mouse") {
      readRDS("../../Species_Reference/MouseACref_v4.rds")
    } else if(species == 'Zebrafish'){
      readRDS("../../Species_Objects/ZebrafishAC_no_HCs_v6_converted.rds")
    } else {
      readRDS(paste0("../../Species_Objects/", species, "AC_v6.rds"))
    }
  })
names(objectList) = speciesList

objectList$Killifish = readRDS('../../Species_Objects/Killifish_ncbiAC_v6.rds')
objectList$Killifish = UpperCase_genes(objectList$Killifish)
```

## Annotate ON and OFF SACs

Some naturally split into two clusters:

* Squirrel
* Rat
* Rhabdomys

The rest need a supervised approach: 

* Get DE genes between ON and OFF SACs
* Score each cell
* Use Gaussian mixture models to define valid cutpoints

## Find useful genes

Using mouse P0 from Peng et al. 2020

```{r}
PengAC = readRDS("../../Species_Reference/PengAC_v3.rds")
BrowseSeurat(PengAC, batch = "batch")
# Cluster 0 = ON SACs
# Cluster 1 = OFF SACs
PengAC$lit_type = ifelse(PengAC$seurat_clusters == 0, "ON", "OFF")
Idents(PengAC) = "lit_type"

plotSACMarkers(PengAC, feature.plot = TRUE)
```

```{r}
PengMarkers = wilcoxauc(PengAC, 'seurat_clusters')
OnSacMarkers = subset(PengMarkers, group == 0 & logFC > 0.25 & padj < 0.001)$feature
OffSacMarkers = subset(PengMarkers, group == 1 & logFC > 0.25 & padj < 0.001)$feature

OnSacMarkers
OffSacMarkers
```

## Score the cells (average z-scored expression)

```{r}
# Subset to SACs
SACList = lapply(seq_along(objectList), function(index) {
  print(names(objectList)[[index]])
  subset(objectList[[index]], lit_type == "SAC")
})
names(SACList) = speciesList

# Add mouse with only Yan SACs to avoid batch effects 
# mouseAC = readRDS("../../Species_Reference/MouseACref_v4.rds")
SACList$Mouse = subset(SACList$Mouse, lit_type == "SAC" & source == "Yan")

ScoredList = lapply(SACList, function(object){
  object = AssignName(object)
  object = ScaleData(object, features = c(OnSacMarkers, OffSacMarkers))
  object$ON.SAC.score = ScoreCelltype(object, OnSacMarkers)
  object$OFF.SAC.score = ScoreCelltype(object, OffSacMarkers)
  return(object)
})
```

# Successful ON and OFF SAC annotations

## Mouse

Poor separation between ON and OFF

```{r, fig.height=5, fig.width=8, eval = TRUE}
ScoredList$Mouse = SubclusterSACs(ScoredList$Mouse, group.by = "batch", resolution = 0.3, harmonize = FALSE)

# Remove the bad quality cells
ScoredList$Mouse = subset(ScoredList$Mouse, seurat_clusters %in% c(1,2))
```

## Rat

```{r}
ScoredList$Rat = SubclusterSACs(ScoredList$Rat, resolution = 0.1)
```

## Peromyscus 

```{r}
ScoredList$Peromyscus = SubclusterSACs(ScoredList$Peromyscus, resolution = 0.1)
```

## Squirrel

```{r}
ScoredList$Squirrel = SubclusterSACs(ScoredList$Squirrel, resolution = 0.1)
```

## Rhabdomys

```{r}
ScoredList$Rhabdomys = SubclusterSACs(ScoredList$Rhabdomys, group.by = "orig.file", resolution = 0.1)
```

## Treeshrew 

```{r}
ScoredList$TreeShrew = SubclusterSACs(ScoredList$TreeShrew, resolution = 0.1)
```

## Cow

```{r}
ScoredList$Cow = SubclusterSACs(ScoredList$Cow, resolution = 0.1)
```

## Sheep

```{r}
ScoredList$Sheep = SubclusterSACs(ScoredList$Sheep, resolution = 0.08)
```

## Opossum

```{r}
ScoredList$Opossum = SubclusterSACs(ScoredList$Opossum, resolution = 0.08)

# Remove the bad quality cells
ScoredList$Opossum = subset(ScoredList$Opossum, seurat_clusters %in% c(0,1))
```

## Chicken

```{r}
ScoredList$Chicken = SubclusterSACs(ScoredList$Chicken, resolution = 0.2, group.by = "orig.file")
```

## Lizard

```{r}
ScoredList$Lizard = SubclusterSACs(ScoredList$Lizard, resolution = 0.1, group.by = "orig.file")
```

## Zebrafish

```{r}
ScoredList$Zebrafish = SubclusterSACs(subset(ScoredList$Zebrafish, sample == 'LD_14D'), resolution = 0.1, group.by = "sample", harmonize = FALSE)
# DimPlotLabeled(ScoredList$Zebrafish, group.by = 'sample')
# DimPlotLabeled(ClusterSeurat(subset(ScoredList$Zebrafish, sample == 'Control_2'), cluster_resolution = 0.1))
```

## Killifish

```{r}
# Convert to zebrafish symbols 
ScoredList$Killifish = RenameFeatures(ScoredList$Killifish, gsub('ZFHX3B', 'ZFHX3', gsub('RND3A', 'RND3', rownames(ScoredList$Killifish))))
ScoredList$Killifish = SubclusterSACs(ScoredList$Killifish, resolution = 0.1, group.by = "alias", harmonize = TRUE)

# VlnPlot(ScoredList$Killifish, features = c('TENM3', 'ZFHX3B', 'RND3A'))
```

## Lamprey

The MSTRG ID for TENM3 is MSTRG.4792 and the LOC ID is LOC116944204.

```{r}
LampreySAC = ScoredList$Lamprey
LampreySAC = Harmonize(LampreySAC, batch = "orig.file", cluster_resolution = 0.1, show.plots = FALSE)
BrowseSeurat(LampreySAC, batch = "orig.file")
LampreySAC = TransferSACLabels(LampreySAC)
FeaturePlot(LampreySAC, features = c("LOC116944204", "FEZF2"))
rownames(LampreySAC@assays$RNA@counts)[rownames(LampreySAC@assays$RNA@counts) == "LOC116944204"] = "TENM3"
rownames(LampreySAC@assays$RNA@data)[rownames(LampreySAC@assays$RNA@data) == "LOC116944204"] = "TENM3"
ScoredList$Lamprey = LampreySAC
```

# Failures

## Human

```{r}
ScoredList$Human = SubclusterSACs(ScoredList$Human, group.by = "donor")
```

## Macaque

```{r}
ScoredList$Macaque = SubclusterSACs(ScoredList$Macaque)
```

## Marmoset

```{r}
ScoredList$Marmoset = SubclusterSACs(ScoredList$Marmoset, resolution = 0.5)
```

## Mouse lemur

Separate by quality

```{r}
ScoredList$MouseLemur = SubclusterSACs(ScoredList$MouseLemur, resolution = 0.8)
```

## Pig

Separate into two clusters but not ON and OFF

```{r}
ScoredList$Pig = SubclusterSACs(ScoredList$Pig, resolution = 0.4)
```


```{r}
pigac = readRDS('../../Species_Objects/PigAC_v6.rds')
pigsacs = subset(pigac, type == '14_SAC')
pigsacs

pigsacs = ClusterSeurat(pigsacs, cluster_resolution = 1)
DimPlotLabeled(pigsacs)
FeaturePlot(pigsacs, c('TENM3', 'ZFHX3', 'FEZF1'))
DotPlot(pigsacs, features = c('TENM3', 'ZFHX3', 'RND3', 'TLE1', 'FEZF1', 'FEZF2'))
```

## Save 

```{r, eval = FALSE}
saveRDS(ScoredList, "../../Species_Reference/ScoredSacList.rds")
```

Addition of zebrafish and lizard

```{r}
ze = ScoredList$Zebrafish
ScoredList_old = readRDS("../../Species_Reference/ScoredSacList.rds")
ScoredList_old$Zebrafish = ScoredList$Zebrafish
saveRDS(ScoredList_old, "../../Species_Reference/ScoredSacList_v2.rds")
```

Addition of re-aligned lizard

```{r}
ScoredList = readRDS("../../Species_Reference/ScoredSacList_v2.rds")

# Run lizard again with v6 object before saving
saveRDS(ScoredList, "../../Species_Reference/ScoredSacList_v3.rds")
```

Addition of killifish

```{r}
ScoredList = readRDS("../../Species_Reference/ScoredSacList_v3.rds")

# Run killifish
ScoredList$Lizard = lizard
ScoredList = append(ScoredList, list(Killifish = kfish), after = 17)
saveRDS(ScoredList, "../../Species_Reference/ScoredSacList_v4.rds")
```

# Summary

Checkpoint

```{r}
manual_annotation = eval(parse(text = params$manual_annotation))

# options(ggplot2.discrete.colour = list(c("red3", "gold")), 
        # ggplot2.discrete.fill = list(c("red3", "gold")))
```

```{r}
ScoredList = readRDS("../../Species_Reference/ScoredSacList_v4.rds")

successful_species = names(manual_annotation)
successList = ScoredList[names(ScoredList) %in% successful_species]
successList = lapply(seq_along(successList), function(index){
  species = names(successList)[[index]]
  on_sac_clusters = manual_annotation[[species]]
  successList[[index]]$lit_subtype = "OFF"
  successList[[index]]$lit_subtype[successList[[index]]$seurat_clusters %in% on_sac_clusters] = "ON"
  return(successList[[index]])
  })
names(successList) = names(ScoredList[names(ScoredList) %in% successful_species])
```

## UMAP

```{r sac-umap, fig.height=9, fig.width=15}
spacer = 0.1
angle.list = c(0, 0, 0, 120, -70, 170, 135, 0, 0,-45, 160, -60, -90, 130) * pi/180
umap.list = lapply(seq_along(successList), function(index) 
  # theme_umap(DimPlot(successList[[index]], group.by = "lit_subtype", label = TRUE), #cols = c(params$color1, params$color2), pt.size = 0.5), 
             # remove.axes = FALSE, title = names(successList)[[index]]) + 
  TitlePlot(PrettyUmap2(successList[[index]], group.by = 'lit_subtype', label = TRUE, 
                        pad = 1, geom.label = geom_label, color = 'black', fill = 'white', 
                      pt.alpha = 1, label.size = 0, alpha = 0.8, label.padding = unit(0, "lines"), 
                      label.r = unit(0.2, "lines"), angle = angle.list[[index]]), 
            names(successList)[[index]])+
    theme(plot.margin = unit(c(spacer, spacer, spacer, spacer), "cm")))
# ggarrange(plotlist = umap.list, ncol = 5, nrow = 3, common.legend = TRUE, legend = "right")
ggarrange(plotlist = PrepStack(plt.list = umap.list, ncol = 5, nrow = 3), 
          ncol = 5, nrow = 3, common.legend = TRUE, legend = "right", align = 'hv')
ggsave("../../figures/my_figs/sac-umap-1.pdf", width=15/1.4, height = 9/1.4)
```

# Scatter plots

```{r sac-cor, fig.height=12, fig.width=16}
scatterplot.list = lapply(seq_along(successList), function(index){
                      object = successList[[index]]
                      avg_exp = as.data.frame(log1p(AverageExpression(object, verbose = FALSE, group.by = "lit_subtype")$RNA))
                      p1 = ExpressionScatter(avg_exp, x = "ON", y="OFF", logX = FALSE, logY = FALSE, 
                                             genes.to.label = c("TENM3", "ZFHX3", "RND3", "FEZF1", "FEZF2"), 
                                             title = names(successList)[[index]])
                      return(p1)
})

ggarrange(plotlist = scatterplot.list, ncol = 5, nrow = 3) %>% suppressWarnings()
```

# ON and OFF type dotplot

```{r, fig.height=5, fig.width=11}
merged <- merge(successList[[1]], y = successList[2:length(successList)], add.cell.ids = names(successList))

# merged = Reduce(merge, successList)
merged$species = rep(names(successList), sapply(successList, ncol))

# DimPlot(merged, split.by = "species", group.by = "lit_subtype", cols = c("blue", "red"))
```

```{r sac-violin, fig.height=7, fig.width=6}
# Scale by species
split = SplitObject(merged, split.by = "species")
merged$`FEZF1_FEZF2` = unlist(lapply(split, function(object) rescale(object@assays$RNA@data["FEZF1",]+object@assays$RNA@data["FEZF2",])))
merged$`TENM3_ZFHX3_RND3` = unlist(lapply(split, function(object) rescale(object@assays$RNA@data["TENM3",]+object@assays$RNA@data["ZFHX3",]+object@assays$RNA@data["RND3",])))
merged$species = factor(merged$species, levels = rev(phylogenetic_order2))

plot_grid(VlnPlot(merged, 
                  features = c("FEZF1_FEZF2", "TENM3_ZFHX3_RND3"), 
                  group.by = "species", 
                  cols = c(params$color1, params$color2),
                  split.by = "lit_subtype", 
                  split.plot = FALSE, 
                  pt.size = 0, 
                  stack = TRUE) + NoLegend() + theme(axis.title.y = element_blank()), 
          stackedBarGraph(merged, feature.2 = "species", feature.1 = "lit_subtype") + 
            labs(y = 'Proportion')+
            guides(fill = guide_legend(title = "Subtype"))+
            # scale_fill_manual(values = c(params$color1, params$color2))+
            coord_flip() + 
            theme(axis.text.y = element_blank(), axis.title.y = element_blank()), 
          ncol = 2, rel_widths = c(1.2,1), align = "h", axis = "bt")

ggsave('../../figures/my_figs/sac-violn-1.pdf', height = 7.5, width = 6)
# VlnPlot(merged, split.by = "lit_subtype", features = c("FEZF1", "FEZF2", OFF_SAC_markers), group.by = "species", split.plot = FALSE, ncol = 5)
# VlnPlot(merged, split.by = "lit_subtype", features = c("ON.SAC.score", "OFF.SAC.score"), group.by = "species", split.plot = FALSE, ncol = 5)
# ggsave("../../figures/files/sac-violin-1.png", height=4.5, width = 8)
```


```{r}
saveRDS(merged, '../../Ortho_Objects/sac-merged_v1.rds')
```

# Cross-species analysis of ON and OFF SACs

```{r, fig.height=5, fig.width=5}
BIGSEURAT = readRDS("../../Ortho_Objects/vertebrateAC_BIGSEURAT.rds")
merged = readRDS('../../Ortho_Objects/sac-merged_v1.rds')
BIGSEURAT$lit_subtype = merged$lit_subtype[match(Cells(BIGSEURAT), Cells(merged))]

sac_subset = subset(BIGSEURAT, lit_subtype %in% c('ON', 'OFF'))
table(sac_subset$species)
saveRDS(sac_subset, '../../Ortho_Objects/sac_subset_v4.rds')
```

```{r, fig.height=3, fig.width=7}
sac_subset = readRDS('../../Ortho_Objects/sac_subset_v3.rds')
# de.on.off = FindAllMarkersFast(sac_subset, group.by = 'lit_subtype')
de.on.off = FindMarkersFast(sac_subset, group.by = 'lit_subtype', ident.1 = 'ON', ident.2 = 'OFF', p_val_adj_cutoff = 1, avg_log2FC_cutoff = 0)

# geneDotPlot(sac_subset, 'TENM3', group.by = 'lit_subtype')
# geneDotPlot(sac_subset, 'TRHDE', group.by = 'lit_subtype')
# geneDotPlot(sac_subset, 'LRP1B', group.by = 'lit_subtype')

merged$species_rev = factor(merged$species, levels = (phylogenetic_order))
VlnPlot(merged, features = c('TENM3', 'ZFHX3', 'TRHDE', 'LRP1B', 'POU6F2'), group.by = "species", cols = c(params$color1, params$color2), split.by = "lit_subtype", split.plot = FALSE, pt.size = 0, stack = TRUE)
VlnPlot(merged, features = c('ADGRL2', 'KIRREL3', 'TLE1'), group.by = "species", cols = c(params$color1, params$color2), split.by = "lit_subtype", split.plot = FALSE, pt.size = 0, stack = TRUE, flip = TRUE)

VlnPlot(merged, features = c('TENM3', 'ADGRL2'), group.by = "species_rev", cols = c(params$color1, params$color2), split.by = "lit_subtype", split.plot = FALSE, pt.size = 0, stack = TRUE, flip = TRUE)
```

```{r, fig.height=5, fig.width=7}
sac_subset = readRDS('../../Ortho_Objects/sac_subset_v4.rds')
# de.on.off = FindAllMarkersFast(sac_subset, group.by = 'lit_subtype')
de.on.off = FindMarkersFast(DownsampleSeurat(sac_subset, group.by = c('species', 'lit_subtype'), size = 50), group.by = 'lit_subtype', ident.1 = 'ON', ident.2 = 'OFF', p_val_adj_cutoff = 1, avg_log2FC_cutoff = 0)

# volcanoPlot(de.on.off, fc_cutoff = 0.25, fdr_cutoff = 0.01, labels = TRUE, max.overlaps = 100, size = 2, max_fdr = 1e-10)
subset(de.on.off, gene == 'ADGRL2')
subset(de.on.off, gene == 'SLIT2')

# pdf('../../figures/my_figs/sac-volcano-plot.pdf', height = 5, width = 7)
TitlePlot(volcanoPlot(de.on.off, 
                      fc_cutoff = 0.25, 
                      fdr_cutoff = 0.01, 
                      labels = TRUE, 
                      max.overlaps = 10, 
                      max_fdr = 1e-11, 
                      color.signif = 'deeppink',
                      name1 = 'OFF-enriched', 
                      name2 = 'ON-enriched', 
                      anno.y = 0.4), 'ON vs OFF SAC')
# dev.off()
ggsave('../../figures/my_figs/sac-volcano-plot-2.pdf', height = 5, width = 7)
```

## Module score

```{r, fig.height=6, fig.width=8}
rownames(de.on.off) = de.on.off$gene
on_genes <- nDEGs(de.on.off, up = TRUE, down = FALSE, log2FC.threshold = 0.25, p_val_adj.threshold = 0.05, value = TRUE)
off_genes <- nDEGs(de.on.off, up = FALSE, down = TRUE, log2FC.threshold = 0.25, p_val_adj.threshold = 0.05, value = TRUE)
sac_subset <- AddModuleScore(sac_subset, features = list(on_genes), name="ON_SAC_score")
# DimPlot(sac_subset, split.by = 'species', group.by = 'lit_subtype')

write.csv(on_genes, 'specificity_analysis/on_sac_genes.csv')
write.csv(off_genes, 'specificity_analysis/off_sac_genes.csv')

umap.list = lapply(seq_along(successList), function(index) {
  successList[[index]] = AddModuleScore(successList[[index]], features = list(off_genes), name="OFF_SAC_score")
  successList[[index]] = AddModuleScore(successList[[index]], features = list(on_genes), name="ON_SAC_score")
  successList[[index]]$SAC_SCORE = successList[[index]]$ON_SAC_score1 - successList[[index]]$OFF_SAC_score1
  theme_umap(ClusterFeaturePlot(successList[[index]], group.by = 'lit_subtype', features = 'SAC_SCORE'), 
             remove.axes = FALSE, title = names(successList)[[index]])
  })
ggarrange(plotlist = umap.list, ncol = 4, nrow = 3, common.legend = TRUE, legend = "right")
```

## Correlation between log2FC of ON vs OFF SACs across species

```{r, fig.height=7, fig.width=7}
sac_subset = readRDS('../../Ortho_Objects/sac_subset.rds')
genes.of.interest = c('FEZF1', 'FEZF2', 'TENM3', 'ZFHX3', 'RND3')
```


```{r, fig.height=7, fig.width=7}
obj1 = subset(sac_subset, species == 'Cow')
obj2 = subset(sac_subset, species == 'Sheep')

de1 = FindMarkersFast(obj1, group.by = 'lit_subtype', ident.1 = 'ON', ident.2 = 'OFF', p_val_adj_cutoff = 1, avg_log2FC_cutoff = 0)
de2 = FindMarkersFast(obj2, group.by = 'lit_subtype', ident.1 = 'ON', ident.2 = 'OFF', p_val_adj_cutoff = 1, avg_log2FC_cutoff = 0)

logfc_dat = merge(de1, de2, by = 'gene')
logfc_dat$label = logfc_dat$gene
logfc_dat$label[!logfc_dat$label %in% genes.of.interest] = NA

ScatterPlot(logfc_dat, 'avg_log2FC.x', 'avg_log2FC.y', labels = 'label', max.overlaps = Inf, label_func = geom_label_repel)
```

# TLE1 conservation

## Discovery in mouse P0

```{r, fig.height=4, fig.width=4}
peng_sac = readRDS("../../Species_Reference/PengAC_v3.rds")
peng_sac$annotated = factor(peng_sac$annotated, levels = c("OFF", "ON"))
# peng_sac$annotated = ifelse(peng_sac$seurat_clusters == 0, "ON", "OFF")
# Idents(peng_sac) = "annotated"
de.sac = FindMarkers(peng_sac, ident.1 = "ON", ident.2 = "OFF")
ExpressionScatter(peng_sac, x = "ON", y = "OFF", genes.to.label = c("FEZF1", "TENM3", "ZFHX3", "RND3", "TLE1"))
```

```{r, fig.height=4, fig.width=4}
ClusterFeaturePlot(peng_sac, features = "TLE1")
DotPlot2(peng_sac, features = "TLE1")
VlnPlot(peng_sac, features = "TLE1", group.by = 'annotated')
```

## Adult mouse

```{r, fig.height=4, fig.width=4}
Idents(successList$Mouse) = "lit_subtype"
ClusterFeaturePlot(successList$Mouse, features = "TLE1")
DotPlot2(successList$Mouse, features = "TLE1")
# successList$Mouse$lit_subtype = factor(successList$Mouse$lit_subtype, levels = c("OFF", "ON"))
VlnPlot(successList$Mouse, features = "TLE1", group.by = "lit_subtype")
```

## Other species

```{r, fig.height=12, fig.width=10}
ggarrange(plotlist = lapply(names(successList), function(species) {
  tryCatch(TitlePlot(VlnPlot(successList[[species]], features = c("TLE1"), group.by = "lit_subtype"), species), error=function(e) NULL)
  }), common.legend = TRUE, legend = 'right'
)
```

# Tenuerin-3 and latrophilin-2

## BCs: OFF BCs

```{r, fig.height=4, fig.width=7}
bc.ortho = readRDS('../../Ortho_Objects/vertebrateBC_BIGSEURAT.rds')
bc.ortho$species = factor(bc.ortho$species, levels = phylogenetic_order)
geneDotPlotFast(bc.ortho, gene = 'TENM3', group.by = 'NOG')
ggsave('../../figures/my_figs/bc-tenm3.pdf', width = 6, height = 3)
geneDotPlotFast(bc.ortho, gene = 'ADGRL2', group.by = 'NOG') # latrophilin 2
ggsave('../../figures/my_figs/bc-adgrl2.pdf', width = 6, height = 3)
```

How surprising is this? 

```{r}
BIGSEURAT = readRDS('../../Ortho_Objects/vertebrateAC_BIGSEURAT.rds')
gene2go = fread('../../reference_files/human_gene2go_symbol.csv')

neurotransmitter = c(gene2go[grepl("neurotransmitter", gene2go$GoTerm)]$Symbol %>% unique %>% sort, 'GAD1', 'GAD2')
connexins = sort(rownames(BIGSEURAT)[grepl('^GJA|^GJB|^GJC|^GJD|^GJE', rownames(BIGSEURAT))])
# adhesion = intersect(sort(str_split(fread('../../reference_files/KEGG_CELL_ADHESION_MOLECULES_CAMS.v2023.2.Hs.tsv')$KEGG_CELL_ADHESION_MOLECULES_CAMS[17], ',')[[1]]), rownames(BIGSEURAT))
subset(gene2go, Symbol == 'TENM3')
subset(gene2go, Symbol == 'ADGRL2')
adhesion = sort(intersect(c(gene2go[gene2go$GoID == 'GO:0005886']$Symbol %>% unique %>% sort), rownames(BIGSEURAT)))
adhesion
```


```{r, fig.height=5, fig.width=8, dpi=300, dev = 'CairoPNG'}
bc.ortho = readRDS('../../Ortho_Objects/vertebrateBC_BIGSEURAT.rds')
bc.conservation.summary = readRDS('../../Ortho_Objects/bc.conservation.summary.rds')
bc.ortho$species_NOG = paste0(bc.ortho$species, '-', bc.ortho$NOG)

# Remove genes with no expression
# datExpr = datExpr[goodSamplesGenes(t(datExpr))$goodGenes,]
# dim(datExpr)

# Keep only genes expressed robustly 
# good.genes = SelectFeatures(bc.ortho, group.by = 'NOG', features = rownames(datExpr))
# datExpr = datExpr[good.genes,]
# dim(datExpr)

# Keep only conserved genes 
good.genes = sort(bc.conservation.summary$gene[which(bc.conservation.summary$cor.score > 0.3)])

# test for anticorrelated cell adhesion molecules in BCs
# datExpr = LogAvgExpr(bc.ortho, group.by = 'species_NOG', assay = 'RNA', features = good.genes)
# dim(datExpr)

# Use species-normalized log-normalized and scaled expression values
matrix.list = readRDS("../../Ortho_Objects/BC_OT_expression_log1p_scaled.rds")
matrix.list.subset = matrix.list[good.genes]

# adhesion.cor = outer(matrix.list.subset, matrix.list.subset, function(x, y) {
#   print(x)
#   cor(as.vector(as.matrix(x)), as.vector(as.matrix(y)), use = 'complete.obs')
#   }) # cor(t(datExpr))

compute.cor <- function(matrix.list.subset) {
  # Initialize an empty matrix to store the correlation results
  cor_matrix <- matrix(NA, nrow = length(matrix.list.subset), ncol = length(matrix.list.subset), 
                       dimnames = list(names(matrix.list.subset), names(matrix.list.subset)))
  
  # Loop over all pairs of matrices in the list
  for (i in 1:length(matrix.list.subset)) {
    for (j in 1:length(matrix.list.subset)) {
      x <- matrix.list.subset[[i]]
      y <- matrix.list.subset[[j]]
      
      # Print the matrices being compared (optional)
      # print(list(x = x, y = y))
      
      # if(names(matrix.list.subset)[[i]] == 'B3GNT2' & names(matrix.list.subset)[[j]] == 'ABCA4') browser()
      
      # Compute the correlation between the vectorized matrices
      cor_matrix[i, j] <- cor(as.vector(as.matrix(x)), as.vector(as.matrix(y)), use = 'complete.obs')
    }
  }
  
  return(cor_matrix)
}

adhesion.cor = compute.cor(matrix.list.subset)
diag(adhesion.cor) = NA
adhesion.cor[upper.tri(adhesion.cor)] = NA
cor.df = reshape2::melt(adhesion.cor) %>% setNames(c('Gene1', 'Gene2', 'Correlation'))

cor.df$number = 1:nrow(cor.df)
cor.df = subset(cor.df, !is.na(Correlation)) %>% arrange(desc(Correlation))

# top10.pos = head(cor.df, 20) %>% pull(gene.combination)
# top10.neg = head(cor.df %>% arrange(Correlation), 20) %>% pull 

MultigeneDotPlot(bc.ortho, genes = c('ISL1', 'GRM6', 'GNB3'), group.by = 'NOG')
MultigeneDotPlot(bc.ortho, genes = c('XKR4', 'ZBTB20'), group.by = 'NOG')
MultigeneDotPlot(bc.ortho, genes = c('DSCAM', 'DSCAML1'), group.by = 'NOG')
MultigeneDotPlot(bc.ortho, genes = c('ZEB2', 'NTNG1'), group.by = 'NOG')

cor.df$gene.combination = paste(cor.df$Gene1, '-' , cor.df$Gene2)
# cor.df$rand.number = runif(nrow(cor.df))

compute_pvalues <- function(correlations, sample_sizes) {
  # Initialize an empty vector to store p-values
  p_values <- numeric(length(correlations))
  
  # Loop over the correlations and sample sizes to compute p-values
  for (i in 1:length(correlations)) {
    r <- correlations[i]
    n <- sample_sizes[i]
    
    # Compute t statistic
    t_value <- (r * sqrt(n - 2)) / sqrt(1 - r^2)
    
    # Compute p-value using the t-distribution with df = n - 2
    p_values[i] <- 2 * pt(-abs(t_value), df = n - 2)  # Two-tailed test
  }
  
  return(p_values)
}

cor.df$p.value = compute_pvalues(cor.df$Correlation, rep(181, nrow(cor.df)))
cor.df$p.adj = p.adjust(cor.df$p.value)
cor.df$rank = rank(abs(cor.df$Correlation))

ggscatter(cor.df, x = 'number', y = 'Correlation', label = 'gene.combination', font.label = c(2, "plain"), size = 1, repel = FALSE) + 
  theme(axis.text.x = element_blank(), plot.margin = unit(c(1,1,1,1), units = 'in'), axis.ticks = element_blank()) +
  scale_x_discrete(expand = expansion(mult = c(0.01,0.01)))+
  xlab('Gene combinations')+
  theme_dario()

top.genes = head(cor.df %>% arrange(p.adj), 50)

pdf('specificity_analysis/bc-cor-anticor-genes.pdf', height = 5, width = 6)
for(i in 1:nrow(top.genes)){
  print(MultigeneDotPlot(bc.ortho, genes = as.character(c(top.genes$Gene1[[i]], top.genes$Gene2[[i]])), group.by = 'NOG'))
}
dev.off()

cor.df = cor.df %>% arrange(p.adj)
saveRDS(cor.df, '../../Ortho_Objects/bc-cor-anticor-df.rds')
```

```{r, fig.height=3, fig.width=8, dpi=300, dev = 'CairoPNG'}
HistogramAnnotation(PrettyHistogram(cor.df$Correlation, xlab = 'Correlation'), 'p = 4.3e-34', subset(cor.df, gene.combination == 'TENM3 - ADGRL2')$Correlation, linetype = 'dashed')
```

String interaction

```{r}
alias.db = fread('~/Downloads/9606.protein.aliases.v12.0.txt')
gene.key = as.data.frame(alias.db[,c('#string_protein_id', 'alias')] %>% unique %>% arrange(alias))
subset(gene.key, alias == 'TENM3')
# nrow(unique(gene.key[,1]))
# nrow(unique(gene.key[,2]))
string.db = fread('~/Downloads/9606.protein.links.v12.0.txt')
head(string.db)

query_stringdb = function(gene1, gene2){
  # Not all proteins will be found
  df = subset(string.db, protein1 == subset(gene.key, alias == gene1)[1,1] & protein2 == subset(gene.key, alias == gene2)[1,1])
  df$gene1 = gene1
  df$gene2 = gene2
  if(nrow(df) > 0) df else data.frame(matrix(NA, nrow = 1, ncol = 5)) %>% setNames(colnames(df))
}

# All significant interactions (1589)
cor.df = readRDS('../../Ortho_Objects/bc-cor-anticor-df.rds')
cor.df.sig = subset(cor.df, p.adj < 1e-6)

# options(warn = 2)         # Treat warnings as errors
# options(error = recover)
string.res = do.call(rbind, apply(cor.df.sig, 1, function(x) {
  print(x['gene.combination'])
  query_stringdb(x['Gene1'], x['Gene2'])
  }))

cor.df.sig$Interaction = string.res$combined_score/1000
cor.df.sig$direction = ifelse(cor.df.sig$Correlation > 0, '+', '-')
cor.df.sig$gene.combination = gsub(' - ', '/', cor.df.sig$gene.combination)

ScatterPlot(cor.df.sig, x = 'Correlation', y = 'Interaction', r_pval = FALSE, lm = FALSE, 
            labels = 'gene.combination', fill = 'direction', ylab = 'Protein interaction score', max.overlaps = 10) + 
  scale_fill_manual(name = 'Direction', values = c('red', 'blue')) +
  geom_hline(yintercept = c(0.4, 0.7, 0.9), linetype = 'dashed', color = c('lightgrey', 'grey', 'black'))


top.genes = subset(cor.df.sig, Interaction > 0.7)
pdf('specificity_analysis/bc-cor-anticor-genes-interaction.pdf', height = 5, width = 6)
for(i in 1:nrow(top.genes)){
  print(MultigeneDotPlot(bc.ortho, genes = as.character(c(top.genes$Gene1[[i]], top.genes$Gene2[[i]])), group.by = 'NOG'))
}
dev.off()
```

```{r, fig.height=6, fig.width=20}
plot_grid(MultigeneDotPlot(bc.ortho, c('DSCAM', 'DSCAML1'), group.by = 'NOG', no.legend = TRUE), 
          MultigeneDotPlot(ac.ortho, c('DSCAM', 'DSCAML1'), no.legend = TRUE), 
          MultigeneDotPlot(rgc.ortho, c('DSCAM', 'DSCAML1'), group.by = 'NOG'), 
          nrow = 1, rel_widths = c(1,2,1.5), labels = c('BC', 'AC', 'RGC'))
```


```{r, fig.height=20, fig.width=5}
MultigeneDotPlot(bc.ortho, c('TENM3', 'ADGRL2', 'UNC5D', 'FLRT3', 'DSCAM', 'DSCAML1', 'SDK1', 'SDK2'), group.by = 'NOG')
```


```{r, fig.height=16, fig.width=20}
plot_grid(MultigeneDotPlot(bc.ortho, c('TENM1', 'TENM2', 'TENM3', 'TENM4', 'ADGRL1', 'ADGRL2', 'ADGRL3'), group.by = 'NOG', no.legend = TRUE), 
          MultigeneDotPlot(ac.ortho, c('TENM1', 'TENM2', 'TENM3', 'TENM4', 'ADGRL1', 'ADGRL2', 'ADGRL3'), no.legend = TRUE), 
          MultigeneDotPlot(rgc.ortho, c('TENM1', 'TENM2', 'TENM3', 'TENM4', 'ADGRL1', 'ADGRL2', 'ADGRL3'), group.by = 'NOG'), 
          nrow = 1, rel_widths = c(1,2,1.5), labels = c('BC', 'AC', 'RGC'))
```

Slit and ROBO

```{r, fig.height=16, fig.width=20}
bc.ortho = readRDS('../../Ortho_Objects/vertebrateBC_BIGSEURAT.rds')
ac.ortho = readRDS('../../Ortho_Objects/vertebrateAC_BIGSEURAT.rds')
rgc.ortho = readRDS('../../Ortho_Objects/vertebrateRGC_BIGSEURAT.rds')

plot_grid(MultigeneDotPlot(bc.ortho, c('SLIT1', 'SLIT2', 'SLIT3', 'ROBO1', 'ROBO2', 'ROBO3', 'ROBO4'), group.by = 'NOG', no.legend = TRUE), 
          MultigeneDotPlot(ac.ortho, c('SLIT1', 'SLIT2', 'SLIT3', 'ROBO1', 'ROBO2', 'ROBO3', 'ROBO4'), no.legend = TRUE), 
          MultigeneDotPlot(rgc.ortho, c('SLIT1', 'SLIT2', 'SLIT3', 'ROBO1', 'ROBO2', 'ROBO3', 'ROBO4'), group.by = 'NOG'), 
          nrow = 1, rel_widths = c(1,2,1.5), labels = c('BC', 'AC', 'RGC'))
```

## ACs

```{r, fig.height=10, fig.width=8}
ac.ortho = LoadAC()
MultigeneDotPlot(ac.ortho, c('TENM3', 'ADGRL2', 'DSCAM', 'DSCAML1'))
```

## SACs

```{r, fig.height=4, fig.width=10}
bc.ortho = readRDS('../../Ortho_Objects/vertebrateBC_BIGSEURAT.rds')
sac_subset = readRDS('../../Ortho_Objects/sac_subset_v4.rds')
sac_subset$subtype = factor(paste0(sac_subset$lit_subtype, ' SAC'), levels = c('ON SAC', 'OFF SAC'))

plot_grid(
  (geneDotPlotFast(sac_subset, gene = 'TENM3', group.by = 'subtype') + NoLegend() | 
  geneDotPlotFast(sac_subset, gene = 'ADGRL2', group.by = 'subtype') + NoLegend() + theme(axis.text.y = element_blank())), 
  (geneDotPlotFast(bc.ortho, group.by = 'NOG', gene = 'TENM3') + NoLegend() | 
    geneDotPlotFast(bc.ortho, group.by = 'NOG', gene = 'ADGRL2') + theme(axis.text.y = element_blank())), 
  nrow = 1, rel_widths = c(1,3), align = 'h', axis = 'bt'
  )
ggsave('../../figures/my_figs/tenm3-adgrl2-sac-bc.pdf', height=4, width=10)
```

## RGCs: broad expression

C10 is low TENM3 and projects to ON IPL
C18 is ON and low 
Check 25 and 26

Doesn't seem to be a big effect in RGCs

```{r, fig.height=12, fig.width=8}
rgc.ortho = readRDS('../../Ortho_Objects/vertebrateRGC_BIGSEURAT.rds')
MultigeneDotPlot(rgc.ortho, c('TENM3', 'ADGRL2', 'DSCAM', 'DSCAML1', 'DCC'), group.by = 'NOG')


mouse.rgc = readRDS('../../HahnObjects/RGCs/MouseRGC_integrated_v2.rds')
DotPlot2(mouse.rgc, features = 'TENM3')
```

# ON vs OFF RGCs

```{r, fig.height=4, fig.width=11}
rgc.prop = read.csv('../RGC/RGC subclasses - Sheet1.csv')
rgc.prop

on.types = setdiff(as.vector(str_split_fixed(ExtractString(c(rgc.prop$ON.sustained, rgc.prop$ON.transient), after = ' '), '/', 2)), '')
off.types = setdiff(ExtractString(c(rgc.prop$OFF.sustained, rgc.prop$OFF.transient), after = ' '), '')
on.types
off.types

mouse.rgc = readRDS('../../HahnObjects/RGCs/MouseRGC_integrated_v2.rds')
mouse.rgc$annotated2 = ExtractString(mouse.rgc$annotated, after = '_')
mouse.rgc$polarity = ifelse(mouse.rgc$annotated2 %in% on.types, 'ON', ifelse(mouse.rgc$annotated2 %in% off.types, 'OFF', 'other'))
mouse.rgc$annotated2 = factor(mouse.rgc$annotated2, levels = Metadata(mouse.rgc, 'polarity', 'annotated2')$annotated2)

TitlePlot(DotPlot3(mouse.rgc, features = c('TENM3', 'ADGRL2', 'DSCAM', 'DSCAML1', 'SDK1', 'SDK2', 'FLRT3', 'UNC5D'), group.by = 'annotated2') +
  geom_hline(yintercept = cumsum(as.vector(table(Metadata(mouse.rgc, 'polarity', 'annotated2')$polarity))) + 0.5),
  'OFF                             ON                                                      other                          ')
```

# Co-embedding with more cells

```{r}
BIGSEURAT = readRDS('../../Ortho_Objects/vertebrateAC-allcells-counts.rds')
BIGSEURAT = RenameCells(BIGSEURAT, add.cell.id = BIGSEURAT$species)
merged = readRDS('../../Ortho_Objects/sac-merged_v1.rds')
BIGSEURAT = RenameCells(BIGSEURAT, new.names = paste0(BIGSEURAT$species, '_', BIGSEURAT$Cell))
BIGSEURAT$lit_subtype = merged$lit_subtype[match(Cells(BIGSEURAT), Cells(merged))]
sac_subset = subset(BIGSEURAT, lit_subtype %in% c('ON', 'OFF'))
table(sac_subset$species)
saveRDS(sac_subset, '../../Ortho_Objects/sac_subset_v5.rds')
```


```{r}
cow_sheep = subset(sac_subset, species %in% c('Cow', 'Sheep'))
cow_sheep = ClusterSeurat(cow_sheep, integrate.by = 'species')
DimPlot(cow_sheep, group.by = 'lit_subtype')

rodents = subset(sac_subset, species %in% c('Mouse', 'Rhabdomys', 'Rat', 'Peromyscus', 'Squirrel'))
rodents = ClusterSeurat(rodents, integrate.by = 'species')
DimPlot(rodents, group.by = 'lit_subtype')
DefaultAssay(rodents) = 'integrated'
rodents = FindClusters(rodents, resolution = 0.01)
DimPlotLabeled(rodents)

# Add the primates
primates = subset(BIGSEURAT, species %in% c('Human', 'Macaque', 'Marmoset') & type == '8_SAC*')
primates = ClusterSeurat(primates, integrate.by = 'species')
DimPlotLabeled(primates)

# Coembedding fails
pr_rd = merge(primates, rodents)
pr_rd = ClusterSeurat(pr_rd, integrate.by = 'species')
DimPlot(pr_rd, group.by = 'lit_subtype')

sac_subset = ClusterSeurat(sac_subset, integrate.by = 'species')
DimPlot(sac_subset, group.by = 'lit_subtype')
DefaultAssay(sac_subset) = 'integrated'
sac_subset = FindClusters(sac_subset, resolution = 0.3)
DimPlotLabeled(sac_subset)
```

# LDA

```{r}
# -----------------------------------------------------------
# LDA on the iris dataset
# -----------------------------------------------------------

library(MASS)
library(ggplot2)
library(dplyr)
library(gridExtra)

# -----------------------------------------------------------
# 1. Load data
# -----------------------------------------------------------
df <- iris
head(df)

# -----------------------------------------------------------
# 2. Fit Linear Discriminant Analysis
# -----------------------------------------------------------
lda_model <- lda(Species ~ ., data = df)

# Variance explained by each LDA axis
prop_var <- lda_model$svd^2 / sum(lda_model$svd^2)
round(prop_var, 3)

# Project data into LDA space
lda_proj <- predict(lda_model)$x
lda_df <- data.frame(lda_proj, Species = df$Species)

# -----------------------------------------------------------
# 3. Plot LDA projection
# -----------------------------------------------------------
p_lda <- ggplot(lda_df, aes(LD1, LD2, color = Species)) +
  geom_point(size = 3, alpha = 0.8) +
  theme_minimal(base_size = 14) +
  labs(
    title = "LDA projection of iris data",
    x = paste0("LD1 (", round(prop_var[1] * 100, 1), "% var)"),
    y = paste0("LD2 (", round(prop_var[2] * 100, 1), "% var)")
  ) +
  scale_color_brewer(palette = "Dark2")

# -----------------------------------------------------------
# 4. Compare with PCA projection
# -----------------------------------------------------------
pca_res <- prcomp(df[, 1:4], scale. = TRUE)
pca_df <- data.frame(pca_res$x[, 1:2], Species = df$Species)

p_pca <- ggplot(pca_df, aes(PC1, PC2, color = Species)) +
  geom_point(size = 3, alpha = 0.8) +
  theme_minimal(base_size = 14) +
  labs(
    title = "PCA projection of iris data",
    x = paste0("PC1 (", round(summary(pca_res)$importance[2, 1] * 100, 1), "% var)"),
    y = paste0("PC2 (", round(summary(pca_res)$importance[2, 2] * 100, 1), "% var)")
  ) +
  scale_color_brewer(palette = "Dark2")

# -----------------------------------------------------------
# 5. Display side-by-side
# -----------------------------------------------------------
gridExtra::grid.arrange(p_pca, p_lda, ncol = 2)

```

```{r, fig.height=4, fig.width=10}
RunLDA(iris[,1:4], iris[,1:4], iris$Species)

rodents_minus = subset(rodents, species == 'Squirrel', invert = TRUE)
squirrel = subset(rodents, species == 'Squirrel')
squirrel.lda = RunLDA(rodents_minus@reductions$pca@cell.embeddings, squirrel@reductions$pca@cell.embeddings, rodents_minus$lit_subtype)

# Works marvelously
EvaluateModel(table(squirrel.lda[[2]]$class, squirrel$lit_subtype))

PrettyHistogram2(squirrel.lda[[2]]$x[,1])
```

## In the primate/rodent integrated space

```{r, fig.height=4, fig.width=10}
# Train on rodents
train_rodents = subset(pr_rd, species %in% c('Human', 'Macaque', 'Marmoset'), invert = TRUE)

# Test on primates
test_primates = subset(pr_rd, species %in% c('Human', 'Macaque', 'Marmoset'), invert = FALSE)

# Run LDA
rodent.primate.lda = RunLDA(train_rodents@reductions$pca@cell.embeddings, test_primates@reductions$pca@cell.embeddings, train_rodents$lit_subtype)
rodent.lda = rodent.primate.lda[[1]]
primate.lda = rodent.primate.lda[[2]]

PrettyHistogram2(rodent.lda$x[,1])

train_rodents$lda.class = rodent.lda$class
train_rodents$posterior = rodent.lda$posterior[,1]
train_rodents$ld1 = rodent.lda$x[,1]

test_primates$lda.class = primate.lda$class
test_primates$posterior = primate.lda$posterior[,1]
test_primates$ld1 = primate.lda$x[,1]
# hist(subset(test_primates, species %in% c('Human'))$ld1)
# hist(subset(test_primates, species %in% c('Macaque'))$ld1)

pseudo.mat = as.data.frame(primate.lda$x)
pseudo.mat[['LD2']] = rnorm(nrow(primate.lda$x))

test_primates[["lda"]] <- CreateDimReducObject(
  embeddings = as.matrix(pseudo.mat),
  key = "lda_",
  assay = DefaultAssay(test_primates)
)

PrettyHistogram2(subset(test_primates, species %in% c('Human'))$ld1)
PrettyHistogram2(subset(test_primates, species %in% c('Macaque'))$ld1)
PrettyHistogram2(subset(test_primates, species %in% c('Marmoset'))$ld1)

PrettyHistogram2(subset(test_primates, species %in% c('Human'))$posterior)
PrettyHistogram2(subset(test_primates, species %in% c('Macaque'))$posterior)
PrettyHistogram2(subset(test_primates, species %in% c('Marmoset'))$posterior)

DotPlot3(test_primates, features = c('TENM3', 'ZFHX3', 'RND3', 'FEZF1', 'FEZF2'), group.by = 'lda.class')

PrettyHistogram2(subset(train_rodents, lda.class == 'ON' & species %in% c('Rat'))$ld1, subset(train_rodents, lda.class == 'OFF' & species %in% c('Rat'))$ld1)
PrettyHistogram2(subset(test_primates, lda.class == 'ON')$ld1, subset(test_primates, lda.class == 'OFF')$ld1)

barplot(col_norm(table(test_primates$lda.class, test_primates$species)))
saveRDS(test_primates, '../../Ortho_Objects/test_primates-v1.rds')
```

Jitter them and make the hue proportional to the posterior

```{r, fig.height=3, fig.width=7}
test_primates$lda.class.filt = ifelse(test_primates$posterior < 0.2 | test_primates$posterior > 0.8, as.character(test_primates$lda.class), NA)
DimPlot(test_primates, split.by = 'species', group.by = 'lda.class.filt', reduction = 'lda')
FeaturePlot(test_primates, split.by = 'species', reduction = 'lda', features = 'TENM3', order = TRUE)
FeaturePlot(test_primates, split.by = 'species', reduction = 'lda', features = 'ZFHX3', order = TRUE)

dat.sum = cbind(as.data.frame(test_primates@reductions$lda@cell.embeddings), 
                lda.class = test_primates$lda.class, 
                posterior = test_primates$posterior, 
                species = test_primates$species)
dat.sum$score = abs(dat.sum$posterior-0.5) + 0.5 # transform so that its the posterior of the best class

table(test_primates$lda.class.filt, test_primates$species)
```


```{r, fig.height=2.5, fig.width=8}
ggscatter(dat.sum, 'lda_1', 'lda_2', facet.by = 'species', color = 'lda.class', alpha = 'score', size = 1, shape = 16) + 
  ylab('') + 
  xlab('LD 1')+
  ArialFont()+
  scale_alpha_continuous(name = 'Posterior', range = c(0, 0.8))+
  guides(color = "none")+
  theme(axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        # plot.subtitle = element_text(size = 20),
        axis.ticks = element_blank(), 
        legend.position = 'right',
        strip.background = element_blank(),  # removes the grey rectangle
        strip.text = element_text(color = "black", size = 12)) # keep the text) + 

ggsave('../../figures/my_figs/fig-primate-sac.pdf')
```

## One primate at a time? 

```{r}
LDA_on_primate = function(this.species){
  
  exclude = setdiff(c('Human', 'Macaque', 'Marmoset'), this.species)
  rodent_marmoset = subset(pr_rd, species %in% exclude, invert = TRUE)
  rodent_marmoset = ClusterSeurat(rodent_marmoset, integrate.by = 'species')
  DimPlotLabeled(rodent_marmoset, group.by = 'lit_subtype')
  
  # LDA
  rodents_minus = subset(rodent_marmoset, species == this.species, invert = TRUE)
  this_species = subset(rodent_marmoset, species == this.species)
  this.lda = RunLDA(rodents_minus@reductions$pca@cell.embeddings, this_species@reductions$pca@cell.embeddings, rodents_minus$lit_subtype)
  
  print(table(this.lda[[2]]$class)) # 65:35 ratio
  
  
  rodent.lda = this.lda[[1]]
  rodents_minus$lda.class = rodent.lda$class
  rodents_minus$posterior = rodent.lda$posterior[,1]
  rodents_minus$ld1 = rodent.lda$x[,1]
  
  primate.lda = this.lda[[2]]
  this.lda$lda.class = primate.lda$class
  this.lda$posterior = primate.lda$posterior[,1]
  this.lda$ld1 = primate.lda$x[,1]

  list(rodent.lda, primate.lda)
}

# Only 99 cells in mouse lemur

marmoset.lda = LDA_on_primate('Marmoset')
macaque.lda = LDA_on_primate('Macaque') 
human.lda = LDA_on_primate('Human') # gives more OFF than ON, not expected
```

## Projection into the rodent integrated space

```{r}
# Scale all data
loadings = rodents@reductions$pca@feature.loadings
primates = ScaleData(primates, features = rownames(primates))
scale.data = primates@assays$RNA@scale.data

# Subset to rodent genes (hopefully rodent genes are resistant to batch effects)
common.genes = rodents@assays$integrated@var.features
scale.data.common = scale.data[common.genes,]

# 2000 x 50
query.pca.embeddings <- Matrix::t(scale.data.common) %*% loadings
primates[["pca_rodent"]] <- CreateDimReducObject(
  embeddings = as.matrix(query.pca.embeddings),
  key = "PC_",
  assay = DefaultAssay(primates)
)

DimPlot(primates, reduction = 'pca_rodent', group.by = 'species')
DimPlot(rodents, reduction = 'pca', group.by = 'species') | DimPlot(rodents, reduction = 'pca', group.by = 'lit_subtype')
```

# Summarize with clade panels

## Primates

```{r, fig.height=3.5, fig.width=4}
BIGSEURAT = readRDS('../../Ortho_Objects/vertebrateAC-allcells-counts.rds')
BIGSEURAT = RenameCells(BIGSEURAT, add.cell.id = BIGSEURAT$species)
sac_subset = readRDS('../../Ortho_Objects/sac_subset_v5.rds')

primateSAC = subset(BIGSEURAT, species %in% species_metadata$Species[species_metadata$Clade == 'Primate'] & type == '8_SAC*')

# Balance species
primateSAC = DownsampleSeurat(primateSAC, 'species', size = 300)
primateSAC = ClusterSeurat(primateSAC, integrate.by = 'species')
PrettyUmap2(primateSAC, group.by = 'species', cols = species_palette3, 
            label = FALSE, show.legend = TRUE, pad = 0.4) + 
  LegendLowerLeft() 
saveRDS(primateSAC, '../../Ortho_Objects/primateSAC.rds')
```

## Rodents

```{r}
rodentSAC = subset(sac_subset, species %in% species_metadata$Species[species_metadata$Clade == 'Rodent'])

# Balance species
rodentSAC = DownsampleSeurat(rodentSAC, 'species', size = 1000)
rodentSAC = ClusterSeurat(rodentSAC, integrate.by = 'species')
DimPlot(rodentSAC, group.by = 'species', cols = species_palette3) | DimPlot(rodentSAC, group.by = 'lit_subtype')
saveRDS(rodentSAC, '../../Ortho_Objects/rodentSAC.rds')
```

## Laurasia

```{r, fig.height=4, fig.width=6}
# lauraSAC = subset(BIGSEURAT, species %in% species_metadata$Species[species_metadata$Clade == 'Laurasiatheria'] & type == '8_SAC*')
lauraSAC = subset(sac_subset, species %in% species_metadata$Species[species_metadata$Clade == 'Laurasiatheria'])
lauraSAC = MapBack(lauraSAC, sac_subset, group.by = 'lit_subtype')

# Remove ferret and pig (too few cells)
lauraSAC = subset(lauraSAC, species %in% c('Ferret', 'Pig'), invert = TRUE)

# Balance species
lauraSAC = DownsampleSeurat(lauraSAC, 'species', size = 1000)
lauraSAC = ClusterSeurat(lauraSAC, integrate.by = 'species', numPCs = 5)
DimPlot(lauraSAC, group.by = 'species', cols = species_palette3) | DimPlot(lauraSAC, group.by = 'lit_subtype')
saveRDS(lauraSAC, '../../Ortho_Objects/lauraSAC.rds')
```

# Other mammals

```{r}
otherSAC = subset(sac_subset, species %in% c('Opossum', 'TreeShrew'))

# Balance species
table(otherSAC$species)
otherSAC = DownsampleSeurat(otherSAC, 'species', size = 1000)
otherSAC = ClusterSeurat(otherSAC, integrate.by = 'species')
DimPlot(otherSAC, group.by = 'species', cols = species_palette3) | DimPlot(otherSAC, group.by = 'lit_subtype')
saveRDS(otherSAC, '../../Ortho_Objects/otherSAC.rds')
```

## Sauropsids 

```{r, fig.height=4, fig.width=6}
# squaSAC = subset(BIGSEURAT, species %in% species_metadata$Species[species_metadata$Clade == 'Squamate'] & type == '8_SAC*')
squaSAC = subset(sac_subset, species %in% species_metadata$Species[species_metadata$Clade == 'Sauropsid'])
# squaSAC = MapBack(squaSAC, sac_subset, group.by = 'lit_subtype')

# Balance species
table(squaSAC$species)
squaSAC = DownsampleSeurat(squaSAC, 'species', size = 500)
squaSAC = ClusterSeurat(squaSAC, integrate.by = 'species', numPCs = 10)
DimPlot(squaSAC, group.by = 'species', cols = species_palette3) | DimPlot(squaSAC, group.by = 'lit_subtype')
DimPlotLabeled(squaSAC)
saveRDS(squaSAC, '../../Ortho_Objects/sauroSAC.rds')
```


## Amphibians

```{r}
axolotl = readRDS('../../Species_Objects/AxolotlAC_v6.rds')
newt = readRDS('../../Species_Objects/NewtAC_v6.rds')

DotPlot3(axolotl, features = 'CHAT', group.by = 'type')
DotPlot3(newt, features = 'CHAT', group.by = 'type')

ax.SAC = subset(axolotl, type %in% c('gabaAC-1', 'gabaAC-17'))
newt.SAC = subset(newt, type %in% c('gabaAC-1', 'gabaAC-14'))

# Integrate with 12k common genes
common.genes = intersect(rownames(ax.SAC), rownames(newt.SAC))
amph.SAC = merge(ax.SAC, newt.SAC)
amph.SAC = SubsetSeuratGenes(amph.SAC, common.genes)
amph.SAC = ClusterSeurat(amph.SAC, integrate.by = 'species', numPCs = 20)
DimPlot(amph.SAC, group.by = 'species', cols = species_palette3) | DimPlotLabeled(amph.SAC)
DimPlot(amph.SAC, group.by = 'species', reduction = 'pca', cols = species_palette3)
FeaturePlot(amph.SAC, features = c(ON_SAC_markers, OFF_SAC_markers))

# Contamination in cluster 2
DotPlot3(amph.SAC, features = Genes(major_annotation))

amph.SAC = subset(amph.SAC, seurat_clusters == 2, invert = T)
DimPlot(amph.SAC, group.by = 'species', cols = species_palette3) | DimPlotLabeled(amph.SAC)
saveRDS(amph.SAC, '../../Ortho_Objects/amph.SAC.rds')
```


## Zebrafish only

```{r}
objectList = list()
# objectList$Killifish = readRDS('../../Species_Objects/Killifish_ncbiAC_v6.rds')
objectList$Zebrafish = readRDS('../../Species_Objects/ZebrafishAC_no_HCs_v6_converted.rds')
ze = readRDS('../../Full_Objects/Zebrafish_lyu_full_v2.rds')
# cbind(rownames(objectList$Zebrafish), rownames(ze))

# Rename zebrafish symbols
objectList$Zebrafish = RenameFeatures(objectList$Zebrafish, rownames(ze))
SAC.2 = subset(objectList$Zebrafish, lit_type %in% c('SAC') & sample %in% c('LD_14D'))
SAC.2$species = 'Zebrafish'
SAC.2 = ClusterSeurat(SAC.2) # Harmonize(SAC.2, batch = 'sample')
DimPlotLabeled(SAC.2) | DimPlotLabeled(SAC.2, group.by = 'sample')
saveRDS(SAC.2, '../../Ortho_Objects/zeSAC.rds')
```

## Zebrafish and killifish

Killifish struggles to split well

```{r}
objectList = list()
objectList$Killifish = readRDS('../../Species_Objects/Killifish_ncbiAC_v6.rds')
objectList$Zebrafish = readRDS('../../Species_Objects/ZebrafishAC_no_HCs_v6_converted.rds')
ze = readRDS('../../Full_Objects/Zebrafish_lyu_full_v2.rds')
# cbind(rownames(objectList$Zebrafish), rownames(ze))

# Rename zebrafish symbols
objectList$Zebrafish = RenameFeatures(objectList$Zebrafish, rownames(ze))
SAC.2 = subset(objectList$Zebrafish, lit_type %in% c('SAC') & sample %in% c('LD_14D'))
SAC.2$species = 'Zebrafish'
SAC.2 = ClusterSeurat(SAC.2) # Harmonize(SAC.2, batch = 'sample')
DimPlotLabeled(SAC.2) | DimPlotLabeled(SAC.2, group.by = 'sample')
 
# Annotate killifish SACs
# objectList$Killifish = AssignAnnotations(objectList$Killifish, list(SAC = 'AC-1'), use = 'annotated')
```


```{r}
SAC.1 = subset(objectList$Killifish, lit_type == 'SAC') # & animal %in% c('SRR27923005', 'SRR27923006')) # 11, 17
FeaturePlot(objectList$Killifish, features = 'sox2') | DimPlotLabeled(objectList$Killifish)
SAC.1$species = 'Killifish'
SAC.1 = ClusterSeurat(SAC.1, numPCs = 10)
# SAC.1 = Harmonize(SAC.1, batch = 'animal') # subset(SAC.1, animal %in% c('SRR27923003', 'SRR27923004')))
DimPlotLabeled(SAC.1) | DimPlotLabeled(SAC.1, group.by = 'animal')
FeaturePlot(SAC.1, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers), 'zfhx3b'))
# DotPlot3(SAC.1, features = Genes(major_annotation_pr_ze))
# DotPlot3(SAC.1, features = c('slc5a7a', 'chata', 'slc6a9', 'sox2', 'tenm3', 'fezf1'))

SAC.2 = subset(objectList$Zebrafish, lit_type %in% c('SAC')) # & sample %in% c('Control_2'))
SAC.2$species = 'Zebrafish'
SAC.2 = ClusterSeurat(SAC.2) #Harmonize(SAC.2, batch = 'sample')
DimPlotLabeled(SAC.2) | DimPlotLabeled(SAC.2, group.by = 'sample')
FeaturePlot(SAC.2, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers)))
# FeaturePlot(SAC.2, features = c('chata', 'slc5a7a', 'slc6a9'))
# DimPlotLabeled(objectList$Zebrafish) | FeaturePlot(objectList$Zebrafish, features = 'megf11')
# DimPlotLabeled(objectList$Zebrafish, group.by = 'lit_type')

# Load orthology graphs
ze2kf = readRDS('../cones/orthology_graphs/ze2kf.rds')
int = OrthotypeAnalysis2(SAC.2, list(SAC.1), homologyList = list(t(ze2kf)), 
                   group.by = 'type', # doesn't matter  
                   numPCs = 10,
                   downsample = 10000 # no downsampling
                   )
FeaturePlot(int, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers), 'zfhx3b'))
DimPlot(int, group.by = 'species', cols = species_palette3) | DimPlotLabeled(int)
```

## Shark and lamprey

Shark doesn't split

```{r}
objectList$Shark = readRDS('../../Species_Objects/SharkAC_v6.rds')
sh2ze = readRDS('../cones/orthology_graphs/sh2ze.rds')

SAC.1 = subset(objectList$Shark, type == 'gabaAC-3')
# SAC.1 = Harmonize(SAC.1, batch = 'animal')
SAC.1 = ClusterSeurat(SAC.1, integrate.by = 'animal')
DimPlotLabeled(SAC.1)
FeaturePlot(SAC.1, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers)))
```

## Zebrafish and lamprey

```{r}
objectList$Lamprey = readRDS('../../Full_Objects/Lamprey_Wang_full_v2.rds')
DotPlot3(objectList$Lamprey, features = 'CHAT', group.by = 'annotated')
pm2ze = readRDS('../cones/orthology_graphs/pm2ze.rds')
objectList$Lamprey = ConvertGeneSymbols2(objectList$Lamprey, pm2ze)

SAC.pm = subset(objectList$Lamprey, annotated %in% paste0('gabaAC-', 1:4))
SAC.pm = Harmonize(SAC.pm, batch = 'animal')
DimPlotLabeled(SAC.pm) | DimPlot(SAC.pm, group.by = 'animal')
FeaturePlot(SAC.pm, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers)), order = F)
DotPlot3(SAC.pm, features = Genes(major_annotation_pr_ze))
VlnPlot(SAC.pm, 'nFeature_RNA')

SAC.ze = subset(objectList$Zebrafish, lit_type %in% c('SAC') & sample %in% c('LD_14D'))
SAC.ze$species = 'Zebrafish'
SAC.ze = ClusterSeurat(SAC.ze) #Harmonize(SAC.2, batch = 'sample')
DimPlotLabeled(SAC.ze) | DimPlotLabeled(SAC.ze, group.by = 'sample')
FeaturePlot(SAC.ze, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers)))
```

Just lamprey

```{r}
SAC.pm.clean = subset(SAC.pm, seurat_clusters %in% c(4,5,6,7), invert = T)
SAC.pm.clean = RenameFeatures(SAC.pm.clean, SAC.pm.clean@misc$orig.features)
SAC.pm.clean$species = 'Lamprey'
DimPlotLabeled(SAC.pm.clean)
saveRDS(SAC.pm.clean, '../../Ortho_Objects/SAC.pm.clean.rds')
int = OrthotypeAnalysis2(SAC.ze, list(SAC.pm.clean), homologyList = list(pm2ze), 
                   group.by = 'type', # doesn't matter  
                   numPCs = 10,
                   downsample = 10000 # no downsampling
                   )
DimPlot(int, group.by = 'species', cols = species_palette3) | DimPlotLabeled(int)
FeaturePlot(int, features = c(tolower(ON_SAC_markers), tolower(OFF_SAC_markers), 'zfhx3b'))
```


